---
title: "Host Response"
author: "Jesse Garrett-Larsen"
date: "2024-10-15"
output: html_document
---
EEID 1a Heterogenetiy in Antibodies + Disease Response
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Packages}
rm(list=ls())
library(tidyverse)
library(dplyr)
library(glmmTMB)
library(effects)
library(AICcmodavg)
library(emmeans)
library(MASS)
library(DHARMa)
library(lme4)
library(purrr)
#install.packages("remotes")
#remotes::install_github("T-Engel/CValternatives")
library(CValternatives)
library(patchwork)
library(gridExtra)
library(car)
```

```{r Import Data}
source("dataCleaning_EEID1A.R")
```

```{r Format Theme}
#set colors for primary treatment
pri_colors <- c("#1B9E77", "#7570B3", "#D95F02")
#set color scheme secondary treatment
sec_colors <- c("#8C754B", "#77AB59", "#59A5D8", "#9F77D9", "#FA8072")
#set theme
theme_set(theme_bw())
```

```{r Raw Data Sample Sizes}
m.ab %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  )%>%
  modify_header(
    label ~ "**All Birds Before Removal**"
  )
```

####Generate Infection Data + Format For Analysis
Rules: A bird is considered infected if it had an eye score > 0.5 for more than one day, a pathogen load > 50 copies, or both
```{r Threshold Cutoffs}
m.ab$threshold_cutoff = 50
m.ab$seropos_cutoff = 0.061
m.ab$pathology_cutoff = 0
```

inf = if quantity (where any NA is replaced with 0) is greater than cutoff of 50 copies OR total eye score (where any NAs are ignored) is greater than 0.5 for more than one consecutive sampling day, or total eye score is greater or equal to 1 for at least one day, return 1, else return 0
```{r Generate Infection Data}
m.ab <- m.ab %>%
  # Sort data by band_number and dpi
  arrange(band_number, dpi) %>%
  # Replace NA values for quantity only
  mutate(quantity_clean = coalesce(quantity, 0)) %>%
  # Flag as infected if quantity > 50 on any single day
  mutate(quantity_flag = ifelse(quantity_clean > 50, 1, 0)) %>%
  # Check consecutive days for tes > 0.5, ignoring NAs for tes
  group_by(band_number) %>%
  mutate(
    tes_flag = ifelse(tes > 0.5 & !is.na(tes), 1, 0),
    tes_consecutive = tes_flag + lag(tes_flag, default = 0),
    tes_single_high = ifelse(tes >= 1 & !is.na(tes), 1, 0)
  ) %>%
  # Combine the three criteria: quantity > 50 OR tes > 0.5 on consecutive days OR tes >= 1 on any day
  mutate(inf = ifelse(quantity_flag == 1 | tes_consecutive > 1 | tes_single_high == 1, 1, 0)) %>%
  # Calculate inf_pri and inf_sec based on dpi
  ungroup() %>%
  group_by(band_number) %>%
  mutate(
    inf_pri = ifelse(any(inf == 1 & dpi < 42), 1, 0),
    inf_sec = ifelse(any(inf == 1 & dpi > 42), 1, 0)
  ) %>%
  # Un-group the data if needed
  ungroup() %>%
  # Select relevant columns (optional)
  select(-quantity_clean, -quantity_flag, -tes_flag, -tes_consecutive, -tes_single_high)


ggplot(m.ab, aes(x=dpi, y=tes, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_line(aes(groups=band_number))+
  geom_hline(yintercept=log10(50))+
facet_wrap(~inf)

ggplot(m.ab, aes(x=dpi, y=tes, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_line(aes(groups=band_number))+
  geom_hline(yintercept=log10(50))+
facet_wrap(~inf_pri)

ggplot(m.ab, aes(x=dpi, y=log10(quantity1), color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_line(aes(groups=band_number))+
  geom_hline(yintercept=log10(50))+
facet_wrap(~inf_sec)

```

```{r Omit Birds n=10}
prob_birds <- m.ab %>% filter(band_number %in% c(2274, 2514, 2469, 2520, 2494, 2452, 2562, 2419, 2434, 2505))%>%
  select(band_number, dpi, primary_treatment, secondary_dose, quantity, tes, l_eye_score, r_eye_score, elisa_od)

prob_birds %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(by=dpi)%>%
  modify_header(
    label ~ "**Birds to Remove**"
  )

#omit 2505 from data set for analysis as it was positive at quarantine (n=1)
m.ab <- m.ab %>%
  filter(band_number != 2505)

#omit 2452 and 2562 bc positive qpcr and they are shams (n=2)
m.ab <- m.ab %>% filter(!band_number %in% c(2452, 2562)) 

#omit sham birds with eye scores (0.5) (2419, 2434) (n=2)
m.ab <- m.ab %>% filter(!band_number %in% c(2419, 2434))

#omit sham bird with path load (2514) (n=1)
m.ab <- m.ab %>% filter(band_number != 2514)

m.ab %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  ) %>%
  modify_header(
    label ~ "**Sample Sizes for Primary Analysis**"
  )
#Final Primary Sample Sizes: Sham = 46; Low = 51; High = 53
```
```{r Idenity Birds to Remove For Secondary}
##Some birds did not fully recover after primary infection. These birds were included in primary analysis, but will be excluded from secondary analysis. I have identified these birds below, but *will not remove them until secondary analysis*.

#check birds that were not recovered by secondary infection: 2274, 2469, 2520, 2494 (2514 not recovered but omitted above)
not_recovered <- m.ab %>% 
  filter(band_number %in% c(2274, 2469, 2520, 2494))%>%
  dplyr::select(band_number, dpi, primary_treatment, secondary_dose, quantity, tes)

not_recovered %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  )%>%
  modify_header(
    label ~ "**Birds Not Recovered DPI 41**"
  )

#confirm quantity > cutoff at dpi 41
ggplot(not_recovered, aes(x=dpi, y=quantity, color=primary_treatment))+
  geom_point()+
  geom_path()+
  geom_hline(yintercept = 50, alpha=0.5)+
  geom_vline(xintercept = 41, alpha=0.5, lty = "dashed")+
  facet_wrap(~band_number)+
  ylim(min=0, max=700)
```

####Primary Challenge
```{r Final Primary Sample Size}
m.ab %>%
  dplyr::select(dpi, primary_treatment)%>%
  tbl_summary(
    by=dpi
  ) %>%
  modify_header(
    label ~ "**PRIMARY ANALYSIS FINAL SAMPLE SIZES**"
  )
```

>*Fifteen plasma samples were lost from DPPI 14 sampling and one from DPPI 41 sampling, resulting in reduced sample sizes for antibody analyses*. Of the 15 samples lost from DPPI 14, 14 were sham inoculated and one was inoculated with a low dose. The one plasma sample lost from DPPI 41 was from a bird inoculated with a low dose.
  >I went back and compared analyses with removal of these birds from the entire dataset (see *removal_analysis.R*) and found that *removing these 15 birds did not change     variability, effect sizes, or significance.* Additionally, *the models without full removal had lower AICc values*, indicating better model performance. Removal makes models perform worse.

```{r p.ab Data Frame With Only Primary Infection}
#data frame with only primary infection
p.ab <- m.ab %>%
  filter(dpi <=41)

p.ab %>%
  dplyr::select(dpi, primary_treatment, elisa_od)%>%
  tbl_summary(
    by=dpi
  ) %>%
  modify_header(
    label ~ "**PRIMARY ANALYSIS FINAL SAMPLE SIZES**"
  )

#Which samples did not have plasma samples?
p.ab.missing <- p.ab %>% 
  filter(dpi %in% c(-8, 14, 41) & is.na(elisa_od)) %>%
  dplyr::select(band_number, dpi, primary_treatment, secondary_dose, elisa_od)

# Remove elisa_od NAs 
p.ab <- p.ab %>% 
  filter(!(dpi %in% c(-8, 14, 41) & is.na(elisa_od)))
```

```{r Do baseline antibody levels differ between treatment groups?}
#Baseline antibody levels did not differ
lm0 <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi <0), family=Gamma())
summary(lm0)
plot(allEffects(lm0))
simulateResiduals(lm0, plot=T)

emm_r <- emmeans(lm0, ~primary_treatment)
pairs(emm_r)
```
```{r Neither mass nor sex predict antibody response}
p.ab$mass <- as.numeric(p.ab$mass)
max.ab <- p.ab %>%
  group_by(band_number, primary_treatment) %>%
  reframe(max_od = max(elisa_od, na.rm = TRUE),
            mass = as.numeric(mass[dpi == -8]),
            dpi = dpi,
            sex = sex,
            inf_pri = inf_pri)

#Neither Mass nor Sex predict max antibody response
lmm <- glm(max_od ~ mass * sex + primary_treatment, data=max.ab %>%filter(dpi == -8), family=Gamma())
summary(lmm)
plot(allEffects(lmm))
simulateResiduals(lmm, plot=T)

ggplot(max.ab%>%filter(dpi == -8), aes(x=mass, y=max_od, color=sex))+
  geom_boxplot(aes(x=mass, y=0.03, group=sex, color=sex), width=0.01)+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family="Gamma"))+
  facet_wrap(~primary_treatment, nrow=1)

#Neither Mass nor Sex predict infection status
lmi <- glm(inf_pri ~ mass + primary_treatment + sex, data=max.ab %>%filter(dpi == -8), family=binomial())
summary(lmi)
simulateResiduals(lmi, plot=T)

```

####Antibody Levels Across Primary Treatment
```{r Antibody Dataframe}
#I treat days post inoculation as a factor in these models and use band_number as a random effect
p.abt <- p.ab %>%
  filter(dpi %in% c(-8, 14, 41))
p.abt$dpi.f <- as.factor(p.abt$dpi)
```

```{r Do antibody levels vary across dpi? Does sex influence antibody response?}
#model comparison 
p1 <- glmmTMB(elisa_od~primary_treatment + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p2 <- glmmTMB(elisa_od~primary_treatment*dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p2.5 <- glmmTMB(elisa_od~primary_treatment*dpi.f + sex + (1|band_number), data=p.abt, family=Gamma(log))
p3<- glmmTMB(elisa_od~primary_treatment + sex + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p4<- glmmTMB(elisa_od~primary_treatment * sex + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p5 <- glmmTMB(elisa_od~1 + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))

aictab(cand.set=list(p1, p2, p2.5, p3, p4, p5), modnames=c("p1", "p2", "p2.5", "p3", "p4", "p5"))
# Interaction between primary_treatment and dpi.f without sex is best model by AICc

#model comparison VGAM
library(VGAM)
drop1(p1, test="Chisq") #primary_treatment and dpi.f contribute uniquely to explaining variation

drop1(p2, test="Chisq") #The interaction between primary_treatment and dpi.f > the effect of primary treatment on the outcome varies significantly by dpi levels

drop1(p2.5, test="Chisq") # sex is not significant

drop1(p3, test="Chisq") # sex is not significant but interaction b/t primary_treatment and dpi.f is better

drop1(p4, test="Chisq") # primary_treatment*sex is not significant

drop1(p5, test="Chisq") # better with primary_treatment
```


```{r}
library(GGally)
ggpairs(p.abt[, c("primary_treatment", "dpi.f", "elisa_od")])
```

```{r Primary Treatment Antibody Final Model}
#make Sham the reference category
p.abt$primary_treatment <- relevel(p.abt$primary_treatment, ref = "Sham")
#final model
library(lme4)
lm1 <- glmmTMB(elisa_od~ primary_treatment*dpi.f + (1|band_number), data=p.abt, family=Gamma(log)) 

#fully interactive does not converge with inverse link- only log link
summary(lm1)
#intercept > expected log-mean elisa_od (-3.074 ~ 0.046) looks good

car::Anova(lm1, type = "III")
simulateResiduals(lm1, plot=T) #but residuals look terrible with log link
hist(resid(lm1))
simulation_output <- simulateResiduals(fittedModel = lm1, n = 1000)

# Plot residuals
plotSimulatedResiduals(simulation_output, rank = TRUE)
#dispersion test
testDispersion(simulation_output)

p.abt$elisa_od_log <- log(p.abt$elisa_od)
hist(p.abt$elisa_od_log)
lm1log <- glmmTMB(elisa_od_log ~ primary_treatment * dpi.f + (1|band_number), data=p.abt)
summary(lm1log)
simulateResiduals(lm1log, plot=T)

emm_results <- emmeans(lm1, ~ primary_treatment|dpi.f, scale="response")
pairwise <- pairs(emm_results, adjust="tukey")
(summary(pairwise))

pairwise_df <- as.data.frame(pairwise)


ggplot(pairwise_df, aes(x = contrast, y = estimate, shape=dpi.f)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Pairwise Comparisons of Primary Treatments within Each dpi.f",
       x = "Comparison",
       y = "Estimated Difference") +
    coord_flip() +  # Flip the coordinates for better readability
  theme_minimal()
```

```{r What if I use a GAM?}
library(mgcv)
p.abt_clean <- p.abt %>% filter(!is.na(elisa_od))

#adapting previous model
gam_model <- gam(elisa_od ~ primary_treatment * dpi.f +                 # Categorical treatment as a fixed main effect
                  s(band_number, bs="re", k=100),             # Random effect for band_number
                  family = Gamma(link = "log"), 
                  data = p.abt_clean)

gam_model_k50 <- gam(elisa_od ~ primary_treatment * dpi.f +                 # Categorical treatment as a fixed main effect
                  s(band_number, bs="re", k=50),             # Random effect for band_number
                  family = Gamma(link = "log"), 
                  data = p.abt_clean)

AIC(gam_model, gam_model_k50)

summary(gam_model)
gam.check(gam_model)

plot(gam_model, pages = 1, all.terms = TRUE, shade = TRUE)

```


```{r Model Predictions lm1}
#specify model to predict
mod.pred <- lm1
#Model predictions
dat.new=expand.grid(primary_treatment=unique(p.abt$primary_treatment),
                    elisa_od = unique(p.abt$elisa_od),
                    dpi.f = unique(p.abt$dpi.f))
#generate yhat values
dat.new$yhat = predict(mod.pred, type="response", newdata=dat.new, re.form=NA)
#generate SEM
preds = predict(mod.pred, type="link", newdata=dat.new, se.fit=TRUE, re.form=NA)
dat.new = cbind(dat.new, preds)

ilink <- family(mod.pred)$linkinv
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))
head(dat.new)
```

```{r Format data for Variability in Antibodies for individuals who were infected}
head(p.ab)
#add small constant to total eye score for calculation of variability
p.ab$tes.new <- p.ab$tes + .01
#new df with only individuals that were infected any time during primary challenge
p.ab$dpi.f <- as.factor(p.ab$dpi)
p.abi <- p.ab %>%
  filter(inf_pri == 1)%>%
  dplyr::select(dpi.f, dpi, primary_treatment, tes.new, elisa_od, quantity1, sex, band_number, inf_pri)

#df with all individuals during primary challenge
p.aba <-  p.ab %>%
  dplyr::select(dpi.f, dpi, primary_treatment, tes.new, elisa_od, quantity1, sex, band_number, inf_pri)
```

I calculate variability the same way across analyses in this experiment. I use the below functions calculate_cv, calculate_pv, and calculate_v2. I then bootstrap the original data by resampling the data with replacement (each bootstrap sample is created by randomly re-sampling all points in a treatment group - some points may appear multiple times in the same bootstrap sample, while others may be excluded). The 95% confidence intervals are then calculated from the bootstrap distribution corresponding to the 2.5th and 97.5th percentiles.

These confidence intervals represent the confidence that we have that the observed variability metric is representative of the population as a whole.

In many of the later tests, the sample size is very low, therefore the confidence intervals are very wide.
```{r Calculate Variability in Antibodies for individuals who were infected}
# Define functions for calculating CV, PV, and V2
calculate_cv <- function(data) {
  return(sd(data) / mean(data))
}

calculate_pv <- function(data) {
  return(PV(data))
}

calculate_v2 <- function(data) {
  mean_x <- mean(data)
  sd_x <- sd(data)
  v2 <- sd_x^2 / (sd_x^2 + mean_x^2)
  return(v2)
}

# Set the number of bootstrap replicates
n_boot <- 1000

# Pipeline to calculate CV, PV, V2, and their 95% confidence intervals
a.cv <- p.aba %>%
  group_by(dpi.f, primary_treatment) %>%
  drop_na(elisa_od) %>%
  dplyr::reframe(
    # Calculate means and standard deviations
    elisa_od = elisa_od,
    mean_od = mean(elisa_od),
    bird_sd = sd(elisa_od),
    band_number = band_number,
    bird_cv = calculate_cv(elisa_od),   # Calculate CV for the original data
    bird_pv = calculate_pv(elisa_od),   # Calculate PV for the original data
    bird_v2 = calculate_v2(elisa_od),   # Calculate V2 for the original data
    
    # Bootstrap for each metric
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(elisa_od, replace = TRUE)))),#1000 new CV calculations from sampling the data 10 times randomly
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(elisa_od, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(elisa_od, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for CV, PV, and V2
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)), #lower confidence interval (2.5th percentile)
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)), #upper confidence interval (97.5th percentile)
    
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975)),
    
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  select(-cv_bootstrap, -pv_bootstrap, -v2_bootstrap) %>%
  ungroup()

#add to df and format names
p.aba.m <- left_join(p.aba, a.cv, by=c("dpi.f", "primary_treatment", "band_number"))
 p.aba.m <- p.aba.m %>%
   select(-elisa_od.y)
 p.aba.m$elisa_od <- p.aba.m$elisa_od.x
# p.abi.m$primary_treatment <- p.abi.m$primary_treatment.x
# p.abi.m <- p.abi.m %>%
#   select(-elisa_od.x, -primary_treatment.x)

#just days samples were taken
p.aba.m <- p.aba.m %>%
  filter(dpi.f %in% c(-8, 14, 41))

summary_tibble <- a.cv %>%
  group_by(dpi.f, primary_treatment) %>%
  summarize(
    mean_bird_cv = mean(bird_cv, na.rm = TRUE),    # Calculate the mean of bird_cv for each group
    mean_bird_sd = mean(bird_sd, na.rm = TRUE),    # Calculate the mean of bird_sd for each group
    mean_bird_pv = mean(bird_pv, na.rm = TRUE),    # Calculate the mean of bird_pv for each group
    avg_elisa_od = mean(elisa_od, na.rm = TRUE),# Calculate the average elisa_od for each group
      lower_ci_pv = mean(pv_lower_ci),
    upper_ci_pv = mean(pv_upper_ci),
    n_individuals = n_distinct(band_number) #calculate number of individuals in each group
  ) %>%
  as_tibble()
print(summary_tibble)
#write.csv(summary_tibble, "/Users/jesse/Documents/GitHub/EEID_1A_Mechanistic_Link/EEID_1A_Antibody_Analysis_files/ELISA Variability Primary/ab_var_not_removed.csv", row.names=FALSE)
```


```{r How does bootstrapping work? Bootstrap Distributions for eyescores}
# Pipeline to calculate CV, PV, V2, and their 95% confidence intervals
n_boot1 <- 1000
a.cv <- p.aba %>%
  group_by(dpi.f, primary_treatment) %>%
  drop_na(elisa_od) %>%
  dplyr::reframe(
    # Calculate means and standard deviations
    elisa_od = elisa_od,
    mean_od = mean(elisa_od),
    bird_sd = sd(elisa_od),
    band_number = band_number,
    dpi.f = dpi.f,
    bird_cv = calculate_cv(elisa_od),   # Calculate CV for the original data
    bird_pv = calculate_pv(elisa_od),   # Calculate PV for the original data
    bird_v2 = calculate_v2(elisa_od),   # Calculate V2 for the original data
    
    # Bootstrap for each metric
    cv_bootstrap = list(replicate(n_boot1, calculate_cv(sample(elisa_od, replace = TRUE)))),#1000 new CV calculations from sampling the data 10 times randomly
    pv_bootstrap = list(replicate(n_boot1, calculate_pv(sample(elisa_od, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot1, calculate_v2(sample(elisa_od, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for CV, PV, and V2
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)), #lower confidence interval (2.5th percentile)
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)), #upper confidence interval (97.5th percentile)
    
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975)),
    
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  #select(-cv_bootstrap, -pv_bootstrap, -v2_bootstrap) %>%
  ungroup()


# Unnest bootstrap samples into a tidy format
bootstrap_long <- a.cv %>%
  select(dpi.f, primary_treatment, cv_bootstrap, pv_bootstrap, v2_bootstrap) %>%
  pivot_longer(
    cols = c(cv_bootstrap, pv_bootstrap, v2_bootstrap),
    names_to = "metric",
    values_to = "bootstrap_values"
  ) %>%
  unnest(cols = c(bootstrap_values))

# Add descriptive labels for metrics
bootstrap_long <- bootstrap_long %>%
  mutate(
    metric = case_when(
      metric == "cv_bootstrap" ~ "Coefficient of Variation (CV)",
      metric == "pv_bootstrap" ~ "Proportional Variance (PV)",
      metric == "v2_bootstrap" ~ "V2"
    )
  )

ggplot(bootstrap_long, aes(x = bootstrap_values, fill = primary_treatment)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ metric, scales = "free") +
  scale_fill_manual(values = c(pri_colors)) +
  labs(
    title = "Bootstrap Distributions of Variability Metrics",
    x = "Metric Value",
    y = "Density",
    fill = "Treatment"
  ) +
  theme_minimal()


ggplot(bootstrap_long, aes(x = bootstrap_values, fill = primary_treatment)) +
  geom_histogram(binwidth = 0.025, position = "identity", alpha = 0.5, color="black") +
  facet_grid(~ metric ~dpi.f) +
  scale_fill_manual(values = c(pri_colors)) +
  labs(
    title = "Bootstrap Distributions of Variability Metrics",
    x = "Metric Value",
    y = "Frequency",
    fill = "Treatment"
  ) +
  theme_minimal()
```

```{r}
ggplot(data = p.aba.m, aes(fill = primary_treatment)) +
  # Original jittered data points for elisa_od
  #geom_jitter(size = 1.5, alpha = 1, width = 0.15, height = 0, shape=1)+
  #geom_density(aes(x=elisa_od), alpha=0.1)+
  geom_histogram(aes(x=elisa_od), binwidth = 0.01, position = "identity", alpha=0.5, color="black")+
  scale_fill_manual(values = c(pri_colors))+
  facet_grid(~primary_treatment~dpi.f)

ggplot(data = p.aba.m, aes(x=dpi, y=elisa_od, color = primary_treatment)) +
  # Original jittered data points for elisa_od
  geom_jitter(size = 1.5, width = 0.15, height = 0, alpha=0.5)+
  scale_color_manual(values = c(pri_colors))+
  facet_grid(~primary_treatment)
```

```{r Antibodies and Variability Figure 2}
#overlay PV and elisa_od
#FIGURE 2 combined
ggplot(data = p.aba.m, aes(x = dpi.f, y = elisa_od, color = primary_treatment)) +
  # Original jittered data points for elisa_od
  geom_jitter(size = 1.5, alpha = 1, width = 0.15, height = 0, shape=1) +
  
  # Line and points with error bars for predicted values
  geom_path(data = dat.new, aes(x = dpi.f, y = yhat, group = primary_treatment,
                                color = primary_treatment), alpha=0.75) +
  geom_errorbar(data = dat.new, aes(ymin = Lower, ymax = Upper, x = dpi.f, y = yhat),
                color = "black", width = 0.1) +
  geom_point(data = dat.new, aes(x = dpi.f, y = yhat, color = primary_treatment),
             size = 2, alpha = 0.75, shape = 16) +
  geom_point(data = dat.new, aes(x = dpi.f, y = yhat),
             color = "black", size = 2, alpha = 1, shape = 1, stroke = 0.1) +
  
  # V2 (CV^2) points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, fill = primary_treatment, shape = "PV"),
             size = 3, alpha = 0.2) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci,
                                 group = primary_treatment, color = primary_treatment),
                width = 0.05, alpha = 0.02) +
  geom_path(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment,
                             group = primary_treatment), lty = "dashed", alpha=0.2) +

  # Labels and axis
  labs(y = "Anti-MG IgY Antibody Levels [OD]", 
       x = "Days Post Primary Inoculation", 
       shape = "Variability Metric", 
       color = "Primary Treatment", 
       fill = "Primary Treatment") +
  
  # Custom color and fill scales
  scale_color_manual(values = c(pri_colors)) +
  scale_fill_manual(values = c(pri_colors)) +

  # Custom shape scale for V2 (CV^2)
  scale_shape_manual(
    name = "Variability Metric",
    values = c("PV" = 17),  # Set shape type for CV^2
    labels = c("PV" = "PV")
  ) +
  
  # Theme settings
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "black", size = 15, face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold", size = 15),
    axis.text.y.right = element_text(color = "black", face = "bold", size = 15),
    axis.title.y.right = element_text(color = "black", size = 20, face = "bold"),
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    legend.position = "top",
    legend.direction = "horizontal"
  )


#FIGURE 2 raw data + predictions
f2.raw<-ggplot(data = p.aba.m, aes(x = dpi.f, y = elisa_od, color = primary_treatment)) +
  # Original jittered data points for elisa_od
  geom_jitter(size = 1.5, alpha = 0.5, width = 0.15, height = 0, shape=16) +
  
  # Line and points with error bars for predicted values
  geom_path(data = dat.new, aes(x = dpi.f, y = yhat, group = primary_treatment,
                                color = primary_treatment), alpha=0.75) +
  geom_errorbar(data = dat.new, aes(ymin = Lower, ymax = Upper, x = dpi.f, y = yhat),
                color = "black", width = 0.01) +
  geom_point(data = dat.new, aes(x = dpi.f, y = yhat, color = primary_treatment),
             size = 2, alpha = 0.75, shape = 16) +
  geom_point(data = dat.new, aes(x = dpi.f, y = yhat),
             color = "black", size = 2, alpha = 1, shape = 1, stroke = 0.1) +
  
  # Labels and axis
  labs(y = "Anti-MG IgY Antibody Levels [OD]", 
       x = "Days Post Primary Inoculation", 
       shape = "Variability Metric", 
       color = "Primary Treatment", 
       fill = "Primary Treatment") +
  
  # Custom color and fill scales
  scale_color_manual(values = c(pri_colors)) +
  scale_fill_manual(values = c(pri_colors)) +
  
  # Theme settings
  theme_bw() +
  theme(
    axis.title.y = element_text(color = "black", size = 15, face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold", size = 15),
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    legend.position = "top",
    legend.direction = "horizontal"
  )

f2.raw

#Figure 2 var only
f2.pv<-ggplot(data = p.aba.m, aes(x = dpi.f, y = elisa_od, color = primary_treatment)) +
  # Original jittered data points for elisa_od
  #geom_jitter(aes(shape=primary_treatment), size = 1.5, alpha = 1, width = 0.15, height = 0) +
  
  # PV
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, fill = primary_treatment),
             size = 1.5, alpha = 0.2, shape=17) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci,
                                 group = primary_treatment, color = primary_treatment),
                width = 0.01, alpha = 0.02) +
  geom_path(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment,
                             group = primary_treatment), lty = "dashed", alpha=0.75) +

  # Labels and axis
  labs(y = "Antibody PV", 
       x = "Days Post Primary Inoculation", 
       shape = "Variability Metric", 
       color = "Primary Treatment", 
       fill = "Primary Treatment") +
  
  # Custom color and fill scales
  scale_color_manual(values = c(pri_colors)) +
  scale_fill_manual(values = c(pri_colors)) +
  
  # Theme settings
  theme_bw() +
  theme(
    axis.title.y = element_text(color = "black", size = 15, face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold", size = 15),
    axis.text.y.right = element_text(color = "black", face = "bold", size = 15),
    axis.title.y.right = element_text(color = "black", size = 20, face = "bold"),
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    legend.position = "top",
    legend.direction = "horizontal"
  )

f2.pv

plot(f2.raw + f2.pv)
```

```{r Antibodies and Variability}
#Just variability
ggplot() +
  # CV points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_cv, color = primary_treatment, shape = "CV"), 
             size = 3, position = position_dodge(width = 0.6), alpha = 0.5) +
  geom_path(data = a.cv, aes(x = dpi.f, y = bird_cv, color = primary_treatment), 
            position = position_dodge(width = 0.6)) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = cv_lower_ci, ymax = cv_upper_ci,
                                 color = primary_treatment, shape = "CV",), lty="dashed",
                width = 0.15, position = position_dodge(width = 0.6), alpha=0.05) +
  
  # PV points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment, shape = "PV"), 
             size = 3, position = position_dodge(width = 0.3), alpha = 0.5) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, 
                                 color = primary_treatment, shape = "PV"), lty="dashed",
                width = 0.15, position = position_dodge(width = 0.3), alpha=0.05) +
  
  # V2 points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_v2, color = primary_treatment, shape = "V2"), 
             size = 3, position = position_dodge(width = 0.9), alpha = 1) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, 
                                 color = primary_treatment, shape = "V2"), 
                width = 0.15, position = position_dodge(width = 0.9)) +
  
  # Custom color and fill scales
  scale_color_manual(values = c(pri_colors)) +
  scale_fill_manual(values = c(pri_colors)) +
  
  # Shape legend to distinguish metrics
  scale_shape_manual(
    name = "Variability Metric",
    values = c("V2" = 16, "CV" = 0, "PV" = 2),
    labels = c("V2" = "CV^2", "CV" = "CV", "PV" = "PV")
  ) +
  
  # Labels and theme
  labs(y = "Variability Metric Value", 
       x = "Days Post Primary Inoculation", 
       color = "Primary Treatment") +
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "black", size = 20, face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold", size = 15),
    axis.text.y.right = element_text(color = "black", face = "bold", size = 15),
    axis.title.y.right = element_text(color = "black", size = 20, face = "bold"),
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    legend.position = "top",
    legend.direction = "horizontal"
  )#+
  coord_flip()

```

```{r CV for Gamma Distributed Antibodies}
library(MASS)

fit_ab <- p.aba %>%
  filter(dpi.f %in% c(-8, 14, 41))%>%
  group_by(dpi.f, primary_treatment) %>%
  summarize(
    shape = fitdistr(na.omit(elisa_od), "gamma")$estimate["shape"],
    rate = fitdistr(na.omit(elisa_od), "gamma")$estimate["rate"]
  ) %>%
  ungroup()

fit_ab <- fit_ab %>%
  mutate(cv = 1 / sqrt(shape))

ggplot(fit_ab, aes(x=dpi.f, y=cv, color=primary_treatment))+
  geom_point()+
  scale_color_manual(values=c(pri_colors))


# Function to calculate CV for gamma distribution


library(MASS)
library(dplyr)
library(purrr)

# Function to calculate gamma CV
calculate_gamma_cv <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return(1 / sqrt(shape))
    }
  }
  return(NA)
}

# Function to calculate gamma V2
calculate_gamma_V2 <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return((1 + shape)^(-1/2))
    }
  }
  return(NA)
}

# Set the number of bootstrap replicates
n_boot <- 1000

# Pipeline to calculate gamma CV, gamma V2, and 95% CI
a.gcv <- p.aba %>%
  group_by(dpi.f, primary_treatment) %>%
  drop_na(elisa_od) %>%
  dplyr::reframe(
    # Original data calculations
    elisa_od = elisa_od,
    mean_od = mean(elisa_od, na.rm = TRUE),
    bird_sd = sd(elisa_od, na.rm = TRUE),
    band_number = band_number,
    
    bird_gamma_cv = calculate_gamma_cv(elisa_od),   # Gamma CV for original data
    bird_gamma_V2 = calculate_gamma_V2(elisa_od),   # Gamma V2 for original data

    # Bootstrap for gamma CV and V2 separately
    gamma_cv_bootstrap = list(replicate(n_boot, calculate_gamma_cv(sample(elisa_od, replace = TRUE)))),
    gamma_V2_bootstrap = list(replicate(n_boot, calculate_gamma_V2(sample(elisa_od, replace = TRUE))))
  ) %>%
  # Extract gamma CV and V2 bootstrap estimates and calculate 95% confidence intervals
  mutate(
    gamma_cv_lower_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_cv_upper_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),

    gamma_V2_lower_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_V2_upper_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  select(-gamma_cv_bootstrap, -gamma_V2_bootstrap) %>%
  ungroup()

# Add to the main dataset and calculate group summaries
p.aba.g <- left_join(p.aba, a.gcv, by = c("dpi.f", "primary_treatment", "band_number")) %>%
  select(-elisa_od.y) %>%
  mutate(elisa_od = elisa_od.x) %>%
  select(-elisa_od.x) %>%
  filter(dpi.f %in% c(-8, 14, 41))

summary_tibble_gamma <- a.gcv %>%
  group_by(dpi.f, primary_treatment) %>%
  summarize(
    mean_bird_gamma_cv = mean(bird_gamma_cv, na.rm = TRUE),  # Mean gamma CV for each group
    mean_bird_gamma_V2 = mean(bird_gamma_V2, na.rm = TRUE),  # Mean V2 for each group
    mean_bird_sd = mean(bird_sd, na.rm = TRUE),              # Mean of bird_sd
    avg_elisa_od = mean(elisa_od, na.rm = TRUE),             # Mean elisa_od for each group
    n_individuals = n_distinct(band_number)                  # Number of individuals in each group
  ) %>%
  as_tibble()

print(summary_tibble_gamma)



ggplot() +
  # Plotting mean gamma CV with error bars
  geom_point(data = summary_tibble_gamma, aes(x = dpi.f, y = mean_bird_gamma_cv, color = primary_treatment, shape = "Gamma CV"), size = 3, position = position_dodge(width = 0.25)) +
  geom_errorbar(data = a.gcv, aes(x = dpi.f, y = bird_gamma_cv, ymin = gamma_cv_lower_ci, ymax = gamma_cv_upper_ci, color = primary_treatment), linetype="dotted", position = position_dodge(width = 0.25), width = 0.2) +

  # Plotting mean gamma V2 with error bars
  geom_point(data = summary_tibble_gamma, aes(x = dpi.f, y = mean_bird_gamma_V2, color = primary_treatment, shape = "Gamma V2"), size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = a.gcv, aes(x = dpi.f, y = bird_gamma_V2, ymin = gamma_V2_lower_ci, ymax = gamma_V2_upper_ci, color = primary_treatment), position = position_dodge(width = 0.5), width = 0.2, linetype = "dashed") +

  # Plotting mean PV with error bars
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment, shape = "PV"), size = 3, position = position_dodge(width = 0.75)) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, color = primary_treatment), linetype = "solid", width = 0.2, alpha = 0.05, position = position_dodge(width = 0.75)) +

  # Custom colors
  scale_color_manual(values = pri_colors) +
  
  # Define shapes for each metric in the legend
  scale_shape_manual(values = c("Gamma CV" = 21, "Gamma V2" = 22, "PV" = 16)) +

  # Labels
  labs(y = "Gamma CV / Gamma V2 / PV", x = "DPI", title = "Gamma CV, Gamma V2, and PV of Antibodies by DPI and Treatment", shape = "Variability Metric") +
  
  # Legends and themes
  theme_minimal() 
```

Histograms of antibodies to compare with susceptibility distributions from Hawley et al., 2024 Figure 3
```{r}
ggplot(data = p.aba.m, aes(x = elisa_od, fill = primary_treatment)) +
  geom_histogram()+
  facet_grid(~dpi ~primary_treatment)+
  scale_fill_manual(values=pri_colors)
```


```{r Calculate Variability in eye score}
#eye score
e.cv <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri == 1)%>%
  reframe(
    bird_cv = calculate_cv(tes.new),
    bird_sd = sd(tes.new),
    bird_pv = calculate_pv(tes.new),
    bird_v2 = calculate_v2(tes.new),
    dpi.f = dpi.f,
    tes.new = tes.new,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(tes.new, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(tes.new, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(tes.new, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Remove bootstrap columns
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()
```

```{r}
e.cv <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri == 1)%>%
  reframe(
    bird_cv = calculate_cv(tes.new),
    bird_sd = sd(tes.new),
    bird_pv = calculate_pv(tes.new),
    bird_v2 = calculate_v2(tes.new),
    dpi.f = dpi.f,
    tes.new = tes.new,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(tes.new, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(tes.new, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(tes.new, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Remove bootstrap columns
  #select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()

# Unnest bootstrap samples into a tidy format
bootstrap_long <- e.cv %>%
  select(dpi.f, primary_treatment, cv_bootstrap, pv_bootstrap, v2_bootstrap) %>%
  pivot_longer(
    cols = c(cv_bootstrap, pv_bootstrap, v2_bootstrap),
    names_to = "metric",
    values_to = "bootstrap_values"
  ) %>%
  unnest(cols = c(bootstrap_values))

# Add descriptive labels for metrics
bootstrap_long <- bootstrap_long %>%
  mutate(
    metric = case_when(
      metric == "cv_bootstrap" ~ "CV",
      metric == "pv_bootstrap" ~ "PV",
      metric == "v2_bootstrap" ~ "V2"
    )
  )

ggplot(bootstrap_long, aes(x = bootstrap_values, fill = primary_treatment)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ metric, scales = "free") +
  scale_fill_manual(values=(pri_colors[c(2, 3)]))+
  labs(
    title = "Bootstrap Distributions of Variability Metrics",
    x = "Metric Value",
    y = "Density",
    fill = "Treatment"
  ) +
  theme_minimal()


ggplot(bootstrap_long %>% filter(!dpi.f %in% c(-8, 41)), aes(x = bootstrap_values, fill = primary_treatment)) +
  geom_histogram(binwidth = 0.1, position = "identity", alpha = 0.5) +
  scale_x_continuous(limits=c(0,1))+
  facet_grid(~ metric ~dpi.f, scales="free") +
  scale_fill_manual(values=(pri_colors[c(2, 3)]))+
  labs(
    title = "Bootstrap Distributions of Variability Metrics",
    x = "Metric Value",
    y = "Frequency",
    fill = "Treatment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(p.aba %>% filter(!dpi.f %in% c(-8, 41) & primary_treatment != "Sham"), aes(x = tes.new, fill = primary_treatment)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  #scale_x_continuous(limits=c(0,1))+
  facet_grid(~ primary_treatment ~dpi.f, scales="free") +
  scale_fill_manual(values=(pri_colors[c(2, 3)]))+
  labs(
    title = "TES",
    x = "Eye Score",
    y = "Frequency",
    fill = "Treatment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r Visualize Variability in Eye Score}
ggplot(e.cv)+
  #geom_jitter(aes(x=dpi.f, y=tes.new, color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV"), size=2)+
  #geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), size=2, position=position_dodge(width=0.5))+
  # geom_point(aes(x=dpi.f, y=bird_cv, color=primary_treatment, shape="CV"), size=2, position=position_dodge(width=0.75))+
  #  geom_errorbar(aes(x=dpi.f, ymin = cv_lower_ci, ymax = cv_upper_ci, y=bird_cv, color=primary_treatment), width=0.2, lty="dotted", position=position_dodge(width=0.75))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed")+
  #geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.5) , width=0.2)+
 scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 15, "PV" = 17, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  scale_color_manual(values=(pri_colors[c(2, 3)]))+
  labs(x="Days Post Inoculation", y="Variability", shape="Variability Metric", color="Primary Treatment")

ggplot(e.cv)+
  geom_jitter(aes(x=dpi.f, y=tes.new, color=primary_treatment), height=0, width=0.1, alpha=0.5)+
  scale_color_manual(values=(pri_colors[c(2, 3)]))+
  facet_wrap(~primary_treatment)
```
Strongly depends on what measure of variability I use - use PV for eye score becuase of zero inflation

```{r Total Eye Score Summed by priming treatment}
e.cv.t <- p.aba %>%
  filter(dpi < 42)%>%
  group_by(band_number)%>%
  reframe(tes_sum = sum(tes.new),
          primary_treatment = primary_treatment,
          dpi = dpi)

e.cv.t <- e.cv.t %>%
  filter(dpi == 41)
ggplot(e.cv.t, aes(x=primary_treatment, y=tes_sum, color=primary_treatment))+
  geom_boxplot()+
  geom_jitter(height=0, width=0.25, alpha=0.25)+
  scale_color_manual(values=pri_colors)+
  labs(x="Primary Treatment", y="Sum of Total Eye Score", color="Primary Treatment")

#unequal variance
leveneTest(tes_sum ~ primary_treatment, data=e.cv.t)

#Shapiro-Wilk test: data are not normally distributed
shapiro_res <- e.cv.t %>%
  filter(primary_treatment != "Sham")%>%
  group_by(primary_treatment)%>%
  summarise(shapiro_p = shapiro.test(tes_sum)$p.value)

#nonparametric Kruskal-Wallis
kruskal.test(tes_sum ~ primary_treatment, data= e.cv.t)

```

--------
I use raw quantity data because I want to capture heterogeneity as it applies to transmission-relevant traits. I don't want to reduce the impact of "outliers" as those data represent individuals that may play an outsized role in seeding/propagating epidemics.
```{r Calculate Raw Path Load Variability in infected birds primary infection}
#path load
q.cv <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri ==1)%>%
  drop_na(quantity1) %>%
  reframe(
    bird_cv = calculate_cv(quantity1),
    bird_pv = calculate_pv(quantity1),
    bird_v2 = calculate_v2(quantity1),
    dpi.f = dpi.f,
    quantity1 = quantity1,
    inf_pri = inf_pri,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(quantity1, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(quantity1, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(quantity1, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Remove bootstrap columns
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()
```

```{r Bootstrap distribution}
#path load
q.cv1 <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri ==1)%>%
  drop_na(quantity1) %>%
  reframe(
    bird_cv = calculate_cv(quantity1),
    bird_pv = calculate_pv(quantity1),
    bird_v2 = calculate_v2(quantity1),
    dpi.f = dpi.f,
    quantity1 = quantity1,
    inf_pri = inf_pri,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(quantity1, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(quantity1, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(quantity1, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Remove bootstrap columns
  #select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()

# Unnest bootstrap samples into a tidy format
bootstrap_long <- q.cv1 %>%
  select(dpi.f, primary_treatment, cv_bootstrap, pv_bootstrap, v2_bootstrap) %>%
  pivot_longer(
    cols = c(cv_bootstrap, pv_bootstrap, v2_bootstrap),
    names_to = "metric",
    values_to = "bootstrap_values"
  ) %>%
  unnest(cols = c(bootstrap_values))

# Add descriptive labels for metrics
bootstrap_long <- bootstrap_long %>%
  mutate(
    metric = case_when(
      metric == "cv_bootstrap" ~ "CV",
      metric == "pv_bootstrap" ~ "PV",
      metric == "v2_bootstrap" ~ "V2"
    )
  )
```


```{r Bootstrap distribution}
ggplot(bootstrap_long, aes(x = bootstrap_values, fill = primary_treatment)) +
  geom_density(alpha = 0.5) +
  facet_grid(~ metric ~dpi.f, scales="free") +
  scale_fill_manual(values=(pri_colors[c(2, 3)]))+
  labs(
    title = "Bootstrap Distributions of Variability Metrics",
    x = "Metric Value",
    y = "Density",
    fill = "Treatment"
  ) +
  theme_minimal()


ggplot(bootstrap_long %>% filter(dpi.f %in% c(7, 41)), aes(x = bootstrap_values, fill = primary_treatment)) +
  geom_histogram(binwidth = 0.1, position = "identity", alpha = 0.5) +
  scale_x_continuous(limits=c(0,1))+
  facet_grid(~ metric ~dpi.f, scales="free") +
  scale_fill_manual(values=(pri_colors[c(2, 3)]))+
  labs(
    title = "Bootstrap Distributions of Variability Metrics",
    x = "Metric Value",
    y = "Frequency",
    fill = "Treatment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(p.aba %>% filter(dpi.f %in% c(7, 41) & primary_treatment != "Sham"), aes(x = quantity1, fill = primary_treatment)) +
  geom_histogram(position = "identity", alpha = 0.5) +
  #scale_x_continuous(limits=c(0,1))+
  facet_grid(~ primary_treatment ~dpi.f, scales="free") +
  scale_fill_manual(values=(pri_colors[c(2, 3)]))+
  labs(
    title = "Quantity1",
    x = "Quantity1",
    y = "Frequency",
    fill = "Treatment"
  ) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r With Gamma Distributed variability metrics}
# Functions for gamma CV and V2 based on gamma distribution parameters
calculate_gamma_cv <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return(1 / sqrt(shape))
    }
  }
  return(NA)
}

calculate_gamma_V2 <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return((1 + shape)^(-1/2))
    }
  }
  return(NA)
}

# Set number of bootstrap replicates
n_boot <- 1000

# Main data processing pipeline
q.cv <- p.aba %>%
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri == 1 & dpi.f %in% c(-8, 7, 41)) %>%
  drop_na(quantity1) %>%
  dplyr::reframe(
    # Original data metrics
    bird_cv = calculate_cv(quantity1),
    bird_pv = calculate_pv(quantity1),
    bird_v2 = calculate_v2(quantity1),
    bird_gamma_cv = calculate_gamma_cv(quantity1),
    bird_gamma_V2 = calculate_gamma_V2(quantity1),
    
    dpi.f = dpi.f,
    quantity1 = quantity1,
    inf_pri = inf_pri,
    
    # Bootstrapping for CV, PV, V2, and gamma metrics
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(quantity1, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(quantity1, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(quantity1, replace = TRUE)))),
    gamma_cv_bootstrap = list(replicate(n_boot, calculate_gamma_cv(sample(quantity1, replace = TRUE)))),
    gamma_V2_bootstrap = list(replicate(n_boot, calculate_gamma_V2(sample(quantity1, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for all metrics
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    
    gamma_cv_lower_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_cv_upper_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    gamma_V2_lower_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_V2_upper_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Remove bootstrap columns to clean up the dataset
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap, -gamma_cv_bootstrap, -gamma_V2_bootstrap) %>%
  ungroup()

# View the resulting data
print(q.cv)

# Create summary table
summary_q_cv <- q.cv %>%
  group_by(dpi, primary_treatment) %>%
  summarize(
    # Mean values for each metric
    mean_bird_cv = mean(bird_cv, na.rm = TRUE),
    mean_bird_pv = mean(bird_pv, na.rm = TRUE),
    mean_bird_v2 = mean(bird_v2, na.rm = TRUE),
    mean_bird_gamma_cv = mean(bird_gamma_cv, na.rm = TRUE),
    mean_bird_gamma_V2 = mean(bird_gamma_V2, na.rm = TRUE),
    
    # 95% Confidence Intervals for each metric
    cv_lower_ci = mean(cv_lower_ci, na.rm = TRUE),
    cv_upper_ci = mean(cv_upper_ci, na.rm = TRUE),
    
    v2_lower_ci = mean(v2_lower_ci, na.rm = TRUE),
    v2_upper_ci = mean(v2_upper_ci, na.rm = TRUE),
    
    pv_lower_ci = mean(pv_lower_ci, na.rm = TRUE),
    pv_upper_ci = mean(pv_upper_ci, na.rm = TRUE),
    
    gamma_cv_lower_ci = mean(gamma_cv_lower_ci, na.rm = TRUE),
    gamma_cv_upper_ci = mean(gamma_cv_upper_ci, na.rm = TRUE),
    
    gamma_V2_lower_ci = mean(gamma_V2_lower_ci, na.rm = TRUE),
    gamma_V2_upper_ci = mean(gamma_V2_upper_ci, na.rm = TRUE)
  ) %>%
  ungroup()

# View the summary table
print(summary_q_cv)
```


Low dose birds are more variable in their pathogen loads during primary infection than high dose birds.
```{r Visualize in infected birds primary infection}
ggplot(q.cv)+
  #geom_jitter(aes(x=dpi.f, y=quantity1, color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV"), size=3, position=position_dodge(width=0.5))+
  #geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), size=3, position=position_dodge(width=0.5))+
  #geom_point(aes(x=dpi.f, y=bird_cv, color=primary_treatment, shape="CV"), size=2, position=position_dodge(width=0.75))+
  #geom_errorbar(aes(x=dpi.f, ymin = cv_lower_ci, ymax = cv_upper_ci, y=bird_cv, color=primary_treatment), width=0.2, lty="dotted", position=position_dodge(width=0.75))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed", position=position_dodge(width=0.5))+
  #geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.5), width=0.2)+
 scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 15, "PV" = 17, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  scale_color_manual(values=c(pri_colors[c(2, 3)]))+
  labs(title="Path Load Variability", y="Variability Metric", x="DPI", lty="Variability Metric", color="Primary Treatment")

ggplot(q.cv)+
    geom_boxplot(aes(x=dpi.f, y=quantity1, color=primary_treatment))+
  geom_jitter(aes(x=dpi.f, y=quantity1, color=primary_treatment), height=0, width=0.25, alpha=0.25)+
  geom_hline(yintercept = 50, alpha=0.25)+
  scale_color_manual(values=c(pri_colors[c(2, 3)]))
```


```{r Log10 Quantity Variability}
p.aba$quantity1 <- p.aba$quantity1+0.0001
q.cv.log <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  drop_na(quantity1) %>%
  reframe(
    bird_cv = calculate_cv(log10(quantity1)),
    bird_pv = calculate_pv(log10(quantity1)),
    bird_v2 = calculate_v2(log10(quantity1)),
    dpi.f = dpi.f,
    quantity1 = quantity1,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(log10(quantity1), replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(log10(quantity1), replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(log10(quantity1), replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Remove bootstrap columns
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()
```

Variability here is funky because there are values below 50 (infection cutoff) that are affecting variability metrics.
```{r Log Quantity Variability}
hist(log10(q.cv.log$quantity1))
ggplot(q.cv.log %>% filter(primary_treatment != "Sham"))+
  #geom_jitter(aes(x=dpi.f, y=log10(quantity1), color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV"), size=3)+
  geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), size=3, position=position_dodge(width=0.2))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed")+
  geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.2) , width=0.2)+
  scale_color_manual(values=c(pri_colors[c(2,3)]))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("PV" = 17, "V2" = 18),  # Assign shapes
                     labels = c("PV" = "PV", "V2" = "CV^2")) +
  labs(title="Log10(Quantity) Variability", y="Variability", x="DPI")

ggplot(q.cv.log)+
  geom_jitter(aes(x=dpi.f, y=log10(quantity1), color=primary_treatment), height=0, width=0.25, alpha=0.5)+
  scale_color_manual(values = c(pri_colors))+
  geom_hline(yintercept = log10(50.001))

ggplot(q.cv.log %>% filter(primary_treatment != "Sham"))+
  #geom_jitter(aes(x=dpi.f, y=log10(quantity1), color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV Log"), position = position_dodge(width=0.3))+
  #geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), position=position_dodge(width=0.2))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), position = position_dodge(width=0.3), width=0.1, lty="solid")+
  # geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.2) , width=0.2)+
   geom_point(data=q.cv, aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV Raw"), size=3, position=position_dodge(width=0.5))+
  geom_errorbar(data=q.cv, aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed", position=position_dodge(width=0.5))+
  scale_color_manual(values=c(pri_colors[c(2,3)]))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("PV Log" = 17, "PV Raw" = 23),  # Assign shapes
                     labels = c("PV Log" = "PV Log", "PV Raw" = "PV Raw")) +
  labs(title="Log10(Quantity) vs Raw Quantity Variability", y="Variability", x="DPI")
```
Use PV for quantity variability

Low dose birds are more variable in their pathogen loads during primary infection than high dose birds, but less variable
#####Secondary Infection

```{r Remove Birds That Did Not Resolve Infection By DPI 42}
#omit from dataset - were still infected (path load) on day 41 prior to reinfection day 42
m.ab <- m.ab %>%
  filter(!(band_number %in% c(2274, 2469, 2520, 2494)))

m.ab %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  )%>%
  modify_header(
    label ~ "**FINAL SAMPLE SIZES SECONDARY**"
  )

m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose, inf_sec) %>%
  group_by(secondary_dose, primary_treatment) %>%
  tbl_summary()%>%
  modify_header(
    label ~ "**By Primary and Secondary Dose**"
  )
m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=primary_treatment
  )
```
#####Do Antibody Levels During Primary Infection Predict Susceptibility to Reinfection?
```{r Set Theme for Susceptibility Graphs}
sus_theme_pre <- theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    axis.title.x = element_blank(),  # Remove the x-axis title
    axis.title.y = element_text(size = 15, face="bold"),
    plot.title = element_text(hjust = 0.5),  # Center the plot title
    legend.position = "top",
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    strip.text = element_text(size = 10)  # Adjust the size of facet labels
  )

sus_theme_14 <- theme(
  axis.text.x = element_text(angle = 45, hjust = 1), 
  axis.title.y = element_text(size = 15, face="bold"),
  legend.position = "none",  # Remove the legend
  plot.title = element_blank(),  # Remove the plot title
  axis.title.x = element_blank(),  # Remove the x-axis title
  legend.text = element_blank(),  # Remove the legend text
  legend.title = element_blank(),  # Remove the legend title
  legend.background = element_blank(),  # Remove the legend background
  strip.text = element_text(size = 10)  # Adjust the size of facet labels
)

sus_theme_41 <- theme(
  axis.text.x = element_text(angle = 45, hjust = 1), 
  axis.title.y = element_text(size = 15, face="bold"),
  legend.position = "none",  # Remove the legend
  plot.title = element_blank(),  # Remove the plot title
  axis.title.x = element_text(size=15, face="bold"), 
  legend.text = element_blank(),  # Remove the legend text
  legend.title = element_blank(),  # Remove the legend title
  legend.background = element_blank(),  # Remove the legend background
  strip.text = element_text(size = 10)  # Adjust the size of facet labels
)
```

```{r Format Pre Antibodies}
m.ab.w <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, elisa_od, inf_sec)%>%
  filter(dpi %in% c(-8, 14, 41))

wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)

wibird$elisa_od_pre <- wibird$`elisa_od_-8`

wibird <- wibird %>% mutate(totalTES = sum())
wibird$sec_dose.n <- wibird$secondary_dose+1
wibird$log10.sec_dose <- round(log10(wibird$sec_dose.n), digits = 2)

wibird$log10.sec_dose <- as.numeric(wibird$log10.sec_dose)

wibird$secondary_dose_fct <- factor(wibird$secondary_dose, levels = c(0, 30, 100, 300, 7000))

sec_dose_names <- c(
  "0" = "0",
  "1.49" = "30",
  "2" = "100",
  "2.48" = "300",
  "3.85" = "7000"
)
```

```{r Is susceptibility to reinfection predicted by antibody levels 8 days before infection?}
glm.ab.pre <- glm(inf_sec ~ elisa_od_pre + log10.sec_dose, data=wibird, family=binomial())
summary(glm.ab.pre)
car::Anova(glm.ab.pre, type = 3)
simulateResiduals(glm.ab.pre, plot=T)

mod <- glm.ab.pre
dat.new=expand.grid(log10.sec_dose=unique(wibird$log10.sec_dose),
                    inf_sec=unique(wibird$inf_sec),
                    elisa_od_pre = unique(wibird$elisa_od_pre))#new grid to put predictions into
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

#names
wibird$dpi <- "DPPI -8"
dpi_namespre <- c(
  "DPPI -8" = "DPPI -8"
)

sus_pre_raw<-ggplot(wibird, aes(x=(elisa_od_pre), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_pre), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_pre, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(y=" ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi ~log10.sec_dose, labeller = as_labeller(c(dpi_namespre, sec_dose_names)))
sus_pre_raw

sus_pre_raw.all<-ggplot(wibird, aes(x=(elisa_od_pre), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_pre), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_pre, y=yhat,
            fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(y=" ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))
  facet_grid(~dpi ~log10.sec_dose, labeller = as_labeller(c(dpi_namespre, sec_dose_names)))
sus_pre_raw.all
```

```{r Format DPI 14 Antibodies}
m.ab.w <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, elisa_od, inf_sec)%>%
  filter(dpi %in% c(-8, 14, 41))

wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)
wibird14 <- wibird %>%
  drop_na(elisa_od_14)

wibird14 <- wibird14 %>% mutate(totalTES = sum())
wibird14$sec_dose.n <- wibird14$secondary_dose+1
wibird14$log10.sec_dose <- round(log10(wibird14$sec_dose.n), digits = 2)

wibird14$log10.sec_dose <- as.numeric(wibird14$log10.sec_dose)
```

```{r Are Antibody Levels on Day 14 Predictive of Susceptibility?}
glm.ab14 <- glm(inf_sec ~ elisa_od_14 + log10.sec_dose, data=wibird14, family=binomial())
summary(glm.ab14)
###Summary(glm) gives the coefficient estimates for the model to help interpret the relationship between predictors and outcome.
##Estimate represents the log odds change in inf_sec for a one-unit change in the predictor
##z value: test statistic for the Wald test (Estimate / SE)
##Pr(>|z|): p-value for whether the coefficient is significantly different from zero
##Null deviance: How well the model predicts without predictors
##Residual deviace: How well the model predicts with the predictors

# Call:
# glm(formula = inf_sec ~ elisa_od_14 + log10.sec_dose, family = binomial(), 
#     data = wibird14)
# 
# Coefficients:
#                Estimate Std. Error z value Pr(>|z|)    
# (Intercept)     -3.6985     1.2568  -2.943  0.00325 ** 
# elisa_od_14    -48.7587    17.3410  -2.812  0.00493 ** 
# log10.sec_dose   1.8746     0.3731   5.024 5.06e-07 ***
# ---
# Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1
# 
# (Dispersion parameter for binomial family taken to be 1)
# 
#     Null deviance: 121.729  on 130  degrees of freedom
# Residual deviance:  70.284  on 128  degrees of freedom
# AIC: 76.284
# 
# Number of Fisher Scoring iterations: 7

# Get the summary of the model
summary_output <- summary(glm.ab14)
# Extract coefficients table
coef_table <- summary_output$coefficients

# Convert to a data frame for easier export
coef_df <- as.data.frame(coef_table)

# Create a summary data frame with additional model information
model_info <- data.frame(
  AIC = AIC(glm.ab14),
  Deviance = summary_output$deviance,
  Null_Deviance = summary_output$null.deviance,
  Residual_Df = summary_output$df.residual,
  Null_Df = summary_output$df.null
)

final_summary <- cbind(coef_df, model_info)

#write.csv(final_summary, "/Users/jesse/Documents/GitHub/EEID_1A_Mechanistic_Link/EEID_1A_Antibody_Analysis_files/Results/glm_ab14_full_summary.csv", row.names = TRUE)

#ANOVA type III gives analysis of deviance table, which evaluates the contribution of each predictor variable after accounting for the other predictors. Type III tests the effect of a variable while adjusting for all other variables, letting you look at each variable's unique contribution.
car::Anova(glm.ab14, type = 3)

# Analysis of Deviance Table (Type III tests)
### Does removing a predictor significantly worsen the model fit?
##LR Chisq = Likelihood Ratio Chi-Squared Test: Measures how much a predictor contributes to reducing the model's deviance (making fit better)
##Df: number of values free to vary for that predictor 
##Pr(>Chisq): p-value for the predictor

# Response: inf_sec
#                LR Chisq Df Pr(>Chisq)    
# elisa_od_14      13.436  1  0.0002468 *** 
# log10.sec_dose   41.135  1  1.421e-10 ***

simulateResiduals(glm.ab14, plot=T)

mod <- glm.ab14
dat.new=expand.grid(log10.sec_dose=unique(wibird14$log10.sec_dose),
                    inf_sec=unique(wibird14$inf_sec),
                    elisa_od_14 = unique(wibird14$elisa_od_14))
# elisa_od_14 = unique(wibird14$elisa_od_14))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird14$dpi <- "DPPI 14"
dpi_names14 <- c(
  "DPPI 14" = "DPPI 14"
)

sus_14_raw<-ggplot(wibird14, aes(x=(elisa_od_14), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_line(data=dat.new, aes(x=(elisa_od_14), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_14, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  geom_point(alpha=0.5, size=2)+
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(y= "Infection Status Upon Secondary Challenge (0|1)", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi~log10.sec_dose, labeller = as_labeller(c(dpi_names14,sec_dose_names)))
sus_14_raw

sus_14_raw.all<-ggplot(wibird14, aes(x=(elisa_od_14), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_line(data=dat.new, aes(x=(elisa_od_14), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_14, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  geom_point(alpha=0.5, size=2)+
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(y= "Infection Status Upon Secondary Challenge (0|1)", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))#+
  facet_grid(~dpi~log10.sec_dose, labeller = as_labeller(c(dpi_names14,sec_dose_names)))
sus_14_raw.all
```

```{r Format DPI 41 Antibodies}
wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)

wibird41 <- wibird %>%
  drop_na(elisa_od_41)

wibird41 <- wibird41 %>% mutate(totalTES = sum())
wibird41$sec_dose.n <- wibird41$secondary_dose+1
wibird41$log10.sec_dose <- round(log10(wibird41$sec_dose.n), digits = 2)

wibird41$log10.sec_dose <- as.numeric(wibird41$log10.sec_dose)

```

```{r Are Antibody Levels on Day 41 Predictive of Susceptibility to Reinfection?}
glm.ab41 <- glm(inf_sec ~ elisa_od_41 + log10.sec_dose, data=wibird41, family=binomial())

summary(glm.ab41)
# Get the summary of the model
summary_output <- summary(glm.ab41)
# Extract coefficients table
coef_table <- summary_output$coefficients

# Convert to a data frame for easier export
coef_df <- as.data.frame(coef_table)

# Create a summary data frame with additional model information
model_info <- data.frame(
  AIC = AIC(glm.ab41),
  Deviance = summary_output$deviance,
  Null_Deviance = summary_output$null.deviance,
  Residual_Df = summary_output$df.residual,
  Null_Df = summary_output$df.null
)

final_summary <- cbind(coef_df, model_info)

#write.csv(final_summary, "/Users/jesse/Documents/GitHub/EEID_1A_Mechanistic_Link/EEID_1A_Antibody_Analysis_files/Results/glm_ab41_full_summary.csv", row.names = TRUE)

#ANOVA type III gives analysis of deviance table, which evaluates the contribution of each predictor variable after accounting for the other predictors. Type III tests the effect of a variable while adjusting for all other variables, letting you look at each variable's unique contribution.
car::Anova(glm.ab41, type = 3)

# Analysis of Deviance Table (Type III tests)
### Does removing a predictor significantly worsen the model fit?
##LR Chisq = Likelihood Ratio Chi-Squared Test: Measures how much a predictor contributes to reducing the model's deviance (making fit better)
##Df: number of values free to vary for that predictor 
##Pr(>Chisq): p-value for the predictor

    #Response: inf_sec
    #               LR Chisq Df Pr(>Chisq)  
    #elisa_od_41      17.102  1  3.543e-05 ***
    #log10.sec_dose   49.068  1  2.473e-12 ***

car::Anova(glm.ab41, type = 3)
simulateResiduals(glm.ab41, plot=T)

mod <- glm.ab41
dat.new=expand.grid(log10.sec_dose=unique(wibird41$log10.sec_dose),
                    inf_sec=unique(wibird41$inf_sec),
                    primary_treatment=unique(wibird41$primary_treatment),
                    #elisa_od_14_new=unique(wibird41$elisa_od_14_new),
                    elisa_od_41 = unique(wibird41$elisa_od_41))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird41$dpi <- "DPPI 41"
dpi_names41 <- c(
  "DPPI 41" = "DPPI 41"
)

sus_41_raw<-ggplot(wibird41, aes(x=(elisa_od_41), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_41), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_41, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(x="IgY Antibody Levels", y= " ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi~log10.sec_dose, labeller = as_labeller(c(dpi_names41,sec_dose_names)))
sus_41_raw

sus_41_raw.all<-ggplot(wibird41, aes(x=(elisa_od_41), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_41), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_41, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(x="IgY Antibody Levels", y= " ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_wrap(~primary_treatment, nrow=3)
sus_41_raw.all
```
```{r Compare Models}
aictab(cand.set=list(glm.ab14, glm.ab41), modnames=c("glm.ab14", "glm.ab41"))
#day 14 was a better fit

#anova(glm.ab14, glm.ab41, test="Chisq")
```


```{r}
library("grid")
library(cowplot)
r1 <- sus_pre_raw + sus_theme_pre
r2 <- sus_14_raw + sus_theme_14
r3 <- sus_41_raw + sus_theme_41
legend <-get_legend(r1)
r1 <- r1 + theme(legend.position = "top")

Raw_all <- arrangeGrob(
  legend,
  arrangeGrob(r1, r2, r3, ncol = 1),
  ncol = 1,
  heights = c(1, 10)
)

grid.newpage()
grid.draw(Raw_all)

```

```{r}
m.ab.w <- m.ab.w %>% mutate(totalTES = sum())
m.ab.w$sec_dose.n <- m.ab.w$secondary_dose+1
m.ab.w$log10.sec_dose <- round(log10(m.ab.w$sec_dose.n), digits = 2)

m.ab.w$log10.sec_dose <- as.numeric(m.ab.w$log10.sec_dose)
m.ab.w$dpi.f <- as.factor(m.ab.w$dpi)

m.ab.w <- m.ab.w %>%
  drop_na(elisa_od)

glm.ab <- glmmTMB(inf_sec ~ elisa_od + log10.sec_dose + dpi.f + (1|band_number), data=m.ab.w, family=binomial())

summary(glm.ab)

car::Anova(glm.ab, type = 3)
simulateResiduals(glm.ab, plot=T)

mod <- glm.ab
dat.new=expand.grid(log10.sec_dose=unique(m.ab.w$log10.sec_dose),
                    inf_sec=unique(m.ab.w$inf_sec),
                    dpi.f=unique(m.ab.w$dpi.f),
                    band_number = unique(m.ab.w$band_number),
                    elisa_od = unique(m.ab.w$elisa_od))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

ggplot(m.ab.w, aes(x=(elisa_od), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(x="IgY Antibody Levels", y= " ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi)
```

Antibodies and susceptibility by priming dose
```{r Antibodies and susceptibility by priming dose dpi 41} 
glm.ab41.pri <- glm(inf_sec ~ elisa_od_41 + log10.sec_dose * primary_treatment, data=wibird41, family=binomial())

summary(glm.ab41.pri)
car::Anova(glm.ab41.pri, type = 3)
simulateResiduals(glm.ab41.pri, plot=T)

mod <- glm.ab41.pri
dat.new=expand.grid(log10.sec_dose=unique(wibird41$log10.sec_dose),
                    inf_sec=unique(wibird41$inf_sec),
                    primary_treatment=unique(wibird41$primary_treatment),
                    elisa_od_41 = unique(wibird41$elisa_od_41))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird41$dpi <- "DPPI 41"
dpi_names41 <- c(
  "DPPI 41" = "DPPI 41"
)

ggplot(wibird41, aes(x=(elisa_od_41), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_41), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_41, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(x="IgY Antibody Levels", y= " ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~log10.sec_dose~primary_treatment)

```

```{r}
glm.ab41.pri.o <- glm(inf_sec ~ elisa_od_41 + primary_treatment, data=wibird41, family=binomial())

summary(glm.ab41.pri.o)
car::Anova(glm.ab41.pri.o, type = 3)
simulateResiduals(glm.ab41.pri.o, plot=T)

mod <- glm.ab41.pri.o
dat.new=expand.grid(#log10.sec_dose=unique(wibird41$log10.sec_dose),
                    inf_sec=unique(wibird41$inf_sec),
                    primary_treatment=unique(wibird41$primary_treatment),
                    elisa_od_41 = unique(wibird41$elisa_od_41))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird41$dpi <- "DPPI 41"
dpi_names41 <- c(
  "DPPI 41" = "DPPI 41"
)

ggplot(wibird41, aes(x=(elisa_od_41), y=(inf_sec), color=as.factor(primary_treatment)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_41), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_41, y=yhat, fill=as.factor(primary_treatment)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(pri_colors), labels=c("Sham", "Low", "High"))+
  scale_fill_manual(values = c(pri_colors), labels=c("Sham", "Low", "High"))+
  labs(x="IgY Antibody Levels", y= " ", color="Primary Dose", fill ="Primary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~primary_treatment)

```

```{r Antibodies and susceptibility by priming dose dpi 14} 
glm.ab14.pri <- glm(inf_sec ~ elisa_od_14 + log10.sec_dose * primary_treatment, data=wibird14, family=binomial())

summary(glm.ab14.pri)
car::Anova(glm.ab14.pri, type = 3)
simulateResiduals(glm.ab14.pri, plot=T)

mod <- glm.ab14.pri
dat.new=expand.grid(log10.sec_dose=unique(wibird41$log10.sec_dose),
                    inf_sec=unique(wibird41$inf_sec),
                    primary_treatment=unique(wibird41$primary_treatment),
                    elisa_od_14 = unique(wibird41$elisa_od_14))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird41$dpi <- "DPPI 41"
dpi_names41 <- c(
  "DPPI 41" = "DPPI 41"
)

ggplot(wibird14, aes(x=(elisa_od_14), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_14), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_14, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(x="IgY Antibody Levels", y= " ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  geom_vline(xintercept = 0.061, alpha=0.25, lty="dashed")+
  facet_grid(~log10.sec_dose~primary_treatment)

```

```{r}
glm.ab14.pri.o <- glm(inf_sec ~ elisa_od_14 + primary_treatment, data=wibird14, family=binomial())

summary(glm.ab14.pri.o)
car::Anova(glm.ab14.pri.o, type = 3)
simulateResiduals(glm.ab14.pri.o, plot=T)

mod <- glm.ab14.pri.o
dat.new=expand.grid(#log10.sec_dose=unique(wibird14$log10.sec_dose),
                    inf_sec=unique(wibird14$inf_sec),
                    primary_treatment=unique(wibird14$primary_treatment),
                    elisa_od_14 = unique(wibird14$elisa_od_14))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird14$dpi <- "DPPI 14"
dpi_names14 <- c(
  "DPPI 14" = "DPPI 14"
)

ggplot(wibird14, aes(x=(elisa_od_14), y=(inf_sec), color=as.factor(primary_treatment)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_14), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_14, y=yhat, fill=as.factor(primary_treatment)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(pri_colors), labels=c("Sham", "Low", "High"))+
  scale_fill_manual(values = c(pri_colors), labels=c("Sham", "Low", "High"))+
  labs(x="IgY Antibody Levels", y= " ", color="Primary Dose", fill ="Primary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~primary_treatment)

```

####Do antibody levels in susceptible birds predict pathogen load or eye score?
```{r Format wide data}
m.ab.w <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, elisa_od, inf_sec, inf_pri)%>%
  filter(dpi %in% c(-8, 14, 41, 56))

wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)

m.ab.w.tes <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, tes, inf_sec, inf_pri)%>%
  filter(dpi %in% c(-8, 14, 41, 56))
wibird.tes <- m.ab.w.tes %>% pivot_wider(names_from = dpi,
                                     names_glue = "{.value}_{dpi}",
                                     values_from = tes)

m.ab.w.quant <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, quantity1, inf_sec, inf_pri)%>%
  filter(dpi %in% c(-8, 14, 41, 56))
wibird.quant <- m.ab.w.quant %>% pivot_wider(names_from = dpi,
                                         names_glue = "{.value}_{dpi}",
                                         values_from = quantity1)

wibird.all <- left_join(wibird, wibird.quant, by="band_number")
wibird.all <- left_join(wibird.all, wibird.tes, by="band_number")
wibird.all <- wibird.all %>%
  select(-primary_treatment.y, -secondary_dose.y, -sex.y, -inf_sec.y, -inf_pri.y, -primary_treatment.x, -secondary_dose.x, -sex.x, -inf_sec.x, -inf_pri.x)
head(wibird.all)
```
Antibody levels do not predict path load or eye score dpi 56
```{r Do antibody levels on dpi 41 predict path load or eye score dpi 56?}
wibird.all$quantity1_56.log <- log10(wibird.all$quantity1_56)
wibird.all.41 <- wibird.all %>%
  drop_na(elisa_od_41)

glm.ab.quant <- glmmTMB(quantity1_56.log+.01 ~ elisa_od_41 + primary_treatment, data=wibird.all.41, family=Gamma())
summary(glm.ab.quant)

glm.ab.tes <- glmmTMB(tes_56+0.001 ~ elisa_od_41 + primary_treatment, data=wibird.all.41, family = Gamma())
summary(glm.ab.tes)
hist(wibird.all.41$tes_56)
ggplot(wibird.all, aes(y=elisa_od_41, x=quantity1_56.log, color=primary_treatment))+
  geom_point()+
  scale_color_manual(values=c(pri_colors))+
  geom_smooth(aes(groups=primary_treatment), method="glm")+
  facet_wrap(~secondary_dose)

```
Antibody levels do not predict path load or eye score dpi 56
```{r Do antibody levels on dpi 14 predict path load or eye score dpi 56?}
wibird.all.14 <- wibird.all %>%
  drop_na(elisa_od_14)

glm.ab.quant <- glmmTMB(quantity1_56.log+.01 ~ elisa_od_14 + primary_treatment, data=wibird.all.14, family=Gamma())
summary(glm.ab.quant)

glm.ab.tes <- glmmTMB(tes_56+0.001 ~ elisa_od_14 + primary_treatment, data=wibird.all.14, family = Gamma())
summary(glm.ab.tes)
```

####Variability combined with Hawley et al., 2024 Susceptibility####
```{r}
####Variability With Hawley et al., 2024####
summary_df <- a.cv %>%
  filter(dpi.f == 41) %>%
  group_by(primary_treatment) %>%
  summarize(
    mean_bird_pv = mean(bird_pv, na.rm = TRUE),
    mean_bird_cv = mean(bird_cv, na.rm = TRUE),
    mean_mean_od = mean(mean_od, na.rm = TRUE)
  )

summary_df
dose <- c("Sham", "Low", "High")
CV <- as.numeric(c(0.899, 1.630, 2.511))
ab.pv <- as.numeric(c(0.04156784, 0.14716858, 0.18885516))
ab.cv <- as.numeric(c(0.03997555, 0.17202405, 0.26176249))
ci_lower_cv <- as.numeric(c(0.0230, 0.119, 0.165))
ci_upper_cv <- as.numeric(c(0.0409, 0.206, 0.339))
ci_lower_pv <- as.numeric(c(0.0246, 0.0998, 0.144))
ci_upper_pv <- as.numeric(c(0.0452, 0.178, 0.231))
mean_sus <- as.numeric(c(1.181, 0.446, 0.192))
mean_ab <- as.numeric(c(0.045, 0.04786, 0.0579))
dpi <- 41



het.df <- data.frame(dose, CV, mean_sus, ab.pv, ab.cv, dpi, mean_ab, ci_lower_cv, ci_upper_cv, ci_lower_pv, ci_upper_pv)

het.df$mean_sus_inv <- 1/het.df$mean_sus

#hawley et al 2024
het.paper.only <- ggplot(het.df, aes(x=Metric))+
  geom_point(size=5,aes(x="CV", y=CV, color=dose), shape = 17)+
  geom_point(size=5,aes(x="CV", y=CV, shape = "CV"), color="black", shape=2, stroke = 1)+
  scale_color_manual(values=c(pri_colors))+
  scale_y_continuous(
    name= "Variability in Susceptibility Hawley et al. 2024"
  )+
  theme_minimal()+
  theme(
    axis.title.y = element_text(color = "black", size=12, face="bold"),
    legend.position = "none"
  ) +
  labs(color = "Primary Treatment", shape= "Variability Metric")
het.paper.only

ab.paper.only <- ggplot(het.df, x=Metric)+
  geom_point(size=5,aes(x="PV", y=ab.pv, color=dose, shape="PV"))+
  geom_errorbar(aes(ymin=ci_lower_pv, ymax=ci_upper_pv, y=ab.pv, x="PV", color=dose), width=0.1)+
  geom_point(size=5,aes(x="CV", y=ab.cv, color=dose, shape = "CV"))+
  geom_errorbar(aes(ymin=ci_lower_cv, ymax=ci_upper_cv, y=ab.cv, x="CV", color=dose), width=0.1)+
  #geom_point(size=3.6,aes(x="PV", y=ab.pv), color="black", shape=1, stroke = 2)+
  scale_color_manual(values=c(pri_colors))+
  scale_y_continuous(
    name="Variability in Antibody Levels DPPI 41"
  )+
  scale_x_discrete(
    name="Metric"
  )+
  scale_shape_manual(values=c(17,19))+
  theme_minimal()+
  theme(
    axis.title.y = element_text(color = "black", size=12, face = "bold"),
    legend.position="top",
    legend.direction = "horizontal",
    legend.box = "vertical",
    legend.background= element_rect(size=0.25, linetype="solid"))+
  labs(color = "Primary Treatment", shape= "Variability Metric")
ab.paper.only

legend <- cowplot::get_legend(ab.paper.only)

# Remove legends from plots
het.paper.only <- het.paper.only + theme(legend.position = "none")
ab.paper.only <- ab.paper.only + theme(legend.position = "none")

#FIGURE 4
Fig4 <- grid.arrange(legend, 
             arrangeGrob(het.paper.only, ab.paper.only, ncol = 2),
             ncol = 1,
             heights = c(1, 4))
```

####Variability in Reinfected Birds####
```{r Secondary Sample Sizes}
sec.ab <- m.ab %>%
  filter(inf_sec == 1 & dpi > 41 & secondary_dose == 7000)

m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose, inf_sec) %>%
  group_by(secondary_dose, primary_treatment) %>%
  tbl_summary()%>%
  modify_header(
    label ~ "**By Primary and Secondary Dose**"
  )

m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(inf_sec, primary_treatment, secondary_dose, band_number) %>%
  group_by(inf_sec, secondary_dose, primary_treatment) %>%
  tbl_summary(
    by = inf_sec, # Grouping by infected secondary
    statistic = list(band_number ~ "{n} ({p}%)") # Count occurrences and show percentages
  ) %>%
  modify_header(label ~ "**Infected Secondary**") 

sec.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose, inf_sec) %>%
  group_by(secondary_dose, primary_treatment) %>%
  tbl_summary()%>%
  modify_header(
    label ~ "**By Primary and Secondary Dose**"
  )
```

```{r Max quantity and eye score secondary}
max.quant.sec <- sec.ab %>%
  filter(dpi > 41 & inf_sec ==1) %>%
  group_by(band_number, primary_treatment, secondary_dose) %>%
  reframe(max_quantity = max(quantity1, na.rm = TRUE))

max.tes.sec <- sec.ab %>%
  group_by(band_number, primary_treatment, secondary_dose)%>%
  reframe(max_tes = max(tes, na.rm=TRUE))

max.sec<-left_join(max.quant.sec, max.tes.sec, by="band_number")

max.sec$primary_treatment <- max.sec$primary_treatment.x
max.sec$secondary_dose <- max.sec$secondary_dose.x

max.sec <- max.sec %>%
  dplyr::select(-c(primary_treatment.y, secondary_dose.y, primary_treatment.x, secondary_dose.x))
max.sec$max_quantity1 <- max.sec$max_quantity+1

```
Pathogen load by priming dose faceted by secondary challenge
```{r}
#only 7000 reinfected
ggplot(sec.ab, aes(x=primary_treatment, y=log10_quantity1, color=primary_treatment))+
  geom_boxplot()+
  geom_jitter(alpha=0.5, width=0.25, height=0)+
  scale_color_manual(values=pri_colors)+
  geom_hline(yintercept = log10(51), alpha=0.5, lty="dashed")+
  facet_wrap(~dpi, nrow=1)+
  theme_bw()

sec.all <- m.ab %>%
  filter(dpi > 41)

#all secondary
ggplot(sec.all, aes(x=dpi, y=log10_quantity1, color=primary_treatment))+
  geom_jitter(alpha=0.5, height=0, width=1)+
  stat_summary(aes(x=dpi, y=log10_quantity1, groups=primary_treatment), geom="point", fun.y = mean)+
  stat_summary(aes(x=dpi, y=log10_quantity1, groups=primary_treatment), geom="errorbar", fun.data=mean_se, width=0.5)+
  stat_summary(aes(x=dpi, y=log10_quantity1, groups=primary_treatment), geom="line", fun.y=mean)+
  scale_color_manual(values=pri_colors)+
  facet_wrap(~secondary_dose, nrow=1)+
  geom_hline(yintercept = log10(51), alpha=0.5, lty="dashed")+
  theme_bw()

#infected secondary
ggplot(sec.all%>% filter(inf_sec == 1), aes(x=dpi, y=log10_quantity1, color=primary_treatment))+
  geom_jitter(alpha=0.5, height=0, width=0)+
  stat_summary(aes(x=dpi, y=log10_quantity1, groups=primary_treatment), geom="point", fun.y = mean)+
  stat_summary(aes(x=dpi, y=log10_quantity1, groups=primary_treatment), geom="errorbar", fun.data=mean_se, width=0.5)+
  stat_summary(aes(x=dpi, y=log10_quantity1, groups=primary_treatment), geom="line", fun.y=mean)+
    geom_hline(yintercept = log10(51), alpha=0.5, lty="dashed")+
  scale_color_manual(values=pri_colors)+
  facet_wrap(~secondary_dose, nrow=1)+
  theme_bw()
```

eye score by priming dose faceted by secondary challenge
```{r}
#only 7000 reinfected
ggplot(sec.ab, aes(x=primary_treatment, y=tes, color=primary_treatment))+
  geom_boxplot()+
  geom_jitter(width=0.25, height=0, alpha=0.5)+
  scale_color_manual(values=pri_colors)+
  facet_wrap(~dpi, nrow=1)+
  theme_bw()

sec.all <- m.ab %>%
  filter(dpi > 41)

#all secondary
ggplot(sec.all, aes(x=dpi, y=tes, color=primary_treatment))+
  geom_jitter(alpha=0.5, height=0, width=1)+
  stat_summary(aes(x=dpi, y=tes, groups=primary_treatment), geom="point", fun.y = mean)+
  stat_summary(aes(x=dpi, y=tes, groups=primary_treatment), geom="errorbar", fun.data=mean_se, width=0.5)+
  stat_summary(aes(x=dpi, y=tes, groups=primary_treatment), geom="line", fun.y=mean)+
  scale_color_manual(values=pri_colors)+
  facet_wrap(~secondary_dose, nrow=1)+
  theme_bw()

#infected secondary
ggplot(sec.all%>% filter(inf_sec == 1), aes(x=dpi, y=tes, color=primary_treatment))+
  geom_jitter(alpha=0.5, height=0, width=0)+
  stat_summary(aes(x=dpi, y=tes, groups=primary_treatment), geom="point", fun.y = mean)+
  stat_summary(aes(x=dpi, y=tes, groups=primary_treatment), geom="errorbar", fun.data=mean_se, width=0.5)+
  stat_summary(aes(x=dpi, y=tes, groups=primary_treatment), geom="line", fun.y=mean)+
  scale_color_manual(values=pri_colors)+
  facet_wrap(~secondary_dose, nrow=1)+
  theme_bw()
```


```{r}
# Pipeline to calculate CV, PV, V2, and their 95% confidence intervals for max_quantity and max_tes

#add small constant for variability calculations
#sec.var$max_tes <- sec.var$max_tes + 0.001

sec.var <- max.sec %>%
  group_by(primary_treatment) %>%
  drop_na(max_quantity, max_tes) %>%
  dplyr::reframe(
    # For max_quantity
    max_quantity = max_quantity,
    mean_quantity = mean(max_quantity),
    median_quantity = median(max_quantity),
    bird_sd_quantity = sd(max_quantity),
    band_number = band_number,
    bird_cv_quantity = calculate_cv(max_quantity),  # Calculate CV for max_quantity
    bird_pv_quantity = calculate_pv(max_quantity),  # Calculate PV for max_quantity
    bird_v2_quantity = calculate_v2(max_quantity),  # Calculate V2 for max_quantity
    
  
    # Bootstrap for each metric for max_quantity
    cv_bootstrap_quantity = list(replicate(n_boot, calculate_cv(sample(max_quantity, replace = TRUE)))),
    pv_bootstrap_quantity = list(replicate(n_boot, calculate_pv(sample(max_quantity, replace = TRUE)))),
    v2_bootstrap_quantity = list(replicate(n_boot, calculate_v2(sample(max_quantity, replace = TRUE)))),

    # For max_tes
    max_tes = max_tes,
    mean_tes = mean(max_tes),
    median_tes = median(max_tes),
    bird_sd_tes = sd(max_tes),
    bird_cv_tes = calculate_cv(max_tes),  # Calculate CV for max_tes
    bird_pv_tes = calculate_pv(max_tes),  # Calculate PV for max_tes
    bird_v2_tes = calculate_v2(max_tes),  # Calculate V2 for max_tes
    
    # Bootstrap for each metric for max_tes
    cv_bootstrap_tes = list(replicate(n_boot, calculate_cv(sample(max_tes, replace = TRUE)))),
    pv_bootstrap_tes = list(replicate(n_boot, calculate_pv(sample(max_tes, replace = TRUE)))),
    v2_bootstrap_tes = list(replicate(n_boot, calculate_v2(sample(max_tes, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for CV, PV, and V2 for both max_quantity and max_tes
  mutate(
    # For max_quantity
    cv_lower_ci_quantity = map_dbl(cv_bootstrap_quantity, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci_quantity = map_dbl(cv_bootstrap_quantity, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci_quantity = map_dbl(pv_bootstrap_quantity, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci_quantity = map_dbl(pv_bootstrap_quantity, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci_quantity = map_dbl(v2_bootstrap_quantity, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci_quantity = map_dbl(v2_bootstrap_quantity, ~ quantile(.x, 0.975, na.rm = TRUE)),

    # For max_tes
    cv_lower_ci_tes = map_dbl(cv_bootstrap_tes, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci_tes = map_dbl(cv_bootstrap_tes, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci_tes = map_dbl(pv_bootstrap_tes, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci_tes = map_dbl(pv_bootstrap_tes, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci_tes = map_dbl(v2_bootstrap_tes, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci_tes = map_dbl(v2_bootstrap_tes, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  select(-cv_bootstrap_quantity, -pv_bootstrap_quantity, -v2_bootstrap_quantity,
         -cv_bootstrap_tes, -pv_bootstrap_tes, -v2_bootstrap_tes) %>%
  ungroup()

# Merge back with the original dataset and cleanup
max.sec.v <- left_join(max.sec, sec.var, by = c("primary_treatment", "band_number"))

# Summarize and print the final table
summary_tibble <- sec.var %>%
  group_by(primary_treatment) %>%
  summarize(
    mean_bird_cv_quantity = mean(bird_cv_quantity, na.rm = TRUE),   # Mean CV for max_quantity
    mean_bird_sd_quantity = mean(bird_sd_quantity, na.rm = TRUE),   # Mean SD for max_quantity
    mean_bird_pv_quantity = mean(bird_pv_quantity, na.rm = TRUE),   # Mean PV for max_quantity
    mean_bird_v2_quantity = mean(bird_v2_quantity, na.rm = TRUE),
    
    mean_bird_cv_tes = mean(bird_cv_tes, na.rm = TRUE),             # Mean CV for max_tes
    mean_bird_sd_tes = mean(bird_sd_tes, na.rm = TRUE),             # Mean SD for max_tes
    mean_bird_pv_tes = mean(bird_pv_tes, na.rm = TRUE),             # Mean PV for max_tes
    mean_bird_v2_tes = mean(bird_v2_tes, na.rm = TRUE),
    
    avg_max_quantity = mean(max_quantity, na.rm = TRUE),            # Average max_quantity for each group
    median_max_quantity = median(max_quantity, na.rm=TRUE),
    avg_max_tes = mean(max_tes, na.rm = TRUE),                      # Average max_tes for each group
    median_max_tes = median(max_tes, na.rm=TRUE),
    n_individuals = n_distinct(band_number)                         # Number of individuals in each group
  ) %>%
  as_tibble()

# Print the summary table
print(summary_tibble)

```

```{r Maximum Pathogen Load and Variability}
ggplot(summary_tibble)+
  geom_point(aes(x="CV", y= mean_bird_cv_quantity, color=primary_treatment))+
  geom_point(aes(x="PV", y= mean_bird_pv_quantity, color=primary_treatment))+
  geom_point(aes(x="V2", y= mean_bird_v2_quantity, color=primary_treatment))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Pathogen Load Variability", y="Variability", x="Metric")

ggplot(max.sec.v %>% filter(primary_treatment != "Sham"))+
  geom_point(aes(x="CV", y= bird_cv_quantity, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="CV", ymin=cv_lower_ci_quantity, ymax=cv_upper_ci_quantity, y=bird_cv_quantity, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="PV", y= bird_pv_quantity, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="PV", ymin=pv_lower_ci_quantity, ymax=pv_upper_ci_quantity, y=bird_pv_quantity, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="V2", y= bird_v2_quantity, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="V2", ymin=v2_lower_ci_quantity, ymax=v2_upper_ci_quantity, y=bird_v2_quantity, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Pathogen Load Variability", y="Variability", x="Metric", color="Primary Treatment")

ggplot()+
  geom_jitter(data=sec.var, aes(x=primary_treatment, y=max_quantity, color=primary_treatment), width=0.05, size=2, alpha=0.5)+
  geom_point(data=summary_tibble, aes(x=primary_treatment, y=median_max_quantity, color=primary_treatment), shape=18, size=3)+
  geom_point(data=summary_tibble, aes(x=primary_treatment, y=avg_max_quantity, color=primary_treatment), shape=17, size=3)+
  scale_color_manual(values=c(pri_colors))+
  labs(y="Max Pathogen Load", x="Primary Treatment", color="Primary Treatment", title = "Maximum Pathogen Load of Reinfected Birds")
```

```{r}
ggplot(summary_tibble)+
  geom_point(aes(x="CV", y= mean_bird_cv_tes, color=primary_treatment))+
  geom_point(aes(x="PV", y= mean_bird_pv_tes, color=primary_treatment))+
  geom_point(aes(x="V2", y= mean_bird_v2_tes, color=primary_treatment))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Eye Score Variability", y="Variability", x="Metric")

ggplot(max.sec.v)+
  geom_point(aes(x="CV", y= bird_cv_tes, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="CV", ymin=cv_lower_ci_tes, ymax=cv_upper_ci_tes, y=bird_cv_tes, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="PV", y= bird_pv_tes, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="PV", ymin=pv_lower_ci_tes, ymax=pv_upper_ci_tes, y=bird_pv_tes, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="V2", y= bird_v2_tes, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="V2", ymin=v2_lower_ci_tes, ymax=v2_upper_ci_tes, y=bird_v2_tes, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Eye Score Variability", y="Variability", x="Metric", color="Primary Treatment")

ggplot()+
  geom_jitter(data=sec.var, aes(x=primary_treatment, y=max_tes, color=primary_treatment), width=0.1, height=0, size=2, alpha=0.5)+
  geom_point(data=summary_tibble, aes(x=primary_treatment, y=avg_max_tes, color=primary_treatment), shape=17, size=3)+
  scale_color_manual(values=c(pri_colors))+
  labs(y="Max Eye Score", x="Primary Treatment", color="Primary Treatment", title = "Maximum Eye Score of Reinfected Birds")
```

