---
title: "Sex Differences"
author: "Jesse Garrett-Larsen"
date: "2024-01-31"
output: 
  html_document: 
    keep_md: yes
---

Load packages
```{r Load Packages}
rm(list=ls())
library(tidyverse)
library(dplyr)
#install.packages("glmmTMB")
library(glmmTMB)
library(effects)
library(AICcmodavg)
#install.packages("emmeans")
library(emmeans)
library(DHARMa)
```

```{r Read In Data + Merge}
master <- read.csv("EEID2021_master_data_20220503.csv")
ab <- read.csv("EEID_1a_ELISA.csv")
m.ab <- left_join(master, ab, by = "bird_ID", keep=TRUE)
```
Format Data
```{r Format}
m.ab[m.ab == "f "] <- "f"

m.ab <- m.ab %>%
  dplyr::select(- date.y, -bird_ID.y, -band_number.y, -dpi, -(X:X.13)) #remove extra columns

#rename columns
names(m.ab)[names(m.ab) == "date.x"] <- "date"
names(m.ab)[names(m.ab) == "band_number.x"] <- "band_number"
names(m.ab)[names(m.ab) == "bird_ID.x"] <- "bird_ID"
names(m.ab)[names(m.ab) == "Avg.OD"] <- "elisa_od"
names(m.ab)[names(m.ab) == "CV"] <- "elisa_cv"
names(m.ab)[names(m.ab) == "dppi"] <- "dpi"


m.ab$sex <- as.factor(m.ab$sex)
m.ab$sex <- factor(m.ab$sex, levels = c("f", "m"), labels = c("Female", "Male"))
m.ab$primary_treatment <- as.factor(m.ab$primary_treatment)
levels(m.ab$primary_treatment) <- c("High", "Low", "Sham")
m.ab$primary_treatment <- factor(m.ab$primary_treatment, levels = c("Sham", "Low", "High"))
m.ab$l_eye_score <- as.numeric(m.ab$l_eye_score)
m.ab$r_eye_score <- as.numeric(m.ab$r_eye_score)
#total eye score = sum of l and r eye score
m.ab$tes <- m.ab$l_eye_score + m.ab$r_eye_score

```
Generate Threshold Cutoffs for Pathogen Load and Antibody Levels
```{r Thresholds}
m.ab$threshold_cutoff = 15
m.ab$seropos_cutoff = 0.061
m.ab$pathology_cutoff = 0
```
Generate new infected and seroconversion columns based off of cutoffs
```{r Infection + Seroconversion + Eye Score}
#generate infected column - if copy number > 50, infected.
#for each timepoint only - can become infected or uninfected at next timepoint
m.ab <- m.ab %>%
  mutate(infected = ifelse(quantity>threshold_cutoff, 1, 0))

#generate infection data; if path load > 50 copies = infected
#1 = bird was successfully infected at any point during the priming phase of the experiment
#coalesce function replaces any NA with 0 here. 
#To be conservative, I am always assuming there is no path load unless it's measured with qPCR
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(infected_prim = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(infected, 0) == 1), 1, 0)) %>%
  ungroup()

#generate infection data; if path load > 50 copies any time after secondary infection -> 1
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate (infected_sec = ifelse(any(coalesce(dpi, 0) > 42 & coalesce(infected, 0) == 1), 1, 0)) %>%
  ungroup()

#generate seropositivity data; if elisa OD > 0.061 = seropositive
#1 = bird seroconverted at any point during the priming phase of the experiment
m.ab <- m.ab %>%
  mutate(seropos = ifelse(elisa_od>seropos_cutoff, 1, 0))

m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(ever_seropos = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(seropos, 0) == 1), 1, 0)) %>%
  ungroup()

#generate eye score data; if eye score > 0 = diseased
#1 = bird had an eye score at that time point
m.ab <- m.ab %>%
  mutate(diseased = ifelse(tes>pathology_cutoff, 1, 0))

#ever diseased during primary infection
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(ever_diseased = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(diseased, 0) == 1), 1, 0)) %>%
  ungroup()

```

```{r Eye Score Formatting}
ggplot(m.ab %>% filter(dpi < 42), aes(x=tes, y=tes, color=band_number))+
  geom_jitter(height=0.1)+
  facet_wrap(~infected_prim)
#2415 has eye score but no path load - super resistant?
```
```{r Omit 2505}
#check for seropositive birds from quarantine from dataset
m.ab %>%
  filter(dpi == -8) %>%
  filter(elisa_od >= 0.061)%>%
  dplyr::select(band_number, elisa_od, primary_treatment, secondary_treatment, infected, seropos, dpi, ELISA.Run.Date)

#omit 2505 from data set for analysis as it was positive at quarantine
m.ab <- m.ab %>%
  filter(band_number != 2505)

#which birds are seropositive but qPCR negative?
m.ab%>%
  filter(infected_prim == 0 & elisa_od > 0.061 & dpi < 42)%>%
  dplyr::select(band_number, dpi, tes, elisa_od, quantity, primary_treatment)
```

```{r Format DPI}
#some sample days were split between two days. 
#One day experimental birds were measured (dpi -8 and dpi 14), the next day infected birds were measured (dpi -7, dpi15)
#this code makes a new column dpi.new that combines dpi 14 and 15
m.ab$dpi.new <- ifelse(m.ab$dpi %in% c(14, 15), "14", as.integer(m.ab$dpi))

#i then use dpi.new to replace the dpi column to combine dpi -7 and -8
m.ab$dpi <- ifelse(m.ab$dpi %in% c(-7, -8), "-8", as.integer(m.ab$dpi.new))

m.ab$dpi.new <- NULL #delete dpi.new column

m.ab$dpi <- as.numeric(m.ab$dpi)

m.ab <- m.ab %>%
  filter(experiment_location == "vt") #include only birds at VT
```
General overview of antibody responses
```{r, warning=FALSE}
ggplot(m.ab %>% filter(dpi <= 41), aes(x=dpi, y=elisa_od, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_hline(yintercept = 0.061)+
  stat_smooth(aes(x=dpi, y=elisa_od, group=band_number), alpha=0.5, size=0.5, span=1)+
  #scale_color_manual(values = c("red", "green4","blue"))+
  labs(x="Days Post Infection", y="ELISA OD", color="Primary Treatment")+
  facet_wrap(~primary_treatment)
```

Sex differences during primary infection
```{r Sex differences during primary infection: Pathogen Load}
#fraction that got infected during primary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment) %>%
  mutate(prim_prop = sum(infected_prim)/n())

#by sex
m.ab <- m.ab %>%
  group_by(primary_treatment, sex) %>%
  mutate(prim_prop_sex = sum(infected_prim)/n())

#by dpi
m.ab <- m.ab %>%
  group_by(primary_treatment, dpi, sex) %>%
  mutate(prim_prop_dpi = sum(infected)/n())

ggplot(m.ab %>% filter(dpi < 42), aes(x=primary_treatment, y=prim_prop, color=primary_treatment))+
  geom_point()+
    labs(x="Primary Treatment", y="Proportion Infected During Primary Infection", color= "Primary Treatment")

ggplot(m.ab %>% filter(dpi < 42), aes(x=primary_treatment, y=prim_prop_sex, color=primary_treatment))+
  geom_point()+
      labs(x="Primary Treatment", y="Proportion Infected During Primary Infection", color= "Primary Treatment")+
  facet_wrap(~sex)

#Proportion infected day 7 (qPCR) by sex
ggplot(m.ab %>% filter(dpi ==7), aes(x=primary_treatment, y=prim_prop_dpi, color=fct_rev(primary_treatment)))+
  geom_point()+
  labs(x="Primary Treatment", y="Proportion Infected Day 7", color= "Primary Treatment")+
  facet_wrap(~sex)

```

Sex differences during primary infection: Eye Score
```{r, warning=FALSE}
#fraction that were diseased during primary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment) %>%
  mutate(prim_prop_dis = sum(ever_diseased)/n())

#by sex
m.ab <- m.ab %>%
  group_by(primary_treatment, sex) %>%
  mutate(prim_prop_dis_sex = sum(ever_diseased)/n())

#by dpi
m.ab <- m.ab %>%
  group_by(primary_treatment, dpi) %>%
  mutate(prim_prop_dis_dpi = sum(diseased)/n())

#by dpi x sex
m.ab <- m.ab %>%
  group_by(primary_treatment, dpi, sex) %>%
  mutate(prim_prop_dis_dpi_sex = sum(diseased)/n())

#all birds
ggplot(m.ab %>% filter(dpi < 42), aes(x=primary_treatment, y=prim_prop_dis, color=primary_treatment))+
  geom_point()+
    labs(x="Primary Treatment", y="Proportion Diseased During Primary Infection", color= "Primary Treatment")

#by sex
ggplot(m.ab %>% filter(dpi < 42), aes(x=primary_treatment, y=prim_prop_dis_sex, color=fct_rev(primary_treatment)))+
  geom_point()+
      labs(title="Proportion with Eye Score > 0 during Primary Infection", x="Primary Treatment", y="Proportion Diseased During Primary Infection", color= "Primary Treatment")+
  facet_wrap(~sex)

#by day
ggplot(m.ab %>% filter(dpi < 42), aes(x=dpi, y=prim_prop_dis_dpi, color=fct_rev(primary_treatment)))+
  geom_point()+
  stat_smooth(aes(group=primary_treatment), method="loess", span=0.5)+
  labs(x="Primary Treatment", y="Proportion Diseased", color= "Primary Treatment")

#by day x sex
ggplot(m.ab %>% filter(dpi < 42), aes(x=dpi, y=prim_prop_dis_dpi_sex, color=fct_rev(primary_treatment)))+
  geom_point()+
  stat_smooth(aes(group=primary_treatment), method="loess", span=0.5)+
  labs(x="Primary Treatment", y="Proportion Diseased", color= "Primary Treatment")+
  facet_wrap(~sex)



```
Duration of disease
```{r}
m.ab$dpi <- as.numeric(m.ab$dpi)
diseased_duration <- m.ab %>%
  filter(diseased == 1) %>%
  group_by(band_number, primary_treatment, sex) %>%
  summarize(diseased_duration = max(dpi) - min(dpi))
diseased_duration
diseased_duration$diseased_duration <- as.numeric(diseased_duration$diseased_duration)


hist(diseased_duration$diseased_duration, breaks = 10)
ggplot(diseased_duration, aes(x=diseased_duration))+
  geom_histogram(aes(fill=sex), position="dodge")
```

Sex differences during primary infection: Antibodies
```{r}
#fraction that seroconverted during primary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment) %>%
  mutate(prim_prop_ab = sum(ever_seropos)/n())

#by sex
m.ab <- m.ab %>%
  group_by(primary_treatment, sex) %>%
  mutate(prim_prop_ab_sex = sum(ever_seropos)/n())
summary(m.ab$prim_prop_ab_sex)
#by dpi
m.ab <- m.ab %>%
  group_by(primary_treatment, dpi) %>%
  mutate(prim_prop_ab_dpi = sum(seropos)/n())

summary(m.ab$prim_prop_ab_dpi)

#by dpi x sex
m.ab <- m.ab %>%
  group_by(primary_treatment, dpi, sex) %>%
  mutate(prim_prop_ab_dpi_sex = sum(seropos)/n())

#all birds
ggplot(m.ab %>% filter(dpi < 42), aes(x=primary_treatment, y=prim_prop_ab, color=primary_treatment))+
  geom_point()+
    labs(x="Primary Treatment", y="Proportion Seropositive During Primary Infection", color= "Primary Treatment")

#by sex
ggplot(m.ab %>% filter(dpi < 42), aes(x=primary_treatment, y=prim_prop_ab_sex, color=fct_rev(primary_treatment)))+
  geom_point()+
      labs(title="Proportion Seropositive during Primary Infection", x="Primary Treatment", y="Proportion Seropositive During Primary Infection", color= "Primary Treatment")+
  facet_wrap(~sex)

#by day
ggplot(m.ab %>% filter(dpi < 42), aes(x=dpi, y=prim_prop_ab_dpi, color=fct_rev(primary_treatment)))+
  geom_point()+
  #stat_smooth(aes(group=primary_treatment), method="loess", span=0.75)+
  labs(x="Primary Treatment", y="Proportion Seropositive", color= "Primary Treatment")

#by day x sex
ggplot(m.ab %>% filter(dpi < 42), aes(x=dpi, y=prim_prop_ab_dpi_sex, color=fct_rev(primary_treatment)))+
  geom_point()+
  stat_smooth(aes(group=primary_treatment), method="loess", span=1)+
  labs(x="Primary Treatment", y="Proportion Seropositive", color= "Primary Treatment")+
  facet_wrap(~sex)
```

Sex differences during secondary infection
```{r Sex differences during secondary infection: Pathogen Load, warning=FALSE}
m.ab$secondary_dose <- as.numeric(m.ab$secondary_dose)
ggplot(m.ab %>% filter(dpi == 56), aes(x=sex, y=elisa_od, color = sex))+
  geom_jitter(alpha=0.25, width=0.25)+
  geom_hline(yintercept=0.061, alpha=0.5, linetype="dashed")+
  stat_summary(aes(x=sex, y=elisa_od, group = secondary_treatment), geom="point", shape = "-", fun=mean, size=10, alpha=1)+
  labs(x="Days Post Infection", y="Antibody Levels", color= "Sex")+
  facet_wrap(~primary_treatment~secondary_dose, ncol=5)

#fraction that got infected during secondary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment, secondary_dose) %>%
  mutate(sec_prop = sum(infected_sec)/n())

#proportion infected during secondary treatment
ggplot(m.ab %>% filter(dpi > 42), aes(x=as.factor(secondary_dose), y=sec_prop, color=fct_rev(primary_treatment)))+
  geom_point()+
  labs(title = "Proportion Infected After Secondary Inoculation", y="Proportion Infected", x= "Secondary Dose", color= "Primary Treatment")

#by sex
#fraction that got infected during secondary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment, secondary_dose, sex) %>%
  mutate(sec_prop_sex = sum(infected_sec)/n())

ggplot(m.ab %>% filter(dpi > 42), aes(x=as.factor(secondary_dose), y=sec_prop_sex, color=fct_rev(primary_treatment)))+
  geom_point()+
  labs(title = "Proportion Infected After Secondary Inoculation", y="Proportion Infected", x= "Secondary Dose", color= "Primary Treatment")+
  facet_wrap(~sex, nrow=2)

#by dpi
#fraction that got infected during secondary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment, secondary_dose, dpi) %>%
  mutate(sec_prop_dpi = sum(infected)/n())


ggplot(m.ab %>% filter(dpi > 42), aes(x=as.factor(secondary_dose), y=sec_prop_dpi, color=fct_rev(primary_treatment)))+
  geom_point()+
  labs(title = "Proportion Infected After Secondary Inoculation", y="Proportion Infected", x= "Secondary Dose", color= "Primary Treatment")+
  facet_wrap(~dpi, nrow=1)

#by dpi and sex
#fraction that got infected during secondary treatment by group
m.ab <- m.ab %>%
  group_by(primary_treatment, secondary_dose, dpi, sex) %>%
  mutate(sec_prop_dpi_sex = sum(infected)/n())

ggplot(m.ab %>% filter(dpi > 42), aes(x=as.factor(secondary_dose), y=sec_prop_dpi_sex, color=fct_rev(primary_treatment)))+
  geom_point(alpha=0.5)+
  labs(title = "Proportion Infected After Secondary Inoculation", y="Proportion Infected", x= "Secondary Dose", color= "Primary Treatment")+
  facet_grid(~sex~dpi)


ggplot(m.ab %>% filter(dpi > 42), aes(x=dpi, y=sec_prop_dpi_sex, color=(as.factor(secondary_dose))))+
  scale_color_manual(values=c("white", "lightblue", "green4", "purple4", "red3"))+
  geom_point(alpha=0.5)+
  geom_line()+
  labs(title = "Proportion Infected After Secondary Inoculation", y="Proportion Infected", x= "DPI", color= "Secondary Dose")+
  facet_grid(~primary_treatment ~sex)+
  theme_bw()

a <- m.ab %>%
  group_by(secondary_dose, primary_treatment)%>%
  filter(dpi > 42 & sex == "Female")%>%
  summarise(n())
a


#eye score
ggplot(m.ab %>% filter(dpi > 42), aes(x=dpi, y=tes, color=(as.factor(secondary_dose))))+
  scale_color_manual(values=c("white", "lightblue", "green4", "purple4", "red3"))+
  geom_point(alpha=0.5)+
  stat_smooth(aes(group = as.factor(band_number), na.rm=T, span=1))+
  labs(y="Total Eye Score", x= "DPI", color= "Secondary Dose")+
  facet_grid(primary_dose ~ sex)+
  theme_bw()