---
title: "EEID 1A Antibody Analysis"
author: "Jesse Garrett-Larsen"
date: "2024-01-26"
fig_width: 10 
fig_height: 6
output: 
  html_document: 
---

Load packages
```{r Load Packages}
rm(list=ls())
library(tidyverse)
library(dplyr)
#install.packages("glmmTMB")
library(glmmTMB)
library(effects)
library(AICcmodavg)
#install.packages("emmeans")
library(emmeans)
library(DHARMa)
```

```{r Read In Data + Merge}
master <- read.csv("EEID2021_master_data_20220503.csv")
ab <- read.csv("EEID_1a_ELISA.csv")
m.ab <- left_join(master, ab, by = "bird_ID", keep=TRUE)
```
Format Data
```{r Format}
m.ab[m.ab == "f "] <- "f"

m.ab <- m.ab %>%
  dplyr::select(- date.y, -bird_ID.y, -band_number.y, -dpi, -(X:X.13)) #remove extra columns

#rename columns
names(m.ab)[names(m.ab) == "date.x"] <- "date"
names(m.ab)[names(m.ab) == "band_number.x"] <- "band_number"
names(m.ab)[names(m.ab) == "bird_ID.x"] <- "bird_ID"
names(m.ab)[names(m.ab) == "Avg.OD"] <- "elisa_od"
names(m.ab)[names(m.ab) == "CV"] <- "elisa_cv"
names(m.ab)[names(m.ab) == "dppi"] <- "dpi"


m.ab$sex <- as.factor(m.ab$sex)
m.ab$sex <- factor(m.ab$sex, levels = c("f", "m"), labels = c("Female", "Male"))
m.ab$primary_treatment <- as.factor(m.ab$primary_treatment)
levels(m.ab$primary_treatment) <- c("High", "Low", "Sham")
m.ab$primary_treatment <- factor(m.ab$primary_treatment, levels = c("Sham", "Low", "High"))
m.ab$l_eye_score <- as.numeric(m.ab$l_eye_score)
m.ab$r_eye_score <- as.numeric(m.ab$r_eye_score)
#total eye score = sum of l and r eye score
m.ab$tes <- m.ab$l_eye_score + m.ab$r_eye_score
m.ab$band_number <- as.factor(m.ab$band_number)
m.ab$secondary_dose <- as.numeric(m.ab$secondary_dose)
```

```{r Format DPI}
#some sample days were split between two days. 
#One day experimental birds were measured (dpi -8 and dpi 14), the next day infected birds were measured (dpi -7, dpi15)
#this code makes a new column dpi.new that combines dpi 14 and 15
m.ab$dpi.new <- ifelse(m.ab$dpi %in% c(14, 15), "14", as.integer(m.ab$dpi))

#i then use dpi.new to replace the dpi column to combine dpi -7 and -8
m.ab$dpi <- ifelse(m.ab$dpi %in% c(-7, -8), "-8", as.integer(m.ab$dpi.new))

m.ab$dpi.new <- NULL #delete dpi.new column

m.ab$dpi <- as.numeric(m.ab$dpi)

#make baseline eyescore 0
m.ab <- m.ab %>%
  mutate(tes = ifelse(dpi == -8 & is.na(tes), 0, tes))

m.ab <- m.ab %>%
  filter(experiment_location == "vt") #include only birds at VT
```

Format Graphs
```{r}
#set color scheme primary treatment
pri_colors <- c("#619CFF", "#00BA38", "#F8766D")
#set color scheme secondary treatment
sec_colors <- c("#8C754B", "#77AB59", "#59A5D8", "#9F77D9", "#FA8072")
sec_default <- c( "#00BA38", "#619CFF", "#F564E3", "#BA00FF", "#F8766D")
def_colors <- c("#F8766D", "#00BA38", "#619CFF", "#F564E3", "#BA00FF", "#7A76A6")


```

Generate Threshold Cutoffs for Pathogen Load and Antibody Levels
```{r Thresholds}
m.ab$threshold_cutoff = 50
m.ab$seropos_cutoff = 0.061
m.ab$pathology_cutoff = 0
```
Generate new infected and seroconversion columns based off of cutoffs
```{r Infection + Seroconversion + Eye Score}
#generate infected column - if copy number > 50, infected.
#for each timepoint only - can become infected or uninfected at next timepoint
m.ab <- m.ab %>%
  mutate(infected = ifelse(quantity>threshold_cutoff, 1, 0))

#generate infection data; if path load > 50 copies = infected
#1 = bird was successfully infected at any point during the priming phase of the experiment
#coalesce function replaces any NA with 0 here. 
#To be conservative, I am always assuming there is no path load unless it's measured with qPCR
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(infected_prim = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(infected, 0) == 1), 1, 0)) %>%
  ungroup()

#generate infection data; if path load > 50 copies any time after secondary infection -> 1
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate (infected_sec = ifelse(any(coalesce(dpi, 0) > 42 & coalesce(infected, 0) == 1), 1, 0)) %>%
  ungroup()

#generate seropositivity data; if elisa OD > 0.061 = seropositive
#can become seropos or not at any point
m.ab <- m.ab %>%
  mutate(seropos = ifelse(elisa_od>seropos_cutoff, 1, 0))

#Seropos at any time during the priming phase
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(seropos_prim = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(seropos, 0) == 1), 1, 0)) %>%
  ungroup()

#Seropos at any time during the secondary phase
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(seropos_sec = ifelse(any(coalesce(dpi, 0) > 42 & coalesce(seropos, 0) == 1), 1, 0)) %>%
  ungroup()

#generate eye score data; if eye score > 0 = diseased
#1 = bird had an eye score at that time point
m.ab <- m.ab %>%
  mutate(diseased = ifelse(tes>pathology_cutoff, 1, 0))

#ever diseased during primary infection
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(diseased_prim = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(diseased, 0) == 1), 1, 0)) %>%
  ungroup()

#ever diseased during secondary infection
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(diseased_sec = ifelse(any(coalesce(dpi, 0) > 42 & coalesce(diseased, 0) == 1), 1, 0)) %>%
  ungroup()

#Rules for infection: We defined a bird as infected if it had eye score > 0, pathogen load > 50 copies, or both
m.ab <- m.ab %>%
  mutate(inf = ifelse((coalesce(quantity, 0) > threshold_cutoff | coalesce(tes, 0) > pathology_cutoff), 1, 0))

#ever inf during primary infection
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(inf_prim = ifelse(any(coalesce(dpi, 0) < 42 & coalesce(inf, 0) == 1), 1, 0)) %>%
  ungroup()

#ever inf during secondary infection
m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(inf_sec = ifelse(any(coalesce(dpi, 0) > 42 & coalesce(inf, 0) == 1), 1, 0)) %>%
  ungroup()

#Rules for infection: We defined a bird as infected if it had eye score > 0, pathogen load > 50 copies, or both
ggplot(m.ab%>% filter(dpi <= 41), aes(x=dpi, y=tes, color=primary_treatment))+
  geom_jitter(height=0)+
  facet_wrap(~inf)

ggplot(m.ab%>% filter(dpi <= 41), aes(x=dpi, y=quantity, color=primary_treatment))+
  geom_jitter(height=0)+
  facet_wrap(~inf)

ggplot(m.ab%>% filter(dpi <= 41), aes(x=tes, y=quantity, color=primary_treatment))+
  geom_point(height=0)+
  facet_wrap(~inf)

who <- m.ab%>%
  filter(dpi <= 41 & inf==0 & tes ==0.5)
who

ggplot(m.ab%>% filter(dpi <= 41), aes(x=dpi, y=tes, color=primary_treatment))+
  geom_jitter(height=0)+
  facet_wrap(~diseased)

```

```{r Eye Score Formatting}
ggplot(m.ab %>% filter(dpi < 42), aes(x=tes, y=tes, color=band_number))+
  geom_jitter(height=0.1)+
  facet_wrap(~inf_prim)
#2415 has eye score but no path load - super resistant?
```
```{r Omit Birds}
#check for seropositive birds from quarantine from dataset
m.ab %>%
  filter(dpi == -8) %>%
  filter(elisa_od >= 0.061)%>%
  dplyr::select(band_number, elisa_od, primary_treatment, secondary_treatment, inf, seropos, dpi, ELISA.Run.Date)

#omit 2505 from data set for analysis as it was positive at quarantine
m.ab <- m.ab %>%
  filter(band_number != 2505)

#which birds are seropositive but qPCR negative?
m.ab%>%
  filter(infected_prim == 0 & elisa_od > 0.061 & dpi < 42)%>%
  dplyr::select(band_number, dpi, tes, elisa_od, quantity, primary_treatment)

#omit birds that were not recovered by secondary infection: 2274, 2514, 2469, 2520
not_recovered <- m.ab %>% 
  filter(band_number %in% c(2274, 2514, 2469, 2520))

#confirm quantity > cutoff at dpi 41
ggplot(not_recovered, aes(x=dpi, y=quantity, color=band_number))+
  geom_point()+
  ylim(min=0, max=400)
```

General overview of antibody responses
```{r, warning=FALSE}
ggplot(m.ab %>% filter(dpi <= 41), aes(x=dpi, y=elisa_od, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_hline(yintercept = 0.061)+
  stat_smooth(aes(x=dpi, y=elisa_od, group=band_number), alpha=0.5, size=0.5, span=1)+
  scale_color_manual(values = pri_colors)+
  labs(x="Days Post Infection", y="Antibody Levels", color="Primary Treatment")+
  facet_wrap(~primary_treatment)

ggplot(m.ab %>% filter(dpi <= 41), aes(x=dpi, y=elisa_od, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_hline(yintercept = 0.061)+
  geom_path(aes(x=dpi, y=elisa_od, group=band_number), alpha=0.5, size=0.5)+
  scale_color_manual(values = pri_colors)+
  labs(x="Days Post Infection", y="Antibody Levels", color="Primary Treatment")

plot <- m.ab %>%
  filter(dpi %in% c(-8, 14, 41))
ggplot(plot, aes(x = dpi, y = elisa_od, color = fct_rev(primary_treatment))) +
  geom_point() +
  geom_hline(yintercept = 0.061, linetype="dashed") +
  geom_line(aes(group=band_number), alpha = 0.5, size = 0.5) +
  stat_summary(aes(group=primary_treatment), geom="line", fun="mean", alpha=1, size=0.5, color="black")+
  stat_summary(aes(group=primary_treatment), geom="line", fun="mean", alpha=0.5)+
  scale_color_manual(values = pri_colors) +
  labs(x = "Days Post Infection", y = "Antibody Levels", color = "Primary Treatment")
str(m.ab)
```
Duration of disease based off of eye score during primary infection
```{r}
m.ab$dpi <- as.numeric(m.ab$dpi)
d.ab <- m.ab %>%
  filter(dpi < 42 & dpi > 0) %>%
  dplyr::select(band_number, dpi, primary_treatment, sex, inf, quantity, diseased)
d.ab

ggplot(d.ab, aes(x=dpi, y=inf))+
  geom_jitter()+
  facet_wrap(~primary_treatment)

m.ab %>% 
  filter(dpi == 41 & inf == 1)

m.ab %>%
  group_by(band_number) %>%
  summarize(inf_count = sum(inf == 1, na.rm = TRUE))

m.ab %>%
  filter(dpi <= 41)%>%
  group_by(band_number, primary_treatment, sex) %>%
  summarize(a=first(dpi[diseased == 1]),
            b=last(dpi[diseased == 1]))

diseased_duration <- m.ab %>%
  filter(dpi <=41)%>%
  group_by(band_number, primary_treatment, sex) %>%
  summarize(diseased_duration = ifelse(sum(diseased == 1) == 0, 0, last(dpi[diseased == 1] + 1) - first(dpi[diseased == 1]))) #add 1 to last diseased to account for birds that only had eye score for 1 day (dpi 7 is first and last eye score > would come up as 0, but had an eye score for at least one day)
diseased_duration
diseased_duration$diseased_duration <- as.numeric(diseased_duration$diseased_duration)


hist(diseased_duration$diseased_duration, breaks = 41)

#histogram showing length with eye score > 0 during primary infection
g.primary.duration <- ggplot(diseased_duration, aes(x=diseased_duration))+
  geom_histogram(aes(fill=fct_rev(primary_treatment)), position="dodge", bins = 48)+
  labs(title = "Duration of Pathology: Primary Infection", x="Duration of Pathology", y="Count", fill="Primary Treatment")
g.primary.duration

ggplot(m.ab, aes(x=dpi, y=tes))+
  geom_jitter()+
  facet_wrap(~primary_treatment)

```
####Does Inoculation Dose Predict Antibody Levels?####
-----
```{r New df p.ab for primary }
#data frame with only primary infection (no PID 56)

p.ab <- m.ab %>%
  filter(dpi <=41)
```

I use a gamma distribution for the antibody data
With models spanning multiple days, band_number is included as a random effect

Do I use a log link function or inverse link function? Compare AIC and logLik
Skipping ahead here with models, but showing that *inverse link works better*
```{r Inverse vs Log Link}
#pre-infection
lm0a <- glm(elisa_od~primary_treatment, data=p.ab %>% filter(dpi <0), family=Gamma())
lm0b <- glm(elisa_od~primary_treatment, data=p.ab %>% filter(dpi <0), family=Gamma(log))

aictab(cand.set=list(lm0a, lm0b), modnames=c("inv","log"))
#no difference between the two

#Across all infection
ps1 <- glmmTMB(elisa_od~primary_treatment + (1|band_number), data=p.ab, family=Gamma())
ps2 <- glmmTMB(elisa_od~primary_treatment + (1|band_number), data=p.ab, family=Gamma(log))

aictab(cand.set=list(ps1, ps2), modnames=c("inv","log"))
#inverse link function is slightly better

#DPI 14
p1a <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi ==14), family=Gamma())
p2a <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi ==14), family=Gamma(log))

aictab(cand.set=list(p1a, p2a), modnames=c("inv","log"))
#no difference 

#DPI 41
p1b <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi ==41), family=Gamma())
p2b <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi ==41), family=Gamma(log))

aictab(cand.set=list(p1b, p2b), modnames=c("inv","log"))
#no difference
```

Baseline antibody levels were not significantly different from each other before inoculation.

emmeans shows estimated marginal means for variables and allows for post-hoc pairwise comparisons using the Tukey-Kramer method to minimize type 1 error.
```{r Baseline antibody levels were not significantly different}
lm0 <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi <0), family=Gamma())
summary(lm0)
plot(allEffects(lm0))

emm_r <- emmeans(lm0, ~primary_treatment)
pairs(emm_r)

```

Using primary_treatment (categorical), I show that birds inoculated with MG produce antibodies in a dose-dependent manner.

Across all days primary infection, inoculation dose 
```{r All days, fig.height=6, fig.width=10}
lm1 <- glmmTMB(elisa_od~primary_treatment + (1|band_number), data=p.ab, family=Gamma())
summary(lm1)

l1r <- simulateResiduals(lm1)
plot(l1r)

#model comparison 
p1 <- glmmTMB(elisa_od~primary_treatment + (1|band_number), data=p.ab, family=Gamma())
p2<- glmmTMB(elisa_od~primary_treatment + sex + (1|band_number), data=p.ab, family=Gamma())
p3<- glmmTMB(elisa_od~primary_treatment * sex + (1|band_number), data=p.ab, family=Gamma())
p4 <- glmmTMB(elisa_od~1 + (1|band_number), data=p.ab, family=Gamma())
p5 <- glmmTMB(elisa_od~primary_treatment + mass + (1|band_number), data=p.ab, family=Gamma())

aictab(cand.set=list(p1, p2, p3, p4, p5), modnames=c("p1","p2", "p3", "p4", "p5"))

emm_results <- emmeans(lm1, ~ primary_treatment, scale="response")
pairs(emm_results)

#antibodies across all days primary infection
ggplot(data=p.ab, aes(x=dpi, y= elisa_od, shape=primary_treatment,
                      color=fct_rev(primary_treatment)))+
  geom_hline(yintercept = 0.061, linetype="dashed", alpha=0.5)+
  geom_point(size=1.5)+
  geom_line(aes(x=dpi, y=elisa_od, group = as.factor(band_number)), alpha=0.3)+
  stat_summary(aes(group=primary_treatment), fun=mean, geom="line", color="black", size=0.5, alpha=1)+
  stat_summary(aes(group=primary_treatment), fun=mean, geom="line", alpha=0.5, size=1)+
  stat_summary(aes(group=primary_treatment, shape = primary_treatment), fun.y=mean, geom="point", shape=3, size=1.5, stroke=1.5, color="black")+
  stat_summary(aes(group=primary_treatment, shape = primary_treatment), fun.y=mean, geom="point", shape=3, size=2.5, stroke=1.5, alpha=0.5)+
    labs(title = "Antibody Levels Primary Infection", y="Antibody Levels", 
       x= "Days Post Infection", shape="Primary Treatment", color="Primary Treatment")+
  guides(shape=FALSE)
```

```{r Antibody levels on day 14 post infection were sig different between groups, fig.height=6, fig.width=10}
#Final Model
lm2 <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi == 14), family=Gamma())
summary(lm2)

emm_results <- emmeans(lm2, ~ primary_treatment, scale="response")
emm_results
pairs(emm_results)

plot(allEffects(lm2))

#model comparison
pr1 <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi == 14), family=Gamma())
pr2<- glmmTMB(elisa_od~primary_treatment + sex , data=p.ab %>% filter(dpi == 14), family=Gamma())
pr3<- glmmTMB(elisa_od~primary_treatment * sex , data=p.ab %>% filter(dpi == 14), family=Gamma())
pr4 <- glmmTMB(elisa_od~primary_treatment * room , data=p.ab %>% filter(dpi == 14), family=Gamma())
pr5 <- glmmTMB(elisa_od~1, data=p.ab %>% filter(dpi == 14), family=Gamma())
pr6 <- glmmTMB(elisa_od~primary_treatment + mass, data=p.ab %>% filter(dpi == 14), family=Gamma())
pr7 <- glmmTMB(elisa_od~primary_treatment + room , data=p.ab %>% filter(dpi == 14), family=Gamma())

#AICc
aictab(cand.set=list(pr1, pr2, pr3, pr4, pr5, pr6, pr7), modnames=c("pr1","pr2", "pr3", "pr4", "pr5", "pr6", "pr7"))


#There were 15 blood samples that were lost before they were run and are not included in this graph
missing_samples <- p.ab %>%
  filter(dpi %in% c(-8, 14, 41) & is.na(elisa_od)) %>%
  dplyr::select(band_number, bird_ID, dpi, elisa_od)
print(missing_samples)

#Model predictions
dat.new=expand.grid(primary_treatment=unique(p.ab$primary_treatment))#new grid to put predictions into
dat.new$yhat = predict(lm2, type="response", newdata=dat.new, re.form=NA) #predicted values based off lm2 model
head(dat.new)

#graph showing antibody levels on day 14 post-infection with model predictions
pid14.pred <- ggplot(data=p.ab %>% filter(dpi == 14), aes(x=primary_treatment, y=elisa_od, shape=primary_treatment, color=fct_rev(primary_treatment)))+
  geom_jitter(size=2, width=0.25)+
  geom_point(data=dat.new, aes(x=primary_treatment, y=yhat, shape=primary_treatment), size=10, shape="-", color="brown")+#model predictions
  scale_shape_manual(values = c(16, 17, 15))+
  geom_hline(yintercept=0.061, linetype="dashed", alpha=0.5)+
  labs(title="MG Antibodies Day 14", x="Primary Treatment", y="Antibody Levels", color="Primary Treatment")+
  guides(shape=FALSE)

pid14.pred

```
I also looked at primary_dose, a continuous variable to test whether inoculation dose affects antibody response.

```{r Primary Dose Continuous Day 14}
#model comparison
prd1 <- glm(elisa_od~primary_dose, data=p.ab %>% filter(dpi == 14), family=Gamma())
prd2<- glm(elisa_od~primary_dose + sex , data=p.ab %>% filter(dpi == 14), family=Gamma())
prd2.5<- glm(elisa_od~primary_dose * sex , data=p.ab %>% filter(dpi == 14), family=Gamma())
prd3 <- glm(elisa_od~primary_dose + room , data=p.ab %>% filter(dpi == 14), family=Gamma())
prd4 <- glm(elisa_od~primary_dose * room , data=p.ab %>% filter(dpi == 14), family=Gamma())
prd5 <- glm(elisa_od~1, data=p.ab %>% filter(dpi == 14), family=Gamma())
prd6 <- glm(elisa_od~primary_dose + mass, data=p.ab %>% filter(dpi == 14), family=Gamma())

#AICc
aictab(cand.set=list(prd1, prd2, prd2.5, prd3, prd4, prd5, prd6), modnames=c("prd1", "prd2", "prd2.5", "prd3", "prd4", "prd5", "prd6"))

summary(prd1)

#continuous model predictions
dat.new2 <- data.frame(primary_dose = seq(min(p.ab$primary_dose), max(p.ab$primary_dose), length.out = 10))
dat.new2$yhat <- predict(prd1, type = "response", interval = "confidence", level = 0.95, newdata = dat.new2)

#plot continuous predicted values over raw data
pid14.dose.pred.c <- ggplot(data=p.ab %>% filter(dpi == 14), aes(x=primary_dose, y=elisa_od, color=primary_treatment))+
  geom_jitter(size=2, width=0.5, aes(shape=as.factor(primary_dose)))+
  geom_line(data=dat.new2, aes(x=primary_dose, y=yhat), color="brown")+#model predictions
  labs(x="Primary Dose", y="Antibody Levels", shape="Primary Dose")

pid14.dose.pred.c
```
Alternatively, I can subset the primary_dose into facors based on primary_dose. This maintains the continuous nature of the data, but allows for pairwise comparisons using emmeans.
```{r Primary Dose continuous pairwise comparisons day 14}
# Create a new factor variable based on primary_dose
p.ab <- p.ab %>% mutate(primary_dose_group = cut(primary_dose, breaks = c(-Inf, 100, 2000, Inf), labels = c("Sham", "Low", "High")))

# Fit glm with  new factor variable
prdc1 <- glm(elisa_od ~ primary_dose_group, data = p.ab %>% filter(dpi == 14), family = Gamma())

# Calculate estimated marginal means (EMMs)
emm_results <- emmeans(prdc1, specs = "primary_dose_group")

# Conduct pairwise comparisons between the groups
pairs(emm_results)

```
Variation in antibody levels on day 14 may be a better predictor of host susceptibility 
```{r Proportion Seroconverted  vs Infected day 14}
#what proportion of birds seroconverted by day 14?
library(gt)

library(gtExtras)
#proportion seropositive
m.abs<- m.ab %>%
  filter(dpi %in% c(-8, 7, 14, 41))%>%
  group_by(dpi, primary_treatment)%>%
  summarise(prop_seropos = sum(seropos == 1, na.rm=TRUE) / n(),
            n_seropos = sum(seropos == 1, na.rm=TRUE),
            n_seroneg = sum(seropos == 0, na.rm=TRUE),
            prop_inf = sum(inf == 1, na.rm=TRUE)/ n(),
            n_inf = sum(inf == 1, na.rm=TRUE),
            n_not_inf = sum(inf == 0),
            seropos_prim = sum(seropos_prim == 1, na.rm=TRUE),
            inf_prim = sum(inf_prim == 1, na.rm=TRUE),
            total_birds =n())
m.abs

table_gt <- m.abs %>%
  gt() %>%
  fmt_number(
    columns=vars(prop_seropos, prop_inf),
    decimals = 2
  )%>%
  tab_header(
    title = "Summary Table",
    subtitle = "Summary of Proportion Seropositive and Infected by Treatment Group"
  ) %>%
  cols_label(
    dpi = "Day Post Infection",
    primary_treatment = "Primary Treatment",
    prop_seropos = "Proportion Seropositive",
    n_seropos = "Number Seropositive",
    n_seroneg = "Number Seronegative",
    prop_inf = "Proportion Infected",
    n_inf = "Number Infected",
    n_not_inf = "Number Not Infected",
    seropos_prim = "Seropositive Primary",
    inf_prim = "Infected Primary",
    total_birds = "Total Birds",
  )

# Display the table
table_gt


#proportion seropositive
ggplot(m.abs, aes(x=dpi, y=prop_seropos, color=primary_treatment, shape=primary_treatment))+
  geom_point(size = 1)+
  scale_color_manual(values = pri_colors)+
  ylim(c(0,1))+
  geom_text(aes(label = round(prop_seropos, 2)), vjust = -0.25, size = 3) +
  labs(title= "Proportion Seropositive", x = "Day Post Infection", y = "Proportion Seropositive", color = "Primary Treatment", shape = "Primary Treatment")#+
  guides(shape = FALSE)

m.ab.p<- m.ab %>%
  filter(dpi %in% c(-8, 7, 14, 41)) %>%
  group_by(dpi, primary_treatment)%>%
  summarise(prop_seropos = sum(seropos == 1, na.rm=TRUE) / n(), na.rm=TRUE,
            n_seropos = sum(seropos == 1, na.rm=TRUE),
            n_seroneg = sum(seropos == 0, na.rm=TRUE),
            prop_inf = sum(inf == 1, na.rm=TRUE)/ n(), na.rm=TRUE,
            n_inf = sum(inf == 1, na.rm=TRUE),
            n_not_inf = sum(inf == 0, na.rm=TRUE),
            seropos_prim = sum(seropos_prim == 1, na.rm=TRUE),
            inf_prim = sum(inf_prim == 1, na.rm=TRUE),
            total_birds =n(), na.rm=TRUE)
m.ab.p

#proportion infected primary
ggplot(m.ab.p, aes(x=dpi, y=prop_inf, color=primary_treatment, shape=primary_treatment))+
  geom_point(size = 1)+
  scale_color_manual(values = pri_colors)+
  ylim(c(0,1))+
  geom_text(aes(label = paste(n_inf, "/", total_birds)), vjust = -0.2, size = 3) +
  #geom_text(aes(label = round(prop_inf, 2)), vjust = -0.2, size = 3) +
  labs(title = "Proportion Infected", x = "Day Post Infection", y = "Proportion Infected", color = "Primary Treatment", shape = "Primary Treatment")

#proportion seropositive primary
ggplot(m.ab.p, aes(x=dpi, y=prop_seropos, color=primary_treatment, shape=primary_treatment))+
  geom_point(size = 1)+
  scale_color_manual(values = pri_colors)+
  ylim(c(0,1))+
  geom_text(aes(label = round(prop_seropos, 2)), vjust = -0.2, size = 3) +
  labs(title = "Proportion Seropositive", x = "Day Post Infection", y = "Proportion Seropositive", color = "Primary Treatment", shape = "Primary Treatment")


m.ab.inf<- m.ab %>%
  filter(dpi %in% c(-8, 7, 14, 41, 46, 49, 56, 63)) %>%
  group_by(dpi, primary_treatment, secondary_dose)%>%
  summarise(prop_seropos = sum(seropos == 1, na.rm=TRUE) / n(), na.rm=TRUE,
            n_seropos = sum(seropos == 1, na.rm=TRUE),
            n_seroneg = sum(seropos == 0, na.rm=TRUE),
            prop_inf = sum(inf == 1, na.rm=TRUE)/ n(), na.rm=TRUE,
            n_inf = sum(inf == 1, na.rm=TRUE),
            n_not_inf = sum(inf == 0, na.rm=TRUE),
            seropos_prim = sum(seropos_prim == 1, na.rm=TRUE),
            inf_prim = sum(inf_prim == 1, na.rm=TRUE),
            total_birds =n(), na.rm=TRUE)

#proportion infected
ggplot(m.ab.inf, aes(x=dpi, y=prop_inf, color=as.factor(secondary_dose), shape=primary_treatment, groups=secondary_dose))+
  geom_point(size = 1)+
  scale_color_manual(values = sec_colors)+
  geom_path(aes(linetype=fct_rev(primary_treatment)))+
  ylim(c(0,1))+
  geom_text(aes(label = round(prop_inf, 2)), vjust = -0.2, size = 3) +
  labs(title = "Proportion Infected", x = "Day Post Infection", y = "Proportion Infected", color = "Primary Treatment", shape = "Primary Treatment")




```
If birds seroconverted, were they protected from secondary pathology?
```{r If birds seroconverted, were they protected from secondary pathology?}
#does seropositivity during primary infection predict whether or not a bird will develop pathology upon secondary infection?
lm.sp1 <- glm(diseased_sec ~ seropos_prim + primary_treatment*secondary_dose, data=m.ab, family=binomial(link= "logit"))
summary(lm.sp1)
plot(allEffects(lm.sp1))

emm_results<- emmeans(lm.sp1, ~ primary_treatment*secondary_dose)
pairs(emm_results)

dat.new=expand.grid(secondary_dose=unique(m.ab$secondary_dose),
                    primary_treatment=unique(m.ab$primary_treatment),
                    diseased_sec=unique(m.ab$diseased_sec),
                    seropos_prim=unique(m.ab$seropos_prim))#new grid to put predictions into
dat.new$yhat = predict(lm.sp1, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)

ggplot(m.ab, aes(y=diseased_sec, x=seropos_prim))+
  geom_jitter(size=1, height=0, width=0.5, aes(shape=primary_treatment, color=primary_treatment))+
  geom_point(data = dat.new, aes(x =seropos_prim, y = yhat, shape = primary_treatment, color = primary_treatment)) +
  geom_path(data = dat.new, aes(x = seropos_prim, y = yhat, color = primary_treatment, linetype = primary_treatment))+
  labs(x="Seroconversion Status During Primary Infection", y="Disease During Secondary Infection", color = "Primary Treatment", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_color_manual(values = c(pri_colors))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  facet_grid(~secondary_dose)
  
  
 

#lm.sp2 <- glm(tes ~ elisa_od + primary_treatment*secondary_dose, data=m.ab, family=binomial(link= "logit"))


#install.packages("vcd")
library(vcd)
#mosaic plot
mosaic(~primary_treatment + secondary_dose + seropos_prim + diseased_sec, data=m.ab,
       highlighting = "primary_treatment", highlighting_fill = pri_colors)

mosaic(~primary_treatment + secondary_dose + seropos_prim + diseased_sec, data=m.ab,
       highlighting = "secondary_dose", highlighting_fill = sec_colors)

mosaic(~primary_treatment + secondary_dose + inf_prim + diseased_sec, data=m.ab,
       highlighting = "primary_treatment", highlighting_fill = pri_colors)

```

```{r infection during primary vs pathology secondary}
#does being infected during primary infeciton predict whether a bird will develop pathology upon secondary infection?
m.ab.inf <- m.ab %>%
  filter(primary_treatment != "Sham")
lm.inf <- glm(diseased_sec ~ inf_prim + primary_treatment*secondary_dose, data=m.ab.inf, family=binomial(link="logit"))
summary(lm.inf)
resid <- simulateResiduals(lm.inf)
plot(resid)
plot(allEffects(lm.inf))

dat.new=expand.grid(secondary_dose=unique(m.ab.inf$secondary_dose),
                    primary_treatment=unique(m.ab.inf$primary_treatment),
                    diseased_sec=unique(m.ab.inf$diseased_sec),
                    inf_prim=unique(m.ab.inf$inf_prim))#new grid to put predictions into
dat.new$yhat = predict(lm.inf, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)




ggplot(m.ab.inf, aes(y=diseased_sec, x=inf_prim))+
  geom_jitter(size=1, height=0, width=0.5, aes(shape=primary_treatment))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, color=as.factor(secondary_dose), shape = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color=as.factor(secondary_dose), linetype = primary_treatment))+
  labs(x="Infection Status During Primary Infection", y="Disease During Secondary Infection", color = "Secondary Dose", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_linetype_manual(values = c("dashed", "solid"))+
  scale_shape_manual(values = c(17, 15))+
  facet_grid(~secondary_dose)

#with sham
lm.dis <- glm(diseased_sec ~ inf_prim + primary_treatment*secondary_dose, data=m.ab, family=binomial(link="logit"))
summary(lm.dis)
resid <- simulateResiduals(lm.dis)
plot(resid)

dat.new=expand.grid(secondary_dose=unique(m.ab$secondary_dose),
                    primary_treatment=unique(m.ab$primary_treatment),
                    diseased_sec=unique(m.ab$diseased_sec),
                    inf_prim=unique(m.ab$inf_prim))#new grid to put predictions into
dat.new$yhat = predict(lm.dis, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)

m.ab.sec<- m.ab %>%
  filter(dpi %in% c(46, 49, 56, 63)) %>%
  group_by(dpi, primary_treatment, secondary_dose, inf_prim)%>%
  summarise(prop_seropos_sec = sum(seropos == 1, na.rm=TRUE) / n(), na.rm=TRUE,
            n_seropos_sec = sum(seropos == 1, na.rm=TRUE),
            n_seroneg_sec = sum(seropos == 0, na.rm=TRUE),
            prop_inf_sec = sum(inf == 1, na.rm=TRUE)/ n(), na.rm=TRUE,
            n_inf_sec = sum(inf == 1, na.rm=TRUE),
            n_not_inf_sec = sum(inf == 0, na.rm=TRUE),
            seropos_sec = sum(seropos_sec == 1, na.rm=TRUE),
            inf_sec = sum(inf_sec == 1, na.rm=TRUE),
            dis_sec = sum(diseased_sec == 1, na.rm =TRUE),
            total_birds =n(), na.rm=TRUE)

ggplot(m.ab, aes(y=diseased_sec, x=inf_prim))+
  geom_jitter(size=1, height=0, width=0.5, aes(shape=primary_treatment))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, color=as.factor(secondary_dose), shape = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color=as.factor(secondary_dose), linetype = primary_treatment))+
  labs(x="Infection Status During Primary Infection", y="Disease During Secondary Infection", color = "Secondary Dose", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  facet_grid(~secondary_dose)

ggplot(m.ab, aes(y=diseased_sec, x=inf_prim))+
  geom_jitter(size=1, height=0, width=0.4, aes(shape=primary_treatment, color = primary_treatment))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, shape = primary_treatment, color = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color = primary_treatment, linetype = primary_treatment))+
    geom_point(data=m.ab.sec, aes(x=inf_prim, y=dis_sec/total_birds, color=primary_treatment), shape = 45, size=10)+
  labs(x="Infection Status During Primary Infection", y="Disease During Secondary Infection", color = "Primary Treatment", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_color_manual(values = c(pri_colors))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  facet_grid(~secondary_dose)
```


```{r infection during primary vs infection secondary}
#infected secondary vs infected primary
lm.dis <- glm(inf_sec ~ inf_prim + primary_treatment*secondary_dose, data=m.ab, family=binomial(link="logit"))
summary(lm.dis)
resid <- simulateResiduals(lm.dis)
plot(resid)

dat.new=expand.grid(secondary_dose=unique(m.ab$secondary_dose),
                    primary_treatment=unique(m.ab$primary_treatment),
                    inf_sec=unique(m.ab$inf_sec),
                    inf_prim=unique(m.ab$inf_prim))#new grid to put predictions into
dat.new$yhat = predict(lm.dis, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)

ggplot(m.ab, aes(y=diseased_sec, x=inf_prim))+
  geom_jitter(size=1, height=0, width=0.5, aes(shape=primary_treatment))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, color=as.factor(secondary_dose), shape = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color=as.factor(secondary_dose), linetype = primary_treatment))+
  labs(x="Infection Status During Primary Infection", y="Disease During Secondary Infection", color = "Secondary Dose", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  facet_grid(~secondary_dose)

ggplot(m.ab, aes(y=diseased_sec, x=inf_prim))+
  geom_jitter(size=1, height=0, width=0.4, aes(shape=primary_treatment, color = primary_treatment))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, shape = primary_treatment, color = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color = primary_treatment, linetype = primary_treatment))+
  labs(x="Infection Status During Primary Infection", y="Disease During Secondary Infection", color = "Primary Treatment", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_color_manual(values = c(pri_colors))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  facet_grid(~secondary_dose)

ggplot(m.ab, aes(x=inf_prim, y=inf_sec, color=primary_treatment))+
  #geom_jitter(width=0.4, height =0.05, alpha = 0.5)+
  geom_count(position=position_jitter(height = 0, width = 0.3))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, shape = primary_treatment, color = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color = primary_treatment, linetype = primary_treatment))+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  scale_color_manual(values = c(pri_colors))+
  labs(x="Infected During Primary Infection", y="Infection Status During Secondary Infection", color = "Primary Treatment", shape = "Primary Treatment", linetype = "Primary Treatment")+
  facet_wrap(~secondary_dose, nrow=1)

ggplot(m.ab, aes(x=inf_prim, y=inf_sec, color=primary_treatment))+
  geom_jitter(width=0.4, height =0, alpha = 0.5)+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, shape = primary_treatment, color = primary_treatment)) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color = primary_treatment, linetype = primary_treatment))+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("0", "1"))+
  scale_color_manual(values = c(pri_colors))+
  scale_linetype_manual(values = c("dotted", "dashed", "solid"))+
  labs(x="Infected During Primary Infection", y="Infection Status During Secondary Infection", color = "Primary Treatment", shape = "Primary Treatment", linetype = "Primary Treatment")+
  facet_wrap(~secondary_dose, nrow=1)
```

```{r eval=FALSE, include=FALSE}
#does being infected during primary infection predict whether a bird will develop pathology upon secondary infection?
m.ab.inf <- m.ab1# %>%
#  filter(primary_treatment != "Sham")
lm.inf <- glm(inf_sec ~ inf_prim + seropos_prim + primary_treatment*secondary_dose, data=m.ab.inf, family=binomial(link="logit"))
summary(lm.inf)
plot(allEffects(lm.inf))

dat.new=expand.grid(secondary_dose=unique(m.ab.inf$secondary_dose),
                    primary_treatment=unique(m.ab.inf$primary_treatment),
                    diseased_sec=unique(m.ab.inf$diseased_sec),
                    inf_prim=unique(m.ab.inf$inf_prim))#new grid to put predictions into
dat.new$yhat = predict(lm.inf, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)

ggplot(m.ab.inf, aes(y=diseased_sec, x=inf_prim))+
  geom_jitter(size=1, height=0, width=0.5, aes(shape=primary_treatment))+
  geom_point(data = dat.new, aes(x =inf_prim, y = yhat, color=as.factor(secondary_dose), shape = fct_rev(primary_treatment))) +
  geom_path(data = dat.new, aes(x = inf_prim, y = yhat, color=as.factor(secondary_dose), linetype = fct_rev(primary_treatment)))+
  labs(x="Infection Status During Primary Infection", y="Relative Likelihood of Disease During Secondary Infection", color = "Secondary Dose", shape = "Primary Treatment", linetype = "Primary Treatment")+
  scale_x_continuous(breaks = c(0, 1), labels = c("N", "Y"))+
  scale_y_continuous(breaks = c(0,1), labels = c("N", "Y"))+
  facet_grid(~secondary_dose)
```

Change in antibody levels from baseline to day 14
```{r change from baseline to day 14}
m.abb <- m.ab %>%
  filter(dpi %in% c(-8, 14)) %>%
  group_by(band_number, primary_treatment)%>%
  mutate(ab_change = elisa_od[dpi == 14] - elisa_od[dpi == -8], na.rm=T)
m.abb

ggplot(m.abb %>% filter(dpi == 14), aes(x=primary_treatment, y=ab_change, color=primary_treatment, shape = primary_treatment))+
  geom_jitter(width=0.25, height=0)+
  scale_color_manual(values = c(pri_colors))+
  labs(x="Day Post Infection", y="Change in Antibody Levels", color="Primary Treatment", shape = "Primary Treatment", title = "Change in Antibody Levels: Day 14 - Baseline")
hist(m.abb$ab_change)
lm.change <- lm(ab_change ~ primary_treatment, data=m.abb)
summary(lm.change)
resid <- simulateResiduals(lm.change)
plot(resid)
plot(allEffects(lm.change))

emm_results<- emmeans(lm.change, ~ primary_treatment)
pairs(emm_results)

```


Antibody levels return to baseline in the low (p = 0.149), but not high dose (p < 0.001) groups by day 41.
```{r Antibody levels on day 41 post infection were sig different between groups, fig.height=6, fig.width=10}
lm3 <- glm(elisa_od~primary_treatment, data=p.ab %>% filter(dpi==41), family=Gamma())
summary(lm3)

emm_results<- emmeans(lm3, ~ primary_treatment)
pairs(emm_results)

#model comparison
p1 <- glm(elisa_od~primary_treatment, data=p.ab %>% filter(dpi==41), family=Gamma())
p2<- glm(elisa_od~primary_treatment + sex , data=p.ab %>% filter(dpi==41), family=Gamma())
p3<- glm(elisa_od~primary_treatment * sex , data=p.ab %>% filter(dpi==41), family=Gamma())
p4 <- glm(elisa_od~primary_treatment + quantity , data=p.ab %>% filter(dpi==41), family=Gamma())
p5 <- glm(elisa_od~1, data=p.ab %>% filter(dpi==41), family=Gamma())
p6 <- glm(elisa_od~primary_treatment + mass, data= p.ab %>% filter(dpi==41), family=Gamma())

#AICc
aictab(cand.set=list(p1, p2, p3, p4, p5, p6), modnames=c("p1", "p2", "p3", "p4", "p5", "p6"))

summary(p2)
#p2 is the best model
#but sex adds complexity and is not significant (0.1445) so remove

#p1 is second best - use
summary(p1)

#model predictions
dat.newer=expand.grid(primary_treatment=unique(p.ab$primary_treatment))#new grid to put predictions into
dat.newer$yhat = predict(lm3, type="response", newdata=dat.newer, re.form=NA) #predicted values based off lm3  model
head(dat.newer)

#plot predicted values over raw data
pid41.pred <- ggplot(data = m.ab %>% filter(dpi == 41), aes(x = primary_treatment, y = elisa_od, shape = primary_treatment, color=fct_rev(primary_treatment))) +
  geom_jitter(size=1.5, width = 0.25) +
  geom_hline(yintercept = 0.061, color = "black", alpha = 0.75, linetype='dashed') +
  labs(title="MG Antibodies Day 41", x="Inoculation Dose", y= "MG Antibodies OD", color = "Primary Treatment")+
  geom_point(data=dat.newer, aes(x=primary_treatment, y=yhat, shape=primary_treatment), size=10, shape="-", color="brown4")+#model predictions
  guides(shape=FALSE)

pid41.pred
```

Antibody levels return to baseline in the low (p = 0.149), but not high dose (p < 0.001) groups by day 41.
```{r Primary Dose Continuous Day 41}
#model comparison
pr1 <- glm(elisa_od~primary_dose, data=p.ab %>% filter(dpi==41), family=Gamma())
pr2<- glm(elisa_od~primary_dose + sex , data=p.ab %>% filter(dpi==41), family=Gamma())
pr3<- glm(elisa_od~primary_dose * sex , data=p.ab %>% filter(dpi==41), family=Gamma())
pr4 <- glm(elisa_od~primary_dose + quantity , data=p.ab %>% filter(dpi==41), family=Gamma())
pr5 <- glm(elisa_od~1, data=p.ab %>% filter(dpi==41), family=Gamma())
pr6 <- glm(elisa_od~primary_dose + mass, data= p.ab %>% filter(dpi==41), family=Gamma())


#AICc
aictab(cand.set=list(pr1, pr2, pr3, pr4, pr5, pr6), modnames=c("pr1", "pr2", "pr3", "pr4", "pr5", "pr6"))

#pr2 best model
#but sex adds complexity and is not significant (0.146) so remove

summary(pr1)

#continuous model predictions
dat.newer2 <- data.frame(primary_dose = seq(min(p.ab$primary_dose), max(p.ab$primary_dose), length.out = 100))
dat.newer2$yhat <- predict(pr1, type = "response", interval = "confidence", level = 0.95, newdata = dat.newer2)
head(dat.newer2)

pid41.dose.pred.c <- ggplot(data=p.ab %>% filter(dpi == 41), aes(x=primary_dose, y=elisa_od))+
  geom_jitter(size=2, width=0.1)+
  geom_line(data=dat.newer2, aes(x=primary_dose, y=yhat), color="brown")+#model predictions
  labs(title = "ELISA OD PID 41", x="Primary Dose", y="ELISA OD", shape="Primary Dose")+
  geom_hline(yintercept = 0.061, color = "black", alpha = 0.75, linetype='dashed')

pid41.dose.pred.c

#pairwise comparisons
p.ab <- p.ab %>% mutate(primary_dose_group = cut(primary_dose, breaks = c(-Inf, 100, 2000, Inf), labels = c("Sham", "Low", "High")))

# Fit glm with  new factor variable
pr1 <- glm(elisa_od~primary_dose_group, data=p.ab %>% filter(dpi==41), family=Gamma())

# Calculate estimated marginal means (EMMs)
emm_results <- emmeans(pr1, specs = "primary_dose_group")

# Conduct pairwise comparisons between the groups
pairs(emm_results)

```

Does prior exposure induce heterogeneity in antibody levels?####
```{r Does prior exposure induce heterogeneity in antibody levels?, fig.height=6, fig.width=10}
p.ab <- p.ab %>%
  filter(dpi !=56)
#subset antibodies by groups to look at their variance
#subset by primary treatment (number doesn't matter) with 1|band_number as random effect
#variance of random effect should be highest in highest heterogeneity groups
library(lme4)

#across all primary infection days, which groups are most heterogeneous?
glm.all <- glmer(elisa_od~1 +(1|band_number), data=p.ab, family=Gamma())
glm.sham <- glmer(elisa_od~1 +(1|band_number), data=subset(p.ab, primary_treatment == "Sham"), family=Gamma())
glm.low <- glmer(elisa_od~1 +(1|band_number), data=subset(p.ab, primary_treatment == "Low"), family=Gamma())
glm.high <- glmer(elisa_od~1 +(1|band_number), data=subset(p.ab, primary_treatment == "High"), family=Gamma())

summary(glm.all); summary(glm.sham); summary(glm.low); summary(glm.high)

AIC(glm.sham,glm.low, glm.high)

##Days 14 and 41 only
#var <- c(11.5496, 0.284649, 8.65716, 5.8528) #variance from table above
#ci <- c(3.39885, 0.534, 2.9423, 2.4192) #Standard Deviation from table above
#g <- c("All", "Sham", "Low", "High") #Groups

##Days -8, 14, and 41 
var <- c(10.89235, 0.267613, 7.58692, 5.2232) #variance from table above
ci <- c(3.3004, 0.51731, 2.7544, 2.2854) #Standard Deviation from table above
g <- c("All", "Sham", "Low", "High") #Groups

#new df with variance and error for error bars (min/max)
variability <- data.frame(g,var, ci)

g.var.primary <- ggplot(variability %>% filter(g != "All"), aes(x=fct_rev(g), y=ci, color=g))+
  geom_point(size=3)+
  #geom_errorbar(aes(ymin=min, ymax=max), width=0.5)+ #add errorbars representing +/- 1 SD
  labs(title = "Primary Infection Variation in Antibody Levels", x="Primary Treatment", y="Standard Deviation", color="Treatment")+
  theme_minimal()

g.var.primary


 m.ab%>%
  group_by(primary_treatment)%>%
  summarise(count=n_distinct(band_number))

m.ab%>%
  group_by(primary_treatment, dpi)%>%
  summarise(count=n_distinct(band_number))

p.ab <- p.ab %>%
  filter(dpi %in% c(-8, 14, 41))%>%
  group_by(band_number) %>%
  mutate(mean_ab = mean(elisa_od))
p.ab$mean_ab
unique(p.ab$band_number)
p.ab$band_number.n<-as.numeric(p.ab$band_number)
ggplot(p.ab %>% filter(dpi <= 41 & band_number %in% c(2274, 2374, 2378)), aes(x=primary_treatment, y=elisa_od, color=fct_rev(primary_treatment), shape=band_number))+
  geom_hline(yintercept = 0.061)+
  geom_jitter(width=0.25)+
  scale_color_manual(values = pri_colors)+
  labs(x="Primary Treatment (All DPI)", y="Antibody Level", color="Primary Treatment")#+
  facet_wrap(~band_number)

```
(chatGPT)
In a mixed-effects model, the random effects capture variability among groups or clusters in the data. The variance component represents the variability of the random effects, while the standard deviation is the square root of the variance.

Mathematically, the standard deviation (σ) is the square root of the variance (σ^2), so σ = √(variance).

For example, if the variance of the random effects for a certain grouping factor is 5.2232, as in your model summary, then the standard deviation would be approximately √5.2232 ≈ 2.2854, as also shown in the summary output.

So, in summary, the standard deviation provides a more interpretable measure of the dispersion of the random effects compared to the variance, as it is in the same units as the original data.

Residual Models: Method to look at the amount of variance within each treatment group while accounting for variance from differences in dpi

```{r Variance across DPI}
####Residual Models####
p.ab <- p.ab %>%
  filter(dpi <= 41 )%>%
  drop_na(band_number, dpi, elisa_od)



#Levene's Test for Equality of Variances
library(car)
leveneTest(elisa_od ~ primary_treatment, data=p.ab)
##Variance is not equal##

#glm <- glmmTMB(elisa_od ~ 1 + (band_number|dpi), data=p.ab, family=Gamma())
glm.a <- glmmTMB(elisa_od ~ 1 + (1|dpi), data=p.ab, family=Gamma())

summary(glm.a)
resid(glm.a)
hist(resid(glm.a))
p.ab$resid <- resid(glm.a)
hist(p.ab$resid)

ggplot(p.ab, aes(x=dpi, y=elisa_od, color=fct_rev(primary_treatment)))+
  geom_point()

ggplot(p.ab, aes(x=dpi, y=resid, color=fct_rev(primary_treatment)))+
  geom_point()+
  labs(x="Days Post Infection", y="Residuals", color="Primary Treatment")+
  facet_wrap(~primary_treatment)

lm1 <- lm(abs(resid) ~ primary_treatment, data=p.ab)
summary(lm1)
```


```{r Variance across DPI}
#subset the data into primary_treatments
#look at the amount of variance within each treatment while accounting for variance from differences in dpi
s.ab <- p.ab %>%
  filter(primary_treatment=="Sham")
glm.s <- glmmTMB(elisa_od ~ 1 + (1|dpi), data=s.ab, family=Gamma())
glm.sf <- glm(elisa_od ~ 1+ dpi, data=s.ab, family=Gamma())
summary(glm.s)
summary(glm.sf)
plot(allEffects(glm.sf))
s.SD <- 0.328

s.ab <- p.ab %>%
  filter(primary_treatment=="Sham")

glm.s <- glmmTMB(elisa_od ~ 1 + (1|dpi), data=s.ab, family=Gamma())

s.ab$resid <- resid(glm.s)

glm.sr <- glm(abs(resid) ~ 1, data=s.ab, family=Gamma())

summary(glm.sr)
s.ab$residresid <- resid(glm.sr)
sd(abs(s.ab$resid))

ggplot(s.ab, aes(x=dpi, y=resid))+
  geom_point(color="blue")+
    labs(title="Sham", x="Days Post Infection", y="Residuals", color="Primary Treatment")

l.ab <- p.ab %>%
  filter(primary_treatment=="Low")
glm.l <- glmmTMB(elisa_od ~ 1 + (1|dpi), data=l.ab, family=Gamma())
glm.lf <- glm(elisa_od ~ 1+ as.factor(dpi), data=l.ab, family=Gamma())
glm.l1 <- glm(elisa_od ~ 1, data=l.ab)
summary(glm.l1)

l.ab$resid <- (resid(glm.l1))

ggplot(l.ab, aes(x=dpi, y=resid))+
  geom_point(color="green3")+
   labs(title="Low", x="Days Post Infection", y="Residuals", color="Primary Treatment")

l.SD <- 2.108

h.ab <- p.ab %>%
  filter(primary_treatment=="High")
glm.h <- glmmTMB(elisa_od ~ 1 + (1|dpi), data=h.ab, family=Gamma())
glm.h1 <- glm(elisa_od ~ 1, data=h.ab)
rh<-abs((resid(glm.h1)))
mean(rh)

ggplot(h.ab, aes(x=dpi, y=resid))+
  geom_point(color="salmon2")+
   labs(title="High", x="Days Post Infection", y="Residuals", color="Primary Treatment")

summary(glm.h)
h.SD <- 4.522

##Days -8, 14, and 41 
sd <- c(s.SD, l.SD, h.SD) #Standard Deviation from table above
g <- c("Sham", "Low", "High") #Groups

#new df with variance and error for error bars (min/max)
sd.dpi <- data.frame(g, sd)

#plot standard deviation from the models above
ggplot(sd.dpi, aes(x=fct_rev(g), y=sd, color=g))+
  geom_point(size=3)+
  labs(title="Variance during Primary Infection", x="Treatment Groups", y="Model Standard Deviation", color="Treatment Groups")


```

```{r eval=FALSE, include=FALSE}
ab.41 <- p.ab %>%
  filter(dpi == 41)

glm.all <- glmmTMB(elisa_od~1 + (1|band_number), data=ab.41, family=Gamma())
glm.sham <- glmmTMB(elisa_od~1, data=subset(ab.41, primary_treatment == "Sham"), family=Gamma())
glm.low <- glmmTMB(elisa_od~1, data=subset(ab.41, primary_treatment == "Low"), family=Gamma())
glm.high <- glmmTMB(elisa_od~1, data=subset(ab.41, primary_treatment == "High"), family=Gamma())

hist(abs(resid(glm.sham)))
hist(abs(resid(glm.low)))
hist(abs(resid(glm.high)))

summary(glm.sham)
ab.1441$resid <- resid(glm.all)
lm.a <- lm(abs(resid)~primary_treatment, data= ab.1441)
summary(lm.a)
ggplot(ab.1441, aes(x=primary_treatment, y=resid))+
  geom_jitter(width=0.25)


ggplot(ab.1441, aes(x=primary_treatment, y=abs(resid)))+
  geom_jitter(width=0.25)

anova(lm.a)
summary(lm.a)
#for only pid 41
ab.41 <- p.ab %>%
  filter(dpi == 41)

glm41 <- glmmTMB(elisa_od ~ 1, data=ab.41, family=Gamma())
ab.41$resid <- resid(glm41)

lm2 <- lm(resid ~ primary_treatment, data=ab.41)

anova(lm2)
summary(lm2)

ggplot(ab.41, aes(x=primary_treatment, y=resid))+
  geom_point()+
  theme_bw()
```

Histograms
```{r Histograms}
#Add variable indicating the elisa_od mean for each primary_treatment
p.ab <- p.ab %>%
  group_by(primary_treatment)%>%
  mutate(groupmean = mean(elisa_od))

#Add variable indicating the elisa_od mean for each primary_treatment
p.ab <- p.ab %>%
  group_by(primary_treatment, dpi)%>%
  mutate(groupmean.dpi = mean(elisa_od))

#Histogram showing distribution of elisa_od across all sample days by primary_treatment
ggplot(p.ab%>% filter(dpi != -8), aes(x=elisa_od, fill=fct_rev(primary_treatment)))+
  #geom_density(color="black", alpha=0.5)+
  geom_histogram(binwidth = 0.005, position="identity", alpha=2, color="black", size=0.35)+
  geom_vline(xintercept=0.061, linetype="dashed", alpha=0.5)+
  xlim(c(0.04, 0.19))+
  geom_vline(data = p.ab %>% filter(dpi != -8), aes(xintercept = groupmean, color = "red"), linetype = "dotted",alpha = 1, show.legend=FALSE) +
  labs(y= "Count", x= "Antibody Levels", fill="Primary Treatment", color= "Average OD")+
  #scale_fill_manual(values=c( "white", "gray","black"), labels=treat_names)+
  scale_color_manual(values = c("red"), guide = guide_legend(title= NULL))+
  facet_grid(~primary_treatment~dpi)

#Histogram showing mean
ggplot(p.ab%>% filter(dpi != -8), aes(x=elisa_od, fill=fct_rev(primary_treatment)))+
  geom_histogram(binwidth = 0.005, position="identity", alpha=2, color="black", size=0.35)+
  geom_vline(xintercept=0.061, linetype="dashed", alpha=0.5)+
  xlim(c(0.04, 0.19))+
  geom_vline(data = p.ab%>% filter(dpi != -8), aes(xintercept = groupmean.dpi, color = "red"), linetype = "dotted",alpha = 1, show.legend=FALSE) +
  labs(y= "Count", x= "Antibody Levels", fill="Primary Treatment", color= "Average OD")+
  #scale_fill_manual(values=c("black", "gray", "white"), labels=treat_names)+
  scale_color_manual(values = c("red"), guide = guide_legend(title= NULL))+
  facet_grid(~primary_treatment~dpi)
```

Coefficient of Variation: Primary Dose
```{r Coefficient of Variation}
#calculate CV
m.cv.all <- p.ab %>% 
  group_by(primary_treatment)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od), #this is variation within groups 
            bird_sd = sd(elisa_od))

#Calculate CV each day
m.cv <- p.ab %>% 
  group_by(dpi, primary_treatment)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od),
            bird_sd = sd(elisa_od))

#Calculate CV each day
m.cv.s <- p.ab %>% 
  group_by(dpi, primary_treatment, sex)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od),
            bird_sd = sd(elisa_od))

#Graph of CV calculated from dpi -8 to 41 with error bars +/- 1 SD
cv.all.primary <- ggplot(m.cv.all, aes(x=primary_treatment, y=bird_cv, color=fct_rev(primary_treatment)))+
  geom_point(size=3)+
  labs(title="CV of Antibody Levels Across Primary Infection", x="Primary Treatment", y="CV", color="Treatment")
  #scale_fill_manual(values=c( "gray90", "gray50", "gray25"))

cv.all.primary

cv.primary <- ggplot(m.cv, aes(x=primary_treatment, y=bird_cv, color=fct_rev(primary_treatment)))+
  geom_point(size=3)+
  labs(x="Primary Treatment", y="CV", color="Primary Treatment")+
  facet_wrap(~dpi, ncol=3)

cv.primary

ggplot(m.cv, aes(x=dpi, y=bird_cv, color=fct_rev(primary_treatment)))+
  geom_point(size=2)+
  geom_line()+
  labs(x="Primary Treatment", y="CV", color="Primary Treatment")

#by sex
ggplot(m.cv.s, aes(x=dpi, y=bird_cv, color=fct_rev(primary_treatment)))+
  geom_point(size=2)+
  geom_line()+
  labs(x="Primary Treatment", y="CV", color="Primary Treatment")+
  facet_wrap(~sex)

```
Proportional Variability
```{r Proportional Variability}
#install.packages("remotes")
remotes::install_github("T-Engel/CValternatives")
library(CValternatives)
library(patchwork)
p.ab$tes.new <- p.ab$tes + 0.0001
m.cv <- p.ab %>% 
  group_by(dpi, primary_treatment)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od),
            bird_sd = sd(elisa_od),
            bird_pv = PV(elisa_od))

e.cv <- p.ab %>% 
  group_by(dpi, primary_treatment)%>%
  summarise(bird_cv = sd(tes.new)/mean(tes.new),
            bird_sd = sd(tes.new),
            bird_pv = PV(tes.new))

q.cv <- p.ab %>% 
  group_by(dpi, primary_treatment)%>%
  summarise(bird_cv = sd(quantity1)/mean(quantity1),
            bird_sd = sd(quantity1),
            bird_pv = PV(quantity1))
qpv <- ggplot(q.cv, aes(x=dpi, y=bird_pv))+
  geom_point(aes(color=fct_rev(primary_treatment)))
qpv
pv <- ggplot(m.cv, aes(x=dpi, y=bird_pv))+
  geom_point(aes(color=fct_rev(primary_treatment)))

cv <- ggplot(m.cv, aes(x=dpi, y=bird_cv))+
  geom_point(aes(color=fct_rev(primary_treatment)))

ggplot(m.cv, aes(x=bird_pv, y=bird_cv, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_path()
print(pv + cv)
pv
cv
```
```{r}
bird.pv <- p.ab %>% 
  group_by(band_number, primary_treatment)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od),
            bird_sd = sd(elisa_od),
            bird_pv = PV(elisa_od))


```

Does prior exposure induce heterogeneity in immune responses just prior to reinfection?
Prior exposure to different doses of MG induces both higher overall antibody levels just prior to infection (p < 0.0001), and more variable antibody levels.
```{r}

```


----
####Secondary Inoculation####
----
```{r Remove birds that weren't recovered}
#omit from dataset - were still infected (path load) on day 41 prior to reinfection day 42
m.ab <- m.ab %>%
  filter(!(band_number %in% c(2274, 2514, 2469, 2520)))
```
Overview of secondary antibody levels
```{r Antibody levels secondary, warning=FALSE}
#primary w/ secondary colors
ggplot(m.ab %>% filter(dpi <=41), aes(x=dpi, y=elisa_od, color=as.factor(secondary_dose)))+
  geom_point()+
  geom_hline(yintercept = 0.061)+
  stat_smooth(aes(x=dpi, y=elisa_od, group=band_number), alpha=0.5, size=0.5, span=1)+
  scale_color_manual(values = sec_colors)+
  labs(x="Days Post Infection", y="Antibody Levels", color="Secondary Dose")+
  facet_wrap(~primary_treatment)

#ab levels over all experiment
ggplot(m.ab, aes(x=dpi, y=elisa_od, color=as.factor(secondary_dose)))+
  geom_point()+
  geom_hline(yintercept = 0.061)+
  stat_smooth(aes(x=dpi, y=elisa_od, group=band_number), alpha=0.5, size=0.5, span=1)+
  scale_color_manual(values = sec_colors)+
  labs(x="Days Post Infection", y="Antibody Levels", color="Secondary Dose")+
  facet_wrap(~primary_treatment)

#ab levels dpi 56
ggplot(m.ab%>%filter(dpi==56), aes(x=as.factor(secondary_dose), y=elisa_od, color=as.factor(secondary_dose)))+
  geom_hline(yintercept=0.061, linetype="dashed", alpha=0.5)+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(color=as.factor(secondary_dose)), size=1.5, shape=1, alpha=0.75)+
  labs(x="Primary Treatment", y="ELISA OD", fill="Treatment")+
  stat_summary(aes(group=secondary_dose), fun=mean, shape = 18, size=.5)+
  scale_color_manual(values=sec_colors)+
  labs(title = "Antibody Levels Post-Secondary Infection", y= "Antibody Levels", x="Secondary Dose", color="Secondary Dose", shape="Secondary Dose")+
  facet_wrap(~primary_treatment)



```

```{r Format ELISA 41 vs 56 comparison}
m.ab1<-m.ab %>%
  dplyr::select(dpi, quantity, elisa_od, band_number, infected_prim, infected_sec, seropos, seropos_prim, diseased, diseased_prim,
                primary_treatment, secondary_treatment, primary_dose, secondary_dose, sex, tes) #make new df with just relevant variables


elisa41 <- m.ab %>%
  filter(dpi == 41 & band_number != 2375)%>% #dpi 41 for elisa
  dplyr::select(dpi, quantity, elisa_od, band_number, infected_prim, infected_sec, seropos, seropos_prim, diseased, diseased_prim,
                primary_treatment, secondary_treatment, primary_dose, secondary_dose, sex, tes) #df with only elisa od, band number, dpi from 41

quant56 <- m.ab %>%
  filter(dpi == 56 & band_number != 2375)%>% #dpi 56 only for path load; drop 2375 because there was no elisa for it on dpi 41
 dplyr::select(dpi, quantity, elisa_od, band_number, infected_prim, infected_sec, seropos, seropos_prim, diseased, diseased_prim,
                primary_treatment, secondary_treatment, primary_dose, secondary_dose, sex, tes)#df with only quantity, band number, dpi from d56

summary(quant56)
summary(elisa41)

#add column to elisa41 df "from quant56 df row quantity, 
#take all data where the band number from elisa41 match band number from quant56
#returns new column with missing points as NA
elisa41$quantity_dpi_56 <- quant56$quantity[match(elisa41$band_number,quant56$band_number)]
elisa41$tes_dpi_56 <- quant56$tes[match(elisa41$band_number,quant56$band_number)]
elisa41$dpi.check <- quant56$dpi[match(elisa41$band_number,quant56$band_number)] #check band numbers to see if match works
elisa41$dpi.check.no <- quant56$band_number[match(elisa41$band_number,quant56$band_number)] #check band numbers to see if match works
#elisa41$infected.d56 <-quant56$infected[match(elisa41$band_number,quant56$band_number)] #add variable specifying infection d56

elisa41$threshold_cutoff = 50

m.ab <- m.ab %>%
  group_by(band_number) %>%
  mutate(elisa_od_41 = elisa_od[dpi==41])

ggplot(m.ab, aes(x=elisa_od, y=elisa_od_41, color=as.factor(dpi)))+
  geom_point()

elisa41 <- elisa41 %>%
  mutate(infected.d56 = ifelse(quantity_dpi_56>threshold_cutoff, 1, 0))

# Get all possible band_number values
all_band_numbers <- m.ab %>% pull(band_number) %>% unique()


# Create a data frame with all combinations of dpi and band_number
all_combinations <- expand.grid(dpi = unique(elisa41$dpi), band_number = all_band_numbers)


# Merge with the actual data to identify missing band_numbers
missing_band_numbers <- all_combinations %>%
  anti_join(elisa41, by = c("dpi", "band_number"))

# Display the missing band_numbers
print(missing_band_numbers)
#2375 missing - didn't have enough plasma on pid 41
summary(elisa41)


# Get all possible band_number values
all_band_numbers <- m.ab %>% pull(band_number) %>% unique()

# Create a data frame with all combinations of dpi and band_number
all_combinations <- expand.grid(dpi = unique(quant56$dpi), band_number = all_band_numbers)

# Merge with the actual data to identify missing band_numbers
missing_band_numbers56 <- all_combinations %>%
  anti_join(quant56, by = c("dpi", "band_number"))

# Display the missing band_numbers
print(missing_band_numbers56)
#2375 omitted b/c there wasn't plasma for an elisa on 41

pr<-m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(quantity, band_number, elisa_od)
pr
count(pr)
ab <- m.ab %>%
  filter(dpi == 41)%>%
  dplyr::select(elisa_od, band_number, primary_treatment)
ab
ab %>%
  group_by(primary_treatment)

elisa41 %>%
  drop_na(quantity)
```


```{r Models}
#do antibody levels on d41 predict infection on day 56, controlling for secondary inoculation?
lm3 <- glm(infected.d56~elisa_od + secondary_dose, data=elisa41, family="binomial")
summary(lm3)
plot(allEffects(lm3))

#does primary treatment predict infection on day 56, controlling for secondary inoculation?
lm3.5 <- glm(infected.d56~primary_treatment + secondary_dose, data=elisa41, family="binomial")
summary(lm3.5)
plot(allEffects(lm3.5))

#do antibody levels on d14 predict infection on day 56?
lm4 <- glm(infected.d56~elisa_od + primary_treatment * secondary_dose, data=elisa41, family="binomial")
summary(lm4)
plot(allEffects(lm4))

#hist(elisa41$quantity_dpi_56)
#elisa41$log10.quantity_56 <- log10(elisa41$quantity_dpi_56 + 0.1)
#hist(elisa41$log10.quantity_56)
#lm5 <- glmmTMB(log10.quantity_56~secondary_dose + elisa_od, data=elisa41, family= nbinom1)
#summary(lm5)
#infd56 vs ab41
dat.new=expand.grid(secondary_treatment=unique(elisa41$secondary_treatment),
                    secondary_dose=unique(elisa41$secondary_dose),
                    primary_treatment=unique(elisa41$primary_treatment),
                    infected.d56=unique(elisa41$infected.d56),
                    elisa_od=unique(elisa41$elisa_od),
                    band_number=unique(elisa41$band_number))#new grid to put predictions into
dat.new$yhat = predict(lm4, type="response", newdata=dat.new, re.form=NA) #predicted values based off glm.phago model
head(dat.new)

#plot showing predicted values - Antibodies on dpi 41 predict infection probability on dpsi 14
ggplot(elisa41 %>% filter(infected.d56 != "NA"), aes(y=infected.d56, x=elisa_od))+
  geom_point(aes(color=fct_rev(as.factor(infected.d56))), shape=1, size = 2, stroke=1.5)+
  geom_line(data = dat.new, aes(x = elisa_od, y = yhat)) +
  #stat_smooth(method = "glm",
  #            method.args = list(family="binomial"), se=FALSE,
  #            fullrange = TRUE)+
  labs(x="ELISA OD Day 41", y="Infection Status Day 56", color = "Infection Status Day 56")+
  geom_vline(xintercept=0.061, linetype="dashed", alpha=0.5)+
  xlim(c(0.04, 0.13))+
  facet_wrap(~primary_treatment~secondary_dose, ncol=5)

#primary + secondary
lm3.5 <- glm(infected.d56~elisa_od + primary_dose * secondary_dose, data=elisa41, family="binomial")
summary(lm3.5)
plot(allEffects(lm3.5))
plot(resid(lm3.5))
qqnorm(resid(lm3.5))
qqline(resid(lm3.5))

lm3.75 <-glm(infected.d56~elisa_od, data=elisa41, family="binomial")
summary(lm3.75)

dat.new=expand.grid(secondary_treatment=unique(elisa41$secondary_treatment),
                    secondary_dose=unique(elisa41$secondary_dose),
                    primary_treatment=unique(elisa41$primary_treatment),
                    primary_dose=unique(elisa41$primary_dose),
                    infected.d56=unique(elisa41$infected.d56),
                    elisa_od=unique(elisa41$elisa_od),
                    band_number=unique(elisa41$band_number))#new grid to put predictions into
dat.new$yhat = predict(lm3.75, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)

ggplot(elisa41%>% filter(infected.d56 != "NA"), aes(y=infected.d56, x=elisa_od))+
  geom_point((aes(color=fct_rev(as.factor(infected.d56)))), shape=1, size = 3)+
  #geom_point(data = dat.new, aes(x = elisa_od, y = yhat, color=as.factor(secondary_dose))) +
  labs(x="ELISA OD Day 41", y="Infection Status Day 56", color = "Infection Status Day 56")+
  geom_vline(xintercept=0.061, linetype="dashed", alpha=0.3)+
  xlim(c(0.04, 0.13))+
  facet_wrap(~secondary_treatment)

ggplot(elisa41, aes(x=elisa_od, y=infected.d56, color = as.factor(primary_dose)))+
  geom_point()+
#  geom_smooth(data = dat.new, aes(x = elisa_od, y = yhat, color = primary_treatment))+
  facet_wrap(~primary_treatment)
unique(m.ab$primary_treatment)
```
```{r Do Antibodies on day 41 protect from Pathology day 56?, warning=FALSE}
#sum of eye scores secondary infection
m.ab <- m.ab %>%
  group_by(band_number)%>%
  mutate(sec_ses = sum(tes[dpi>41], na.rm =TRUE))
m.ab

ggplot(m.ab, aes(x=tes, y=sec_ses, color=as.factor(dpi)))+
  geom_point()

ggplot(m.ab %>%filter(dpi>41), aes(x=dpi, y=tes, color=as.factor(secondary_dose)))+
  geom_jitter()+
  geom_smooth(aes(groups = band_number), span=0.5, size=0.5, se=FALSE)+
  scale_color_manual(values=sec_colors)+
  facet_wrap(~primary_treatment)

ggplot(m.ab, aes(x=dpi, y=tes, color=fct_rev(primary_treatment)))+
  geom_jitter(height=0)+
  geom_line(aes(groups = band_number))+
  #geom_smooth(aes(groups = band_number), span=0.5, size=0.5, se=FALSE)+
  scale_color_manual(values=pri_colors)+
  labs(y="Total Eye Score", x="Days Post Infection", color= "Secondary Dose")


ggplot(m.ab %>% filter(dpi >41), aes(x=dpi, y=tes, color=as.factor(secondary_dose)))+
  geom_jitter(height=0)+
  geom_smooth(aes(groups = band_number), span=0.5, size=0.5)+
  scale_color_manual(values=sec_colors)+
  labs(y="Total Eye Score", x="Days Post Infection", color= "Secondary Dose")
```


```{r Do Antibodies on day 41 protect from Pathology day 56?, warning=FALSE}
#do antibody levels on day 41 predict sum of eye scores post secondary infection when controlling for secondary dose?
lm5 <- glm(sec_ses ~ elisa_od_41 + primary_treatment*secondary_dose, data=m.ab)
summary(lm5)
plot(allEffects(lm5))
resid <- simulateResiduals(lm5)
plot(resid)

dat.new=expand.grid(sec_ses=unique(m.ab$sec_ses),
                    secondary_dose=unique(m.ab$secondary_dose),
                    primary_treatment=unique(m.ab$primary_treatment),
                    elisa_od_41=unique(m.ab$elisa_od_41),
                    band_number=unique(m.ab$band_number))#new grid to put predictions into
dat.new$yhat = predict(lm5, type="response", newdata=dat.new, re.form=NA) #predicted values
head(dat.new)
unique(dat.new$band_number)

ggplot(m.ab %>% filter(dpi >= 42), aes(y=sec_ses, x=elisa_od_41, color=as.factor(secondary_dose)))+
  geom_point(aes(color=as.factor(secondary_dose)), shape=1, size = 2, stroke=1)+
  geom_line(data = dat.new, aes(x = elisa_od_41, y = yhat, groups=as.factor(secondary_dose))) +
  labs(x="ELISA OD Day 41", y="Sum of Eye Score: Secondary Infection", color = "Secondary Dose")+
  geom_vline(xintercept=0.061, linetype="dashed", alpha=0.5)+
  xlim(c(0.04, 0.13))+
  facet_wrap(~primary_treatment)

ggplot(m.ab %>% filter(dpi >= 42), aes(y=sec_ses, x=elisa_od_41, color=fct_rev(primary_treatment)))+
  geom_point(aes(color=fct_rev(primary_treatment)), shape=1, size = 2, stroke=1)+
  geom_line(data = dat.new, aes(x = elisa_od_41, y = yhat, groups=primary_treatment)) +
  labs(x="ELISA OD Day 41", y="Sum of Eye Score: Secondary Infection", color = "Primiary Treatment")+
  geom_vline(xintercept=0.061, linetype="dashed", alpha=0.5)+
  xlim(c(0.04, 0.13))+
  facet_wrap(~secondary_dose, ncol=5)


```
```{r secondary disease duration}
diseased_duration_sec <- m.ab %>%
  filter(dpi>41)%>%
  group_by(band_number, primary_treatment, sex) %>%
  summarize(diseased_duration = ifelse(sum(diseased == 1) == 0, 0, last(dpi[diseased == 1] + 1) - first(dpi[diseased == 1])))

ggplot(diseased_duration_sec, aes(x=diseased_duration))+
  geom_histogram(aes(fill=fct_rev(primary_treatment)), position="dodge", bins = 20)+
  labs(title = "Duration of Pathology: Secondary Infection", x="Duration of Pathology", y="Count", fill="Primary Treatment")
```


CV Secondary Inoculation
```{r Coefficient of Variation}
#Calculate CV each day including secondary
m.ab$secondary_dose <- as.numeric(m.ab$secondary_dose)

s.cv <- m.ab %>% 
  filter(dpi == 56) %>%
  group_by(dpi, primary_treatment,)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od),
            bird_sd = sd(elisa_od),
            bird_pv = PV(elisa_od))

ggplot(s.cv, aes(x=dpi, y=bird_pv, color=primary_treatment))+ 
  geom_point(size=2)+
  geom_line()+
  scale_color_manual(values=pri_colors)+
  #geom_path(aes(group=primary_treatment))+
  labs(x="Days Post Infection", y= "PV", color= "Primary Treatment")


ggplot(s.cv, aes(x=as.factor(secondary_dose), y=bird_cv, color=as.factor(secondary_dose)))+ 
  geom_point(size=2)+
  scale_color_manual(values=sec_colors)+
  #geom_path(aes(group=primary_treatment))+
  labs(x="Secondary Dose", y= "CV", color= "Secondary Dose")+
  facet_wrap(~primary_treatment)

s.pv <- m.ab %>% 
  group_by(dpi, primary_treatment, secondary_dose)%>%
  summarise(bird_cv = sd(elisa_od)/mean(elisa_od),
           bird_sd = sd(elisa_od),
          bird_pv = PV(elisa_od))

ggplot(s.pv%>%filter(dpi==56), aes(x=as.factor(secondary_dose), y=bird_pv, color=as.factor(secondary_dose)))+ 
  geom_point(size=2)+
  scale_color_manual(values=sec_colors)+
  #geom_path(aes(group=primary_treatment))+
  labs(x="Secondary Dose", y= "PV", color= "Secondary Dose")+
  facet_wrap(~primary_treatment)

m.ab$tes.new <- m.ab$tes + 0.001
g.pv <- m.ab %>% 
  filter(dpsi < 0)%>%
  group_by(dpi, primary_treatment, secondary_dose)%>%
  summarise(bird_cv = sd(tes.new)/mean(tes.new),
           bird_sd = sd(tes.new),
          bird_pv = PV(tes.new))

ggplot(g.pv %>% filter(dpsi < 0), aes(x=dpsi, y=bird_pv, color=primary_treatment))+ 
  geom_point(size=2)+
  #geom_point(s.pv, aes(x=primary_treatment, y=bird_pv))+
  scale_color_manual(values=pri_colors)+
  #geom_path(aes(group=primary_treatment))+
  labs(x="Primary Treatment", y= "PV", color= "Primary Treatment")+
  facet_wrap(~secondary_dose)

m.ab$dpsi <- as.numeric(m.ab$dpsi)
quant.raw <- ggplot(m.ab %>% filter(dpsi %in% c(4, 7, 14, 21)), aes(x=dpsi, y=log10(quantity1), color=fct_rev(primary_treatment)))+
  geom_hline(yintercept = log10(1))+
  geom_point(alpha=0.5)+
  stat_summary(aes(x=dpsi, y=log10(quantity1), groups=primary_treatment), fun=mean, geom="point")+
    stat_summary(aes(x=dpsi, y=log10(quantity1), groups=primary_treatment), fun=mean, geom="line")+
  facet_wrap(~secondary_dose)

tes.raw <- ggplot(m.ab %>% filter(dpsi %in% c(4, 7, 14, 21)), aes(x=dpsi, y=tes, color=fct_rev(primary_treatment)))+
  geom_point( alpha=0.5)+
  stat_summary(aes(x=dpsi, y=tes, groups=primary_treatment), fun=mean, geom="point")+
    stat_summary(aes(x=dpsi, y=tes, groups=primary_treatment), fun=mean, geom="line")+
  facet_wrap(~secondary_dose)

quant.raw
tes.raw
print(quant.raw+tes.raw, nrow=2)

```
Are antibody levels prior to infection protective?
```{r}

```

