---
title: "Host Response"
author: "Jesse Garrett-Larsen"
date: "2024-10-15"
output: html_document
---
EEID 1a Heterogenetiy in Antibodies + Disease Response
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Packages}
rm(list=ls())
library(tidyverse)
library(dplyr)
library(glmmTMB)
library(effects)
library(AICcmodavg)
library(emmeans)
library(MASS)
library(DHARMa)
library(lme4)
library(purrr)
#install.packages("remotes")
#remotes::install_github("T-Engel/CValternatives")
library(CValternatives)
library(patchwork)
library(gridExtra)
library(car)
```

```{r Import Data}
source("dataCleaning_EEID1A.R")
```

```{r Format Theme}
#set colors for primary treatment
pri_colors <- c("#1B9E77", "#7570B3", "#D95F02")
#set color scheme secondary treatment
sec_colors <- c("#8C754B", "#77AB59", "#59A5D8", "#9F77D9", "#FA8072")
#set theme
theme_set(theme_minimal())
```

```{r Raw Data Sample Sizes}
m.ab %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  )%>%
  modify_header(
    label ~ "**All Birds Before Removal**"
  )
```

####Generate Infection Data + Format For Analysis
Rules: A bird is considered infected if it had an eye score > 0.5 for more than one day, a pathogen load > 50 copies, or both
```{r Threshold Cutoffs}
m.ab$threshold_cutoff = 50
m.ab$seropos_cutoff = 0.061
m.ab$pathology_cutoff = 0
```

inf = if quantity (where any NA is replaced with 0) is greater than cutoff of 50 copies OR total eye score (where any NAs are ignored) is greater than 0.5 for more than one consecutive sampling day, or total eye score is greater or equal to 1 for at least one day, return 1, else return 0
```{r Generate Infection Data}
m.ab <- m.ab %>%
  # Sort data by band_number and dpi
  arrange(band_number, dpi) %>%
  # Replace NA values for quantity only
  mutate(quantity_clean = coalesce(quantity, 0)) %>%
  # Flag as infected if quantity > 50 on any single day
  mutate(quantity_flag = ifelse(quantity_clean > 50, 1, 0)) %>%
  # Check consecutive days for tes > 0.5, ignoring NAs for tes
  group_by(band_number) %>%
  mutate(
    tes_flag = ifelse(tes > 0.5 & !is.na(tes), 1, 0),
    tes_consecutive = tes_flag + lag(tes_flag, default = 0),
    tes_single_high = ifelse(tes >= 1 & !is.na(tes), 1, 0)
  ) %>%
  # Combine the three criteria: quantity > 50 OR tes > 0.5 on consecutive days OR tes >= 1 on any day
  mutate(inf = ifelse(quantity_flag == 1 | tes_consecutive > 1 | tes_single_high == 1, 1, 0)) %>%
  # Calculate inf_pri and inf_sec based on dpi
  ungroup() %>%
  group_by(band_number) %>%
  mutate(
    inf_pri = ifelse(any(inf == 1 & dpi < 42), 1, 0),
    inf_sec = ifelse(any(inf == 1 & dpi > 42), 1, 0)
  ) %>%
  # Un-group the data if needed
  ungroup() %>%
  # Select relevant columns (optional)
  select(-quantity_clean, -quantity_flag, -tes_flag, -tes_consecutive, -tes_single_high)


ggplot(m.ab, aes(x=dpi, y=tes, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_line(aes(groups=band_number))+
  geom_hline(yintercept=log10(50))+
facet_wrap(~inf)

ggplot(m.ab, aes(x=dpi, y=tes, color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_line(aes(groups=band_number))+
  geom_hline(yintercept=log10(50))+
facet_wrap(~inf_pri)

ggplot(m.ab, aes(x=dpi, y=log10(quantity1), color=fct_rev(primary_treatment)))+
  geom_point()+
  geom_line(aes(groups=band_number))+
  geom_hline(yintercept=log10(50))+
facet_wrap(~inf_sec)

```

```{r Omit Birds n=10}
prob_birds <- m.ab %>% filter(band_number %in% c(2274, 2514, 2469, 2520, 2494, 2452, 2562, 2419, 2434, 2505))%>%
  select(band_number, dpi, primary_treatment, secondary_dose, quantity, tes, l_eye_score, r_eye_score, elisa_od)

prob_birds %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(by=dpi)%>%
  modify_header(
    label ~ "**Birds to Remove**"
  )

#omit 2505 from data set for analysis as it was positive at quarantine (n=1)
m.ab <- m.ab %>%
  filter(band_number != 2505)

#omit 2452 and 2562 bc positive qpcr and they are shams (n=2)
m.ab <- m.ab %>% filter(!band_number %in% c(2452, 2562)) 

#omit sham birds with eye scores (0.5) (2419, 2434) (n=2)
m.ab <- m.ab %>% filter(!band_number %in% c(2419, 2434))

#omit sham bird with path load (2514) (n=1)
m.ab <- m.ab %>% filter(band_number != 2514)

m.ab %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  ) %>%
  modify_header(
    label ~ "**Sample Sizes for Primary Analysis**"
  )
#Final Primary Sample Sizes: Sham = 46; Low = 51; High = 53
```
```{r Idenity Birds to Remove For Secondary}
##Some birds did not fully recover after primary infection. These birds were included in primary analysis, but will be excluded from secondary analysis. I have identified these birds below, but *will not remove them until secondary analysis*.

#check birds that were not recovered by secondary infection: 2274, 2469, 2520, 2494 (2514 not recovered but omitted above)
not_recovered <- m.ab %>% 
  filter(band_number %in% c(2274, 2469, 2520, 2494))%>%
  dplyr::select(band_number, dpi, primary_treatment, secondary_dose, quantity, tes)

not_recovered %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  )%>%
  modify_header(
    label ~ "**Birds Not Recovered DPI 41**"
  )

#confirm quantity > cutoff at dpi 41
ggplot(not_recovered, aes(x=dpi, y=quantity, color=primary_treatment))+
  geom_point()+
  geom_path()+
  geom_hline(yintercept = 50, alpha=0.5)+
  geom_vline(xintercept = 41, alpha=0.5, lty = "dashed")+
  facet_wrap(~band_number)+
  ylim(min=0, max=700)
```

####Primary Challenge
```{r Final Primary Sample Size}
m.ab %>%
  dplyr::select(dpi, primary_treatment)%>%
  tbl_summary(
    by=dpi
  ) %>%
  modify_header(
    label ~ "**PRIMARY ANALYSIS FINAL SAMPLE SIZES**"
  )
```
```{r p.ab Data Frame With Only Primary Infection}
#data frame with only primary infection
p.ab <- m.ab %>%
  filter(dpi <=41)
```

```{r Do baseline antibody levels differ between treatment groups?}
#Baseline antibody levels did not differ
lm0 <- glmmTMB(elisa_od~primary_treatment, data=p.ab %>% filter(dpi <0), family=Gamma())
summary(lm0)
plot(allEffects(lm0))
simulateResiduals(lm0, plot=T)

emm_r <- emmeans(lm0, ~primary_treatment)
pairs(emm_r)
```
```{r Neither mass nor sex predict antibody response}
p.ab$mass <- as.numeric(p.ab$mass)
max.ab <- p.ab %>%
  group_by(band_number, primary_treatment) %>%
  reframe(max_od = max(elisa_od, na.rm = TRUE),
            mass = as.numeric(mass[dpi == -8]),
            dpi = dpi,
            sex = sex,
            inf_pri = inf_pri)

#Neither Mass nor Sex predict max antibody response
lmm <- glm(max_od ~ mass + primary_treatment + sex, data=max.ab %>%filter(dpi == -8), family=Gamma())
summary(lmm)
plot(allEffects(lmm))
simulateResiduals(lmm, plot=T)

ggplot(max.ab%>%filter(dpi == -8), aes(x=mass, y=max_od, color=primary_treatment))+
  geom_boxplot(aes(x=mass, y=0.1), width=0.01, color="black")+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family="Gamma"))+
  facet_wrap(~sex~primary_treatment, nrow=2)

#Neither Mass nor Sex predict infection status
lmi <- glm(inf_pri ~ mass + primary_treatment + sex, data=max.ab %>%filter(dpi == -8), family=binomial())
summary(lmi)
simulateResiduals(lmi, plot=T)

```

####Antibody Levels Across Primary Treatment
```{r Antibody Dataframe}
#I treat days post inoculation as a factor in these models and use band_number as a random effect
p.abt <- p.ab %>%
  filter(dpi %in% c(-8, 14, 41))
p.abt$dpi.f <- as.factor(p.abt$dpi)
```

```{r Do antibody levels vary across dpi? Does sex influence antibody response?}
#model comparison 
p1 <- glmmTMB(elisa_od~primary_treatment + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p2 <- glmmTMB(elisa_od~primary_treatment*dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p2.5 <- glmmTMB(elisa_od~primary_treatment*dpi.f + sex + (1|band_number), data=p.abt, family=Gamma(log))
p3<- glmmTMB(elisa_od~primary_treatment + sex + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p4<- glmmTMB(elisa_od~primary_treatment * sex + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))
p5 <- glmmTMB(elisa_od~1 + dpi.f + (1|band_number), data=p.abt, family=Gamma(log))

aictab(cand.set=list(p1, p2, p2.5, p3, p4, p5), modnames=c("p1", "p2", "p2.5", "p3", "p4", "p5"))
# Interaction between primary_treatment and dpi.f without sex is best model by AICc

#model comparison VGAM
library(VGAM)
drop1(p1, test="Chisq") #primary_treatment and dpi.f contribute uniquely to explaining variation

drop1(p2, test="Chisq") #The interaction between primary_treatment and dpi.f > the effect of primary treatment on the outocme varies significantly by dpi levels

drop1(p2.5, test="Chisq") # sex is not significant

drop1(p3, test="Chisq") # sex is not significant but interaction b/t primary_treatment and dpi.f is better

drop1(p4, test="Chisq") # primary_treatment*sex is not significant

drop1(p5, test="Chisq") # better with primary_treatment
```
```{r Primary Treatment Antibody Final Model}
#make Sham the reference category
p.abt$primary_treatment <- relevel(p.abt$primary_treatment, ref = "Sham")
#final model
library(lme4)
lm1 <- glmmTMB(elisa_od~ primary_treatment*dpi.f + (1|band_number), data=p.abt, family=Gamma(log)) 

#fully interactive does not converge with inverse link- only log link
summary(lm1)
car::Anova(lm1, type = "III")
simulateResiduals(lm1, plot=T) #but residuals look terrible with log link
hist(resid(lm1))
simulation_output <- simulateResiduals(fittedModel = lm1, n = 1000)

# Plot residuals
plotSimulatedResiduals(simulation_output, rank = TRUE)
#dispersion test
testDispersion(simulation_output)

p.abt$elisa_od_log <- log(p.abt$elisa_od)
hist(p.abt$elisa_od_log)
lm1log <- lmer(elisa_od_log ~ primary_treatment * dpi.f + (1|band_number), data=p.abt)
summary(lm1log)

emm_results <- emmeans(lm1, ~ primary_treatment|dpi.f, scale="response")
pairwise <- pairs(emm_results, adjust="tukey")
(summary(pairwise))
pairwise_df <- as.data.frame(pairwise)

ggplot(pairwise_df, aes(x = contrast, y = estimate, shape=dpi.f)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Pairwise Comparisons of Primary Treatments within Each dpi.f",
       x = "Comparison",
       y = "Estimated Difference") +
    coord_flip() +  # Flip the coordinates for better readability
  theme_minimal()
```

```{r Model Predictions lm1}
#specify model to predict
mod.pred <- lm1
#Model predictions
dat.new=expand.grid(primary_treatment=unique(p.abt$primary_treatment),
                    elisa_od = unique(p.abt$elisa_od),
                    dpi.f = unique(p.abt$dpi.f))
#generate yhat values
dat.new$yhat = predict(mod.pred, type="response", newdata=dat.new, re.form=NA)
#generate SEM
preds = predict(mod.pred, type="link", newdata=dat.new, se.fit=TRUE, re.form=NA)
dat.new = cbind(dat.new, preds)

ilink <- family(mod.pred)$linkinv
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))
head(dat.new)
```

```{r Format data for Variability in Antibodies for individuals who were infected}
head(p.ab)
#add small constant to total eye score for calculation of variability
p.ab$tes.new <- p.ab$tes + .01
#new df with only individuals that were infected any time during primary challenge
p.ab$dpi.f <- as.factor(p.ab$dpi)
p.abi <- p.ab %>%
  filter(inf_pri == 1)%>%
  dplyr::select(dpi.f, dpi, primary_treatment, tes.new, elisa_od, quantity1, sex, band_number, inf_pri)

#df with all individuals during primary challenge
p.aba <-  p.ab %>%
  dplyr::select(dpi.f, dpi, primary_treatment, tes.new, elisa_od, quantity1, sex, band_number, inf_pri)
```


```{r Calculate Variability in Antibodies for individuals who were infected}
# Define functions for calculating CV, PV, and V2
calculate_cv <- function(data) {
  return(sd(data) / mean(data))
}

calculate_pv <- function(data) {
  return(PV(data))
}

calculate_v2 <- function(data) {
  mean_x <- mean(data)
  sd_x <- sd(data)
  v2 <- sd_x^2 / (sd_x^2 + mean_x^2)
  return(v2)
}

# Set the number of bootstrap replicates
n_boot <- 1000

# Pipeline to calculate CV, PV, V2, and their 95% confidence intervals
a.cv <- p.aba %>%
  group_by(dpi.f, primary_treatment) %>%
  drop_na(elisa_od) %>%
  dplyr::reframe(
    # Calculate means and standard deviations
    elisa_od = elisa_od,
    mean_od = mean(elisa_od),
    bird_sd = sd(elisa_od),
    band_number = band_number,
    bird_cv = calculate_cv(elisa_od),   # Calculate CV for the original data
    bird_pv = calculate_pv(elisa_od),   # Calculate PV for the original data
    bird_v2 = calculate_v2(elisa_od),   # Calculate V2 for the original data
    
    # Bootstrap for each metric
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(elisa_od, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(elisa_od, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(elisa_od, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for CV, PV, and V2
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975)),
    
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  select(-cv_bootstrap, -pv_bootstrap, -v2_bootstrap) %>%
  ungroup()

#add to df and format names
p.aba.m <- left_join(p.aba, a.cv, by=c("dpi.f", "primary_treatment", "band_number"))
 p.aba.m <- p.aba.m %>%
   select(-elisa_od.y)
 p.aba.m$elisa_od <- p.aba.m$elisa_od.x
# p.abi.m$primary_treatment <- p.abi.m$primary_treatment.x
# p.abi.m <- p.abi.m %>%
#   select(-elisa_od.x, -primary_treatment.x)

#just days samples were taken
p.aba.m <- p.aba.m %>%
  filter(dpi.f %in% c(-8, 14, 41))

summary_tibble <- a.cv %>%
  group_by(dpi.f, primary_treatment) %>%
  summarize(
    mean_bird_cv = mean(bird_cv, na.rm = TRUE),    # Calculate the mean of bird_cv for each group
    mean_bird_sd = mean(bird_sd, na.rm = TRUE),    # Calculate the mean of bird_sd for each group
    mean_bird_pv = mean(bird_pv, na.rm = TRUE),    # Calculate the mean of bird_pv for each group
    avg_elisa_od = mean(elisa_od, na.rm = TRUE),# Calculate the average elisa_od for each group
    n_individuals = n_distinct(band_number) #calculate number of individuals in each group
  ) %>%
  as_tibble()
print(summary_tibble)
```
```{r Antibodies and Variability Figure 2}
#overlay PV and elisa_od
#FIGURE 2
ggplot(data = p.aba.m, aes(x = dpi.f, y = elisa_od, color = primary_treatment)) +
  # Original jittered data points for elisa_od
  geom_jitter(size = 1.5, alpha = 1, width = 0.15, height = 0, shape=1) +
  
  # Line and points with error bars for predicted values
  geom_path(data = dat.new, aes(x = dpi.f, y = yhat, group = primary_treatment,
                                color = primary_treatment), alpha=0.75) +
  geom_errorbar(data = dat.new, aes(ymin = Lower, ymax = Upper, x = dpi.f, y = yhat),
                color = "black", width = 0.1) +
  geom_point(data = dat.new, aes(x = dpi.f, y = yhat, color = primary_treatment),
             size = 2, alpha = 0.75, shape = 16) +
  geom_point(data = dat.new, aes(x = dpi.f, y = yhat),
             color = "black", size = 2, alpha = 1, shape = 1, stroke = 0.1) +
  
  # V2 (CV^2) points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, fill = primary_treatment, shape = "PV"),
             size = 3, alpha = 0.2) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci,
                                 group = primary_treatment, color = primary_treatment),
                width = 0.05, alpha = 0.02) +
  geom_path(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment,
                             group = primary_treatment), lty = "dashed", alpha=0.2) +

  # Labels and axis
  labs(y = "Anti-MG IgY Antibody Levels [OD]", 
       x = "Days Post Primary Inoculation", 
       shape = "Variability Metric", 
       color = "Primary Treatment", 
       fill = "Primary Treatment") +
  
  # Custom color and fill scales
  scale_color_manual(values = c(pri_colors)) +
  scale_fill_manual(values = c(pri_colors)) +

  # Custom shape scale for V2 (CV^2)
  scale_shape_manual(
    name = "Variability Metric",
    values = c("PV" = 17),  # Set shape type for CV^2
    labels = c("PV" = "PV")
  ) +
  
  # Theme settings
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "black", size = 15, face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold", size = 15),
    axis.text.y.right = element_text(color = "black", face = "bold", size = 15),
    axis.title.y.right = element_text(color = "black", size = 20, face = "bold"),
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    legend.position = "top",
    legend.direction = "horizontal"
  )
```


```{r Antibodies and Variability Figure 2}
#Just variability
ggplot() +
  # CV points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_cv, color = primary_treatment, shape = "CV"), 
             size = 3, position = position_dodge(width = 0.6), alpha = 0.5) +
  geom_path(data = a.cv, aes(x = dpi.f, y = bird_cv, color = primary_treatment), 
            position = position_dodge(width = 0.6)) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = cv_lower_ci, ymax = cv_upper_ci,
                                 color = primary_treatment, shape = "CV",), lty="dashed",
                width = 0.15, position = position_dodge(width = 0.6), alpha=0.05) +
  
  # PV points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment, shape = "PV"), 
             size = 3, position = position_dodge(width = 0.3), alpha = 0.5) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, 
                                 color = primary_treatment, shape = "PV"), lty="dashed",
                width = 0.15, position = position_dodge(width = 0.3), alpha=0.05) +
  
  # V2 points and confidence intervals
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_v2, color = primary_treatment, shape = "V2"), 
             size = 3, position = position_dodge(width = 0.9), alpha = 1) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, 
                                 color = primary_treatment, shape = "V2"), 
                width = 0.15, position = position_dodge(width = 0.9)) +
  
  # Custom color and fill scales
  scale_color_manual(values = c(pri_colors)) +
  scale_fill_manual(values = c(pri_colors)) +
  
  # Shape legend to distinguish metrics
  scale_shape_manual(
    name = "Variability Metric",
    values = c("V2" = 16, "CV" = 0, "PV" = 2),
    labels = c("V2" = "CV^2", "CV" = "CV", "PV" = "PV")
  ) +
  
  # Labels and theme
  labs(y = "Variability Metric Value", 
       x = "Days Post Primary Inoculation", 
       color = "Primary Treatment") +
  theme_minimal() +
  theme(
    axis.title.y = element_text(color = "black", size = 20, face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold", size = 15),
    axis.text.y.right = element_text(color = "black", face = "bold", size = 15),
    axis.title.y.right = element_text(color = "black", size = 20, face = "bold"),
    axis.title.x = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    legend.position = "top",
    legend.direction = "horizontal"
  )#+
  coord_flip()

```

```{r CV for Gamma Distributed Antibodies}
library(MASS)

fit_ab <- p.aba %>%
  filter(dpi.f %in% c(-8, 14, 41))%>%
  group_by(dpi.f, primary_treatment) %>%
  summarize(
    shape = fitdistr(na.omit(elisa_od), "gamma")$estimate["shape"],
    rate = fitdistr(na.omit(elisa_od), "gamma")$estimate["rate"]
  ) %>%
  ungroup()

fit_ab <- fit_ab %>%
  mutate(cv = 1 / sqrt(shape))

ggplot(fit_ab, aes(x=dpi.f, y=cv, color=primary_treatment))+
  geom_point()+
  scale_color_manual(values=c(pri_colors))


# Function to calculate CV for gamma distribution


library(MASS)
library(dplyr)
library(purrr)

# Function to calculate gamma CV
calculate_gamma_cv <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return(1 / sqrt(shape))
    }
  }
  return(NA)
}

# Function to calculate gamma V2
calculate_gamma_V2 <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return((1 + shape)^(-1/2))
    }
  }
  return(NA)
}

# Set the number of bootstrap replicates
n_boot <- 1000

# Pipeline to calculate gamma CV, gamma V2, and 95% CI
a.gcv <- p.aba %>%
  group_by(dpi.f, primary_treatment) %>%
  drop_na(elisa_od) %>%
  dplyr::reframe(
    # Original data calculations
    elisa_od = elisa_od,
    mean_od = mean(elisa_od, na.rm = TRUE),
    bird_sd = sd(elisa_od, na.rm = TRUE),
    band_number = band_number,
    
    bird_gamma_cv = calculate_gamma_cv(elisa_od),   # Gamma CV for original data
    bird_gamma_V2 = calculate_gamma_V2(elisa_od),   # Gamma V2 for original data

    # Bootstrap for gamma CV and V2 separately
    gamma_cv_bootstrap = list(replicate(n_boot, calculate_gamma_cv(sample(elisa_od, replace = TRUE)))),
    gamma_V2_bootstrap = list(replicate(n_boot, calculate_gamma_V2(sample(elisa_od, replace = TRUE))))
  ) %>%
  # Extract gamma CV and V2 bootstrap estimates and calculate 95% confidence intervals
  mutate(
    gamma_cv_lower_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_cv_upper_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),

    gamma_V2_lower_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_V2_upper_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  select(-gamma_cv_bootstrap, -gamma_V2_bootstrap) %>%
  ungroup()

# Add to the main dataset and calculate group summaries
p.aba.g <- left_join(p.aba, a.gcv, by = c("dpi.f", "primary_treatment", "band_number")) %>%
  select(-elisa_od.y) %>%
  mutate(elisa_od = elisa_od.x) %>%
  select(-elisa_od.x) %>%
  filter(dpi.f %in% c(-8, 14, 41))

summary_tibble_gamma <- a.gcv %>%
  group_by(dpi.f, primary_treatment) %>%
  summarize(
    mean_bird_gamma_cv = mean(bird_gamma_cv, na.rm = TRUE),  # Mean gamma CV for each group
    mean_bird_gamma_V2 = mean(bird_gamma_V2, na.rm = TRUE),  # Mean V2 for each group
    mean_bird_sd = mean(bird_sd, na.rm = TRUE),              # Mean of bird_sd
    avg_elisa_od = mean(elisa_od, na.rm = TRUE),             # Mean elisa_od for each group
    n_individuals = n_distinct(band_number)                  # Number of individuals in each group
  ) %>%
  as_tibble()

print(summary_tibble_gamma)



ggplot() +
  # Plotting mean gamma CV with error bars
  geom_point(data = summary_tibble_gamma, aes(x = dpi.f, y = mean_bird_gamma_cv, color = primary_treatment, shape = "Gamma CV"), size = 3, position = position_dodge(width = 0.25)) +
  geom_errorbar(data = a.gcv, aes(x = dpi.f, y = bird_gamma_cv, ymin = gamma_cv_lower_ci, ymax = gamma_cv_upper_ci, color = primary_treatment), linetype="dotted", position = position_dodge(width = 0.25), width = 0.2) +

  # Plotting mean gamma V2 with error bars
  geom_point(data = summary_tibble_gamma, aes(x = dpi.f, y = mean_bird_gamma_V2, color = primary_treatment, shape = "Gamma V2"), size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = a.gcv, aes(x = dpi.f, y = bird_gamma_V2, ymin = gamma_V2_lower_ci, ymax = gamma_V2_upper_ci, color = primary_treatment), position = position_dodge(width = 0.5), width = 0.2, linetype = "dashed") +

  # Plotting mean PV with error bars
  geom_point(data = a.cv, aes(x = dpi.f, y = bird_pv, color = primary_treatment, shape = "PV"), size = 3, position = position_dodge(width = 0.75)) +
  geom_errorbar(data = a.cv, aes(x = dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, color = primary_treatment), linetype = "solid", width = 0.2, alpha = 0.05, position = position_dodge(width = 0.75)) +

  # Custom colors
  scale_color_manual(values = pri_colors) +
  
  # Define shapes for each metric in the legend
  scale_shape_manual(values = c("Gamma CV" = 21, "Gamma V2" = 22, "PV" = 16)) +

  # Labels
  labs(y = "Gamma CV / Gamma V2 / PV", x = "DPI", title = "Gamma CV, Gamma V2, and PV of Antibodies by DPI and Treatment", shape = "Variability Metric") +
  
  # Legends and themes
  theme_minimal() 
```


```{r Calculate Variability in eye score}
#eye score
e.cv <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri == 1)%>%
  reframe(
    bird_cv = calculate_cv(tes.new),
    bird_sd = sd(tes.new),
    bird_pv = calculate_pv(tes.new),
    bird_v2 = calculate_v2(tes.new),
    dpi.f = dpi.f,
    tes.new = tes.new,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(tes.new, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(tes.new, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(tes.new, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Remove bootstrap columns
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()
```

```{r Visualize Variability in Eye Score}
ggplot(e.cv)+
  #geom_jitter(aes(x=dpi.f, y=tes.new, color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV"), size=2)+
  geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), size=2, position=position_dodge(width=0.5))+
  # geom_point(aes(x=dpi.f, y=bird_cv, color=primary_treatment, shape="CV"), size=2, position=position_dodge(width=0.75))+
  #  geom_errorbar(aes(x=dpi.f, ymin = cv_lower_ci, ymax = cv_upper_ci, y=bird_cv, color=primary_treatment), width=0.2, lty="dotted", position=position_dodge(width=0.75))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed")+
  geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.5) , width=0.2)+
 scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 15, "PV" = 17, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  scale_color_manual(values=(pri_colors[c(2, 3)]))+
  labs(x="Days Post Inoculation", y="Variability", shape="Variability Metric", color="Primary Treatment")

ggplot(e.cv)+
  geom_jitter(aes(x=dpi.f, y=tes.new, color=primary_treatment), height=0, width=0.1, alpha=0.5)+
  scale_color_manual(values=(pri_colors[c(2, 3)]))+
  facet_wrap(~primary_treatment)
```
Strongly depends on what measure of variability I use - use PV for eye score becuase of zero inflation
--------
I use raw quantity data because I want to capture heterogeneity as it applies to transmission-relevant traits. I don't want to reduce the impact of "outliers" as those data represent individuals that may play an outsized role in seeding/propagating epidemics.
```{r Calculate Raw Path Load Variability in infected birds primary infection}
#path load
q.cv <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri ==1)%>%
  drop_na(quantity1) %>%
  reframe(
    bird_cv = calculate_cv(quantity1),
    bird_pv = calculate_pv(quantity1),
    bird_v2 = calculate_v2(quantity1),
    dpi.f = dpi.f,
    quantity1 = quantity1,
    inf_pri = inf_pri,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(quantity1, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(quantity1, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(quantity1, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975))
  ) %>%
  # Remove bootstrap columns
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()
```

```{r With Gamma Distributed variability metrics}
# Functions for gamma CV and V2 based on gamma distribution parameters
calculate_gamma_cv <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return(1 / sqrt(shape))
    }
  }
  return(NA)
}

calculate_gamma_V2 <- function(data) {
  data <- na.omit(data)
  if (length(data) > 1) {
    fit <- tryCatch(
      fitdistr(data, "gamma"),
      error = function(e) return(NA)
    )
    if (!is.na(fit[1])) {
      shape <- fit$estimate["shape"]
      return((1 + shape)^(-1/2))
    }
  }
  return(NA)
}

# Set number of bootstrap replicates
n_boot <- 1000

# Main data processing pipeline
q.cv <- p.aba %>%
  group_by(dpi, primary_treatment) %>%
  filter(inf_pri == 1 & dpi.f %in% c(-8, 7, 41)) %>%
  drop_na(quantity1) %>%
  dplyr::reframe(
    # Original data metrics
    bird_cv = calculate_cv(quantity1),
    bird_pv = calculate_pv(quantity1),
    bird_v2 = calculate_v2(quantity1),
    bird_gamma_cv = calculate_gamma_cv(quantity1),
    bird_gamma_V2 = calculate_gamma_V2(quantity1),
    
    dpi.f = dpi.f,
    quantity1 = quantity1,
    inf_pri = inf_pri,
    
    # Bootstrapping for CV, PV, V2, and gamma metrics
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(quantity1, replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(quantity1, replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(quantity1, replace = TRUE)))),
    gamma_cv_bootstrap = list(replicate(n_boot, calculate_gamma_cv(sample(quantity1, replace = TRUE)))),
    gamma_V2_bootstrap = list(replicate(n_boot, calculate_gamma_V2(sample(quantity1, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for all metrics
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    
    gamma_cv_lower_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_cv_upper_ci = map_dbl(gamma_cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    gamma_V2_lower_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    gamma_V2_upper_ci = map_dbl(gamma_V2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Remove bootstrap columns to clean up the dataset
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap, -gamma_cv_bootstrap, -gamma_V2_bootstrap) %>%
  ungroup()

# View the resulting data
print(q.cv)

# Create summary table
summary_q_cv <- q.cv %>%
  group_by(dpi, primary_treatment) %>%
  summarize(
    # Mean values for each metric
    mean_bird_cv = mean(bird_cv, na.rm = TRUE),
    mean_bird_pv = mean(bird_pv, na.rm = TRUE),
    mean_bird_v2 = mean(bird_v2, na.rm = TRUE),
    mean_bird_gamma_cv = mean(bird_gamma_cv, na.rm = TRUE),
    mean_bird_gamma_V2 = mean(bird_gamma_V2, na.rm = TRUE),
    
    # 95% Confidence Intervals for each metric
    cv_lower_ci = mean(cv_lower_ci, na.rm = TRUE),
    cv_upper_ci = mean(cv_upper_ci, na.rm = TRUE),
    
    v2_lower_ci = mean(v2_lower_ci, na.rm = TRUE),
    v2_upper_ci = mean(v2_upper_ci, na.rm = TRUE),
    
    pv_lower_ci = mean(pv_lower_ci, na.rm = TRUE),
    pv_upper_ci = mean(pv_upper_ci, na.rm = TRUE),
    
    gamma_cv_lower_ci = mean(gamma_cv_lower_ci, na.rm = TRUE),
    gamma_cv_upper_ci = mean(gamma_cv_upper_ci, na.rm = TRUE),
    
    gamma_V2_lower_ci = mean(gamma_V2_lower_ci, na.rm = TRUE),
    gamma_V2_upper_ci = mean(gamma_V2_upper_ci, na.rm = TRUE)
  ) %>%
  ungroup()

# View the summary table
print(summary_q_cv)
```


Low dose birds are more variable in their pathogen loads during primary infection than high dose birds.
```{r Visualize in infected birds primary infection}
ggplot(q.cv)+
  #geom_jitter(aes(x=dpi.f, y=quantity1, color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV"), size=3, position=position_dodge(width=0.5))+
  geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), size=3, position=position_dodge(width=0.5))+
  #geom_point(aes(x=dpi.f, y=bird_cv, color=primary_treatment, shape="CV"), size=2, position=position_dodge(width=0.75))+
  #geom_errorbar(aes(x=dpi.f, ymin = cv_lower_ci, ymax = cv_upper_ci, y=bird_cv, color=primary_treatment), width=0.2, lty="dotted", position=position_dodge(width=0.75))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed", position=position_dodge(width=0.5))+
  geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.5), width=0.2)+
 scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 15, "PV" = 17, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  scale_color_manual(values=c(pri_colors[c(2, 3)]))+
  labs(title="Path Load Variability", y="Variability Metric", x="DPI", lty="Variability Metric", color="Primary Treatment")

ggplot(q.cv)+
  geom_jitter(aes(x=dpi.f, y=quantity1, color=primary_treatment), height=0, width=0.25, alpha=0.5)+
  geom_hline(yintercept = 50, alpha=0.25)+
  scale_color_manual(values=c(pri_colors[c(2, 3)]))
```


```{r Log10 Quantity Variability}
p.aba$quantity1 <- p.aba$quantity1+0.0001
q.cv.log <- p.aba %>% 
  group_by(dpi, primary_treatment) %>%
  drop_na(quantity1) %>%
  reframe(
    bird_cv = calculate_cv(log10(quantity1)),
    bird_pv = calculate_pv(log10(quantity1)),
    bird_v2 = calculate_v2(log10(quantity1)),
    dpi.f = dpi.f,
    quantity1 = quantity1,
    
    # Bootstrap for confidence intervals
    cv_bootstrap = list(replicate(n_boot, calculate_cv(sample(log10(quantity1), replace = TRUE)))),
    v2_bootstrap = list(replicate(n_boot, calculate_v2(sample(log10(quantity1), replace = TRUE)))),
    pv_bootstrap = list(replicate(n_boot, calculate_pv(sample(log10(quantity1), replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals
  mutate(
    cv_lower_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci = map_dbl(cv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci = map_dbl(v2_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci = map_dbl(pv_bootstrap, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Remove bootstrap columns
  select(-cv_bootstrap, -v2_bootstrap, -pv_bootstrap) %>%
  ungroup()
```

Variability here is funky because there are values below 50 (infection cutoff) that are affecting variability metrics.
```{r Log Quantity Variability}
hist(log10(q.cv.log$quantity1))
ggplot(q.cv.log %>% filter(primary_treatment != "Sham"))+
  #geom_jitter(aes(x=dpi.f, y=log10(quantity1), color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV"), size=3)+
  geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), size=3, position=position_dodge(width=0.2))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed")+
  geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.2) , width=0.2)+
  scale_color_manual(values=c(pri_colors[c(2,3)]))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("PV" = 17, "V2" = 18),  # Assign shapes
                     labels = c("PV" = "PV", "V2" = "CV^2")) +
  labs(title="Log10(Quantity) Variability", y="Variability", x="DPI")

ggplot(q.cv.log)+
  geom_jitter(aes(x=dpi.f, y=log10(quantity1), color=primary_treatment), height=0, width=0.25, alpha=0.5)+
  scale_color_manual(values = c(pri_colors))+
  geom_hline(yintercept = log10(50.001))

ggplot(q.cv.log %>% filter(primary_treatment != "Sham"))+
  #geom_jitter(aes(x=dpi.f, y=log10(quantity1), color=primary_treatment))+
  geom_point(aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV Log"), position = position_dodge(width=0.3))+
  #geom_point(aes(x=dpi.f, y=bird_v2, color=primary_treatment, shape="V2"), position=position_dodge(width=0.2))+
  geom_errorbar(aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), position = position_dodge(width=0.3), width=0.1, lty="solid")+
  # geom_errorbar(aes(x=dpi.f, ymin = v2_lower_ci, ymax = v2_upper_ci, y=bird_v2, color=primary_treatment), position=position_dodge(width=0.2) , width=0.2)+
   geom_point(data=q.cv, aes(x=dpi.f, y=bird_pv, color=primary_treatment, shape="PV Raw"), size=3, position=position_dodge(width=0.5))+
  geom_errorbar(data=q.cv, aes(x=dpi.f, ymin = pv_lower_ci, ymax = pv_upper_ci, y=bird_pv, color=primary_treatment), width=0.2, lty="dashed", position=position_dodge(width=0.5))+
  scale_color_manual(values=c(pri_colors[c(2,3)]))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("PV Log" = 17, "PV Raw" = 23),  # Assign shapes
                     labels = c("PV Log" = "PV Log", "PV Raw" = "PV Raw")) +
  labs(title="Log10(Quantity) vs Raw Quantity Variability", y="Variability", x="DPI")
```
Use PV for quantity variability

Low dose birds are more variable in their pathogen loads during primary infection than high dose birds, but less variable
#####Secondary Infection

```{r Remove Birds That Did Not Resolve Infection By DPI 42}
#omit from dataset - were still infected (path load) on day 41 prior to reinfection day 42
m.ab <- m.ab %>%
  filter(!(band_number %in% c(2274, 2469, 2520, 2494)))

m.ab %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=dpi
  )%>%
  modify_header(
    label ~ "**FINAL SAMPLE SIZES SECONDARY**"
  )

m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose, inf_sec) %>%
  group_by(secondary_dose, primary_treatment) %>%
  tbl_summary()%>%
  modify_header(
    label ~ "**By Primary and Secondary Dose**"
  )
m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose)%>%
  tbl_summary(
    by=primary_treatment
  )
```
#####Do Antibody Levels During Primary Infection Predict Susceptibility to Reinfection?
```{r Set Theme for Susceptibility Graphs}
sus_theme_pre <- theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    axis.title.x = element_blank(),  # Remove the x-axis title
    axis.title.y = element_text(size = 15, face="bold"),
    plot.title = element_text(hjust = 0.5),  # Center the plot title
    legend.position = "top",
    legend.text = element_text(size = 15),
    legend.title = element_text(size = 15),
    legend.background = element_rect(size = 0.25, linetype = "solid"),
    strip.text = element_text(size = 10)  # Adjust the size of facet labels
  )

sus_theme_14 <- theme(
  axis.text.x = element_text(angle = 45, hjust = 1), 
  axis.title.y = element_text(size = 15, face="bold"),
  legend.position = "none",  # Remove the legend
  plot.title = element_blank(),  # Remove the plot title
  axis.title.x = element_blank(),  # Remove the x-axis title
  legend.text = element_blank(),  # Remove the legend text
  legend.title = element_blank(),  # Remove the legend title
  legend.background = element_blank(),  # Remove the legend background
  strip.text = element_text(size = 10)  # Adjust the size of facet labels
)

sus_theme_41 <- theme(
  axis.text.x = element_text(angle = 45, hjust = 1), 
  axis.title.y = element_text(size = 15, face="bold"),
  legend.position = "none",  # Remove the legend
  plot.title = element_blank(),  # Remove the plot title
  axis.title.x = element_text(size=15, face="bold"), 
  legend.text = element_blank(),  # Remove the legend text
  legend.title = element_blank(),  # Remove the legend title
  legend.background = element_blank(),  # Remove the legend background
  strip.text = element_text(size = 10)  # Adjust the size of facet labels
)
```

```{r Format Pre Antibodies}
m.ab.w <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, elisa_od, inf_sec)%>%
  filter(dpi %in% c(-8, 14, 41))

wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)

wibird$elisa_od_pre <- wibird$`elisa_od_-8`

wibird <- wibird %>% mutate(totalTES = sum())
wibird$sec_dose.n <- wibird$secondary_dose+1
wibird$log10.sec_dose <- round(log10(wibird$sec_dose.n), digits = 2)

wibird$log10.sec_dose <- as.numeric(wibird$log10.sec_dose)

wibird$secondary_dose_fct <- factor(wibird$secondary_dose, levels = c(0, 30, 100, 300, 7000))

sec_dose_names <- c(
  "0" = "0",
  "1.49" = "30",
  "2" = "100",
  "2.48" = "300",
  "3.85" = "7000"
)
```

```{r Is susceptibility to reinfection predicted by antibody levels 8 days before infection?}
glm.ab.pre <- glm(inf_sec ~ elisa_od_pre + log10.sec_dose, data=wibird, family=binomial())
summary(glm.ab.pre)
car::Anova(glm.ab.pre, type = 3)
simulateResiduals(glm.ab.pre, plot=T)

mod <- glm.ab.pre
dat.new=expand.grid(log10.sec_dose=unique(wibird$log10.sec_dose),
                    inf_sec=unique(wibird$inf_sec),
                    elisa_od_pre = unique(wibird$elisa_od_pre))#new grid to put predictions into
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

#names
wibird$dpi <- "DPPI -8"
dpi_namespre <- c(
  "DPPI -8" = "DPPI -8"
)

sus_pre_raw<-ggplot(wibird, aes(x=(elisa_od_pre), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_pre), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_pre, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(y=" ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi ~log10.sec_dose, labeller = as_labeller(c(dpi_namespre, sec_dose_names)))
sus_pre_raw
```

```{r Format DPI 14 Antibodies}
m.ab.w <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, elisa_od, inf_sec)%>%
  filter(dpi %in% c(-8, 14, 41))

wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)
wibird14 <- wibird %>%
  drop_na(elisa_od_14)

wibird14 <- wibird14 %>% mutate(totalTES = sum())
wibird14$sec_dose.n <- wibird14$secondary_dose+1
wibird14$log10.sec_dose <- round(log10(wibird14$sec_dose.n), digits = 2)

wibird14$log10.sec_dose <- as.numeric(wibird14$log10.sec_dose)
```

```{r Are Antibody Levels on Day 14 Predictive of Susceptibility?}
glm.ab14 <- glm(inf_sec ~ elisa_od_14 + log10.sec_dose, data=wibird14, family=binomial())
summary(glm.ab14)
car::Anova(glm.ab14, type = 3)
simulateResiduals(glm.ab14, plot=T)

mod <- glm.ab14
dat.new=expand.grid(log10.sec_dose=unique(wibird14$log10.sec_dose),
                    inf_sec=unique(wibird14$inf_sec),
                    elisa_od_14 = unique(wibird14$elisa_od_14))
# elisa_od_14 = unique(wibird14$elisa_od_14))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird14$dpi <- "DPPI 14"
dpi_names14 <- c(
  "DPPI 14" = "DPPI 14"
)

sus_14_raw<-ggplot(wibird14, aes(x=(elisa_od_14), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_line(data=dat.new, aes(x=(elisa_od_14), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_14, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  geom_point(alpha=0.5, size=2)+
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(y= "Infection Status Upon Secondary Challenge (0|1)", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi~log10.sec_dose, labeller = as_labeller(c(dpi_names14,sec_dose_names)))
sus_14_raw

```

```{r Format DPI 41 Antibodies}
wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)

wibird41 <- wibird %>%
  drop_na(elisa_od_41)

wibird41 <- wibird41 %>% mutate(totalTES = sum())
wibird41$sec_dose.n <- wibird41$secondary_dose+1
wibird41$log10.sec_dose <- round(log10(wibird41$sec_dose.n), digits = 2)

wibird41$log10.sec_dose <- as.numeric(wibird41$log10.sec_dose)
```

```{r Are Antibody Levels on Day 41 Predictive of Susceptibility to Reinfection?}
glm.ab41 <- glm(inf_sec ~ elisa_od_41 + log10.sec_dose, data=wibird41, family=binomial())

summary(glm.ab41)
car::Anova(glm.ab41, type = 3)
simulateResiduals(glm.ab41, plot=T)

mod <- glm.ab41
dat.new=expand.grid(log10.sec_dose=unique(wibird41$log10.sec_dose),
                    inf_sec=unique(wibird41$inf_sec),
                    #elisa_od_14_new=unique(wibird41$elisa_od_14_new),
                    elisa_od_41 = unique(wibird41$elisa_od_41))
dat.new$yhat=predict(mod, type="response", newdata = dat.new)
#prediction intervals
preds = predict(mod, type = "link", newdata = dat.new, se.fit =T)
#bind se's and fitted points
dat.new = cbind(dat.new, preds)
#inverse link function
ilink <- family(mod)$linkinv
#back transform CIs
dat.new <- transform(dat.new,
                     Fitted = ilink(fit),
                     Upper = ilink(fit + (2*se.fit)),
                     Lower = ilink(fit - (2*se.fit)))

wibird41$dpi <- "DPPI 41"
dpi_names41 <- c(
  "DPPI 41" = "DPPI 41"
)

sus_41_raw<-ggplot(wibird41, aes(x=(elisa_od_41), y=(inf_sec), color=as.factor(log10.sec_dose)))+
  geom_point(alpha=0.5)+
  geom_line(data=dat.new, aes(x=(elisa_od_41), y=yhat), alpha = 1, size =1)+
  geom_ribbon(data = dat.new, aes(ymin=Lower, ymax = Upper, x= elisa_od_41, y=yhat, fill=as.factor(log10.sec_dose)), alpha = 0.05, linetype="dashed", size=0.1) +
  scale_color_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  scale_fill_manual(values = c(sec_colors), labels=c("0", "30", "100", "300", "7000"))+
  labs(x="IgY Antibody Levels", y= " ", color="Secondary Dose", fill ="Secondary Dose")+
  scale_x_continuous(labels = scales::number_format(accuracy = 0.01))+
  facet_grid(~dpi~log10.sec_dose, labeller = as_labeller(c(dpi_names41,sec_dose_names)))
sus_41_raw
```

```{r}
library("grid")
library(cowplot)
r1 <- sus_pre_raw + sus_theme_pre
r2 <- sus_14_raw + sus_theme_14
r3 <- sus_41_raw + sus_theme_41
legend <-get_legend(r1)
r1 <- r1 + theme(legend.position = "none")

Raw_all <- arrangeGrob(
  legend,
  arrangeGrob(r1, r2, r3, ncol = 1),
  ncol = 1,
  heights = c(1, 10)
)

grid.newpage()
grid.draw(Raw_all)
```

####Do antibody levels in susceptible birds predict pathogen load or eye score?
```{r Format wide data}
m.ab.w <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, elisa_od, inf_sec, inf_pri)%>%
  filter(dpi %in% c(-8, 14, 41, 56))

wibird <- m.ab.w %>% pivot_wider(names_from = dpi,
                                 names_glue = "{.value}_{dpi}",
                                 values_from = elisa_od)

m.ab.w.tes <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, tes, inf_sec, inf_pri)%>%
  filter(dpi %in% c(-8, 14, 41, 56))
wibird.tes <- m.ab.w.tes %>% pivot_wider(names_from = dpi,
                                     names_glue = "{.value}_{dpi}",
                                     values_from = tes)

m.ab.w.quant <- m.ab %>% 
  dplyr::select(dpi, band_number, primary_treatment, secondary_dose,
                sex, quantity1, inf_sec, inf_pri)%>%
  filter(dpi %in% c(-8, 14, 41, 56))
wibird.quant <- m.ab.w.quant %>% pivot_wider(names_from = dpi,
                                         names_glue = "{.value}_{dpi}",
                                         values_from = quantity1)

wibird.all <- left_join(wibird, wibird.quant, by="band_number")
wibird.all <- left_join(wibird.all, wibird.tes, by="band_number")
wibird.all <- wibird.all %>%
  select(-primary_treatment.y, -secondary_dose.y, -sex.y, -inf_sec.y, -inf_pri.y, -primary_treatment.x, -secondary_dose.x, -sex.x, -inf_sec.x, -inf_pri.x)
head(wibird.all)
```
Antibody levels do not predict path load or eye score dpi 56
```{r Do antibody levels on dpi 41 predict path load or eye score dpi 56?}
wibird.all$quantity1_56.log <- log10(wibird.all$quantity1_56)
wibird.all.41 <- wibird.all %>%
  drop_na(elisa_od_41)

glm.ab.quant <- glmmTMB(quantity1_56.log+.01 ~ elisa_od_41 + primary_treatment, data=wibird.all.41, family=Gamma())
summary(glm.ab.quant)

glm.ab.tes <- glmmTMB(tes_56+0.001 ~ elisa_od_41 + primary_treatment, data=wibird.all.41, family = Gamma())
summary(glm.ab.tes)
hist(wibird.all.41$tes_56)
ggplot(wibird.all, aes(x=elisa_od_41, y=quantity1_56.log, color=primary_treatment))+
  geom_point()+
  scale_color_manual(values=c(pri_colors))+
  geom_smooth(aes(groups=primary_treatment), method="glm")

```
Antibody levels do not predict path load or eye score dpi 56
```{r Do antibody levels on dpi 14 predict path load or eye score dpi 56?}
wibird.all.14 <- wibird.all %>%
  drop_na(elisa_od_14)

glm.ab.quant <- glmmTMB(quantity1_56.log+.01 ~ elisa_od_14 + primary_treatment, data=wibird.all.14, family=Gamma())
summary(glm.ab.quant)

glm.ab.tes <- glmmTMB(tes_56+0.001 ~ elisa_od_14 + primary_treatment, data=wibird.all.14, family = Gamma())
summary(glm.ab.tes)
```

####Variability combined with Hawley et al., 2024 Susceptibility####
```{r}
####Variability With Hawley et al., 2024####
summary_df <- a.cv %>%
  filter(dpi.f == 41) %>%
  group_by(primary_treatment) %>%
  summarize(
    mean_bird_pv = mean(bird_pv, na.rm = TRUE),
    mean_bird_cv = mean(bird_cv, na.rm = TRUE),
    mean_mean_od = mean(mean_od, na.rm = TRUE)
  )

summary_df
dose <- c("Sham", "Low", "High")
CV <- as.numeric(c(0.899, 1.630, 2.511))
ab.pv <- as.numeric(c(0.04156784, 0.14716858, 0.18885516))
ab.cv <- as.numeric(c(0.03997555, 0.17202405, 0.26176249))
ci_lower_cv <- as.numeric(c(0.0230, 0.119, 0.165))
ci_upper_cv <- as.numeric(c(0.0409, 0.206, 0.339))
ci_lower_pv <- as.numeric(c(0.0246, 0.0998, 0.144))
ci_upper_pv <- as.numeric(c(0.0452, 0.178, 0.231))
mean_sus <- as.numeric(c(1.181, 0.446, 0.192))
mean_ab <- as.numeric(c(0.045, 0.04786, 0.0579))
dpi <- 41



het.df <- data.frame(dose, CV, mean_sus, ab.pv, ab.cv, dpi, mean_ab, ci_lower_cv, ci_upper_cv, ci_lower_pv, ci_upper_pv)

het.df$mean_sus_inv <- 1/het.df$mean_sus

#hawley et al 2024
het.paper.only <- ggplot(het.df, aes(x=Metric))+
  geom_point(size=5,aes(x="CV", y=CV, color=dose), shape = 17)+
  geom_point(size=5,aes(x="CV", y=CV, shape = "CV"), color="black", shape=2, stroke = 1)+
  scale_color_manual(values=c(pri_colors))+
  scale_y_continuous(
    name= "Variability in Susceptibility Hawley et al. 2024"
  )+
  theme_minimal()+
  theme(
    axis.title.y = element_text(color = "black", size=12, face="bold"),
    legend.position = "none"
  ) +
  labs(color = "Primary Treatment", shape= "Variability Metric")
het.paper.only

ab.paper.only <- ggplot(het.df, x=Metric)+
  geom_point(size=5,aes(x="PV", y=ab.pv, color=dose, shape="PV"))+
  geom_errorbar(aes(ymin=ci_lower_pv, ymax=ci_upper_pv, y=ab.pv, x="PV", color=dose), width=0.1)+
  geom_point(size=5,aes(x="CV", y=ab.cv, color=dose, shape = "CV"))+
  geom_errorbar(aes(ymin=ci_lower_cv, ymax=ci_upper_cv, y=ab.cv, x="CV", color=dose), width=0.1)+
  #geom_point(size=3.6,aes(x="PV", y=ab.pv), color="black", shape=1, stroke = 2)+
  scale_color_manual(values=c(pri_colors))+
  scale_y_continuous(
    name="Variability in Antibody Levels DPPI 41"
  )+
  scale_x_discrete(
    name="Metric"
  )+
  scale_shape_manual(values=c(17,19))+
  theme_minimal()+
  theme(
    axis.title.y = element_text(color = "black", size=12, face = "bold"),
    legend.position="top",
    legend.direction = "horizontal",
    legend.box = "vertical",
    legend.background= element_rect(size=0.25, linetype="solid"))+
  labs(color = "Primary Treatment", shape= "Variability Metric")
ab.paper.only

legend <- cowplot::get_legend(ab.paper.only)

# Remove legends from plots
het.paper.only <- het.paper.only + theme(legend.position = "none")
ab.paper.only <- ab.paper.only + theme(legend.position = "none")

#FIGURE 4
Fig4 <- grid.arrange(legend, 
             arrangeGrob(het.paper.only, ab.paper.only, ncol = 2),
             ncol = 1,
             heights = c(1, 4))
```

####Variability in Reinfected Birds####
```{r Secondary Sample Sizes}
sec.ab <- m.ab %>%
  filter(inf_sec == 1 & dpi > 41 & secondary_dose == 7000)

m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose, inf_sec) %>%
  group_by(secondary_dose, primary_treatment) %>%
  tbl_summary()%>%
  modify_header(
    label ~ "**By Primary and Secondary Dose**"
  )

m.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(inf_sec, primary_treatment, secondary_dose, band_number) %>%
  group_by(inf_sec, secondary_dose, primary_treatment) %>%
  tbl_summary(
    by = inf_sec, # Grouping by infected secondary
    statistic = list(band_number ~ "{n} ({p}%)") # Count occurrences and show percentages
  ) %>%
  modify_header(label ~ "**Infected Secondary**") 

sec.ab %>%
  filter(dpi == 56) %>%
  dplyr::select(dpi, primary_treatment, secondary_dose, inf_sec) %>%
  group_by(secondary_dose, primary_treatment) %>%
  tbl_summary()%>%
  modify_header(
    label ~ "**By Primary and Secondary Dose**"
  )
```

```{r Max quantity and eye score secondary}
max.quant.sec <- sec.ab %>%
  filter(dpi > 41 & inf_sec ==1) %>%
  group_by(band_number, primary_treatment, secondary_dose) %>%
  reframe(max_quantity = max(quantity1, na.rm = TRUE))

max.tes.sec <- sec.ab %>%
  group_by(band_number, primary_treatment, secondary_dose)%>%
  reframe(max_tes = max(tes, na.rm=TRUE))

max.sec<-left_join(max.quant.sec, max.tes.sec, by="band_number")

max.sec$primary_treatment <- max.sec$primary_treatment.x
max.sec$secondary_dose <- max.sec$secondary_dose.x

max.sec <- max.sec %>%
  dplyr::select(-c(primary_treatment.y, secondary_dose.y, primary_treatment.x, secondary_dose.x))
max.sec$max_quantity1 <- max.sec$max_quantity+1

```

```{r}
# Pipeline to calculate CV, PV, V2, and their 95% confidence intervals for max_quantity and max_tes

#add small constant for variability calculations
sec.var$max_tes <- sec.var$max_tes + 0.001

sec.var <- max.sec %>%
  group_by(primary_treatment) %>%
  drop_na(max_quantity, max_tes) %>%
  dplyr::reframe(
    # For max_quantity
    max_quantity = max_quantity,
    mean_quantity = mean(max_quantity),
    median_quantity = median(max_quantity),
    bird_sd_quantity = sd(max_quantity),
    band_number = band_number,
    bird_cv_quantity = calculate_cv(max_quantity),  # Calculate CV for max_quantity
    bird_pv_quantity = calculate_pv(max_quantity),  # Calculate PV for max_quantity
    bird_v2_quantity = calculate_v2(max_quantity),  # Calculate V2 for max_quantity
    
    # Bootstrap for each metric for max_quantity
    cv_bootstrap_quantity = list(replicate(n_boot, calculate_cv(sample(max_quantity, replace = TRUE)))),
    pv_bootstrap_quantity = list(replicate(n_boot, calculate_pv(sample(max_quantity, replace = TRUE)))),
    v2_bootstrap_quantity = list(replicate(n_boot, calculate_v2(sample(max_quantity, replace = TRUE)))),

    # For max_tes
    max_tes = max_tes,
    mean_tes = mean(max_tes),
    median_tes = median(max_tes),
    bird_sd_tes = sd(max_tes),
    bird_cv_tes = calculate_cv(max_tes),  # Calculate CV for max_tes
    bird_pv_tes = calculate_pv(max_tes),  # Calculate PV for max_tes
    bird_v2_tes = calculate_v2(max_tes),  # Calculate V2 for max_tes
    
    # Bootstrap for each metric for max_tes
    cv_bootstrap_tes = list(replicate(n_boot, calculate_cv(sample(max_tes, replace = TRUE)))),
    pv_bootstrap_tes = list(replicate(n_boot, calculate_pv(sample(max_tes, replace = TRUE)))),
    v2_bootstrap_tes = list(replicate(n_boot, calculate_v2(sample(max_tes, replace = TRUE))))
  ) %>%
  # Calculate 95% confidence intervals for CV, PV, and V2 for both max_quantity and max_tes
  mutate(
    # For max_quantity
    cv_lower_ci_quantity = map_dbl(cv_bootstrap_quantity, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci_quantity = map_dbl(cv_bootstrap_quantity, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci_quantity = map_dbl(pv_bootstrap_quantity, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci_quantity = map_dbl(pv_bootstrap_quantity, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci_quantity = map_dbl(v2_bootstrap_quantity, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci_quantity = map_dbl(v2_bootstrap_quantity, ~ quantile(.x, 0.975, na.rm = TRUE)),

    # For max_tes
    cv_lower_ci_tes = map_dbl(cv_bootstrap_tes, ~ quantile(.x, 0.025, na.rm = TRUE)),
    cv_upper_ci_tes = map_dbl(cv_bootstrap_tes, ~ quantile(.x, 0.975, na.rm = TRUE)),
    pv_lower_ci_tes = map_dbl(pv_bootstrap_tes, ~ quantile(.x, 0.025, na.rm = TRUE)),
    pv_upper_ci_tes = map_dbl(pv_bootstrap_tes, ~ quantile(.x, 0.975, na.rm = TRUE)),
    v2_lower_ci_tes = map_dbl(v2_bootstrap_tes, ~ quantile(.x, 0.025, na.rm = TRUE)),
    v2_upper_ci_tes = map_dbl(v2_bootstrap_tes, ~ quantile(.x, 0.975, na.rm = TRUE))
  ) %>%
  # Optionally, remove the bootstrap columns to clean up the dataset
  select(-cv_bootstrap_quantity, -pv_bootstrap_quantity, -v2_bootstrap_quantity,
         -cv_bootstrap_tes, -pv_bootstrap_tes, -v2_bootstrap_tes) %>%
  ungroup()

# Merge back with the original dataset and cleanup
max.sec.v <- left_join(max.sec, sec.var, by = c("primary_treatment", "band_number"))

# Summarize and print the final table
summary_tibble <- sec.var %>%
  group_by(primary_treatment) %>%
  summarize(
    mean_bird_cv_quantity = mean(bird_cv_quantity, na.rm = TRUE),   # Mean CV for max_quantity
    mean_bird_sd_quantity = mean(bird_sd_quantity, na.rm = TRUE),   # Mean SD for max_quantity
    mean_bird_pv_quantity = mean(bird_pv_quantity, na.rm = TRUE),   # Mean PV for max_quantity
    mean_bird_v2_quantity = mean(bird_v2_quantity, na.rm = TRUE),
    
    mean_bird_cv_tes = mean(bird_cv_tes, na.rm = TRUE),             # Mean CV for max_tes
    mean_bird_sd_tes = mean(bird_sd_tes, na.rm = TRUE),             # Mean SD for max_tes
    mean_bird_pv_tes = mean(bird_pv_tes, na.rm = TRUE),             # Mean PV for max_tes
    mean_bird_v2_tes = mean(bird_v2_tes, na.rm = TRUE),
    
    avg_max_quantity = mean(max_quantity, na.rm = TRUE),            # Average max_quantity for each group
    median_max_quantity = median(max_quantity, na.rm=TRUE),
    avg_max_tes = mean(max_tes, na.rm = TRUE),                      # Average max_tes for each group
    median_max_tes = median(max_tes, na.rm=TRUE),
    n_individuals = n_distinct(band_number)                         # Number of individuals in each group
  ) %>%
  as_tibble()

# Print the summary table
print(summary_tibble)

```

```{r Maximum Pathogen Load and Variability}
ggplot(summary_tibble)+
  geom_point(aes(x="CV", y= mean_bird_cv_quantity, color=primary_treatment))+
  geom_point(aes(x="PV", y= mean_bird_pv_quantity, color=primary_treatment))+
  geom_point(aes(x="V2", y= mean_bird_v2_quantity, color=primary_treatment))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Pathogen Load Variability", y="Variability", x="Metric")

ggplot(max.sec.v)+
  geom_point(aes(x="CV", y= bird_cv_quantity, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="CV", ymin=cv_lower_ci_quantity, ymax=cv_upper_ci_quantity, y=bird_cv_quantity, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="PV", y= bird_pv_quantity, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="PV", ymin=pv_lower_ci_quantity, ymax=pv_upper_ci_quantity, y=bird_pv_quantity, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="V2", y= bird_v2_quantity, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="V2", ymin=v2_lower_ci_quantity, ymax=v2_upper_ci_quantity, y=bird_v2_quantity, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Pathogen Load Variability", y="Variability", x="Metric", color="Primary Treatment")

ggplot()+
  geom_jitter(data=sec.var, aes(x=primary_treatment, y=max_quantity, color=primary_treatment), width=0.05, size=2, alpha=0.5)+
  geom_point(data=summary_tibble, aes(x=primary_treatment, y=median_max_quantity, color=primary_treatment), shape=18, size=3)+
  geom_point(data=summary_tibble, aes(x=primary_treatment, y=avg_max_quantity, color=primary_treatment), shape=17, size=3)+
  scale_color_manual(values=c(pri_colors))+
  labs(y="Max Pathogen Load", x="Primary Treatment", color="Primary Treatment", title = "Maximum Pathogen Load of Reinfected Birds")
```

```{r}
ggplot(summary_tibble)+
  geom_point(aes(x="CV", y= mean_bird_cv_tes, color=primary_treatment))+
  geom_point(aes(x="PV", y= mean_bird_pv_tes, color=primary_treatment))+
  geom_point(aes(x="V2", y= mean_bird_v2_tes, color=primary_treatment))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Eye Score Variability", y="Variability", x="Metric")

ggplot(max.sec.v)+
  geom_point(aes(x="CV", y= bird_cv_tes, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="CV", ymin=cv_lower_ci_tes, ymax=cv_upper_ci_tes, y=bird_cv_tes, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="PV", y= bird_pv_tes, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="PV", ymin=pv_lower_ci_tes, ymax=pv_upper_ci_tes, y=bird_pv_tes, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
  geom_point(aes(x="V2", y= bird_v2_tes, color=primary_treatment), position=position_dodge(width=0.5))+
  geom_errorbar(aes(x="V2", ymin=v2_lower_ci_tes, ymax=v2_upper_ci_tes, y=bird_v2_tes, color=primary_treatment), width=0.2, position=position_dodge(width=0.5))+
    scale_color_manual(values=c(pri_colors))+
  scale_shape_manual(name = "Variability Metric",
                     values = c("CV" = 17, "PV" = 19, "V2" = 18),  # Assign shapes
                     labels = c("CV" = "CV", "PV" = "PV", "V2" = "CV^2")) +
  labs(title="Secondary Eye Score Variability", y="Variability", x="Metric", color="Primary Treatment")

ggplot()+
  geom_jitter(data=sec.var, aes(x=primary_treatment, y=max_tes, color=primary_treatment), width=0.1, height=0, size=2, alpha=0.5)+
  geom_point(data=summary_tibble, aes(x=primary_treatment, y=avg_max_tes, color=primary_treatment), shape=17, size=3)+
  scale_color_manual(values=c(pri_colors))+
  labs(y="Max Eye Score", x="Primary Treatment", color="Primary Treatment", title = "Maximum Eye Score of Reinfected Birds")
```

